### 2023-02-28 ucIrvine/presentation.tex
Feb 27

\documentclass{beamer}
\usepackage{graphicx}
\title{Uncertainties in precipitation
estimates from space-borne radar observations}
\author{Mircea Grecu}
\institute{NASA Goddard Space Flight Center and Morgan State University}
\date{March 2nd,2023}

\begin{document}
\frame{\titlepage}
\begin{frame}
\frametitle{Sources of uncertainties}
\begin{columns}
%%in the precipitation estimates from space-borne radar observations
\begin{column}{0.45\textwidth}
    Major sources of uncertainty include:
    \begin{itemize}
        \item Variability in the Z-R relationships
        \item Attenuation in the observed radar reflectivity
        \item Variability in the radar beam footprint
        \item Ground clutter
    \end{itemize}
\end{column}
\begin{column}{0.55\textwidth}  %%<--- here
    \begin{figure}
    \begin{center}
     \includegraphics[width=0.9\textwidth]{Figures/fig1.png}
     \end{center}
    \end{figure}
\end{column}
\end{columns}
\end{frame}
\begin{frame}
\frametitle{Variability in the Z-Precipitation Rate relationships}
\begin{itemize}
    \item At least 2 parameters are required to describe
    Particle size distributions (PSDs).
    \item Consequently, the Z-Precipitation Rate relationships require at
    least 2 parameters. The generalized PSD intercept, $N_w=\frac {4^4} {\pi \rho_w} \frac {PWC} {D_m^4}$,
    greatly simplify Z-Rate relationships.
\end{itemize}
\begin{figure}
\begin{center}
\includegraphics[width=0.9\textwidth]{Figures/fig2.png}
\end{center}
\end{figure}
\end{frame}
\begin{frame}
\frametitle{The $N_w$ problem. General considerations}
\begin{columns}
    \begin{column}{0.5\textwidth}
        \begin{itemize}
            \item The $N_w$ parameter greatly simplifies the formulation, but
            not the problem.
            \item It does not solve it though, as $N_w$ still needs to
            estimated independently of the radar observations.
            \item Ground observations may be used to impose constraints on
            $N_w$.
            \item However, such constraints are not very effective for 
            PSD characterized by large mean particle sizes $D_m$.
        \end{itemize}
    \end{column}
    \begin{column}{0.5\textwidth}
        \begin{figure}
            \begin{center}
            \includegraphics[width=0.9\textwidth]{Figures/fig3.png}
            \end{center}
        \end{figure}
        \begin{figure}
            \begin{center}
            \includegraphics[width=0.9\textwidth]{Figures/fig4.png}
            \end{center}
        \end{figure}
    \end{column}
\end{columns}
\end{frame}
\begin{frame}
\frametitle{The $N_w$ problem. Radar profiling considerations}
\begin{columns}
    \begin{column}{0.55\textwidth}
    \begin{itemize}
        \item Additional constraints can be imposed by using radar
        by considering $N_w$ in the radar profiling context.
        \item That is, the attenuation correction process needs to be
        consistent with the $N_w$ parameter.
        $Z(r)=Z_m(r) /PIA$
        $PIA=(1-\epsilon (N_w) q\int_0^r Z_m^\beta(s)ds)^{1/\beta}$
        \item The analytical PIA needs to be consistent with the
        SRT PIA estimate and Ka-band reflectivity observations 
        when available.
    \end{itemize}
    \end{column}
    \begin{column}{0.5\textwidth}
        \begin{figure}
            \begin{center}
            \includegraphics[width=0.9\textwidth]{Figures/fig5.png}
            \end{center}
    \end{figure}
    \end{column}
\end{columns}
\end{frame}
\begin{frame}
\frametitle{The $N_w$ problem. Further considerations}
\begin{itemize}
\item Surface Reference Technique (SRT) PIA is a useful piece of information when reliable.
\item Dual SRT PIA is more reliable than single SRT PIA, except for heavy convection.
\item Over oceans SRT PIA estimates are significantly more reliable than over land.
\item For snow and light rain, SRT PIA estimates are not reliable.
\end{itemize}
\begin{figure}
\begin{center}
\includegraphics[width=0.85\textwidth]{Figures/fig6.png}
\end{center}
\end{figure}
\end{frame}
\begin{frame}
\frametitle{Snow estimation issues}
\begin{columns}
\begin{column}{0.58\textwidth}
    \begin{itemize}
    \item The variability of $N_w$ is still a problem, but there is very little
    attenuation in snow to use the PIA as a constraint.
    \item When dual frequency observations are available, $D_m$ and hence
    $N_w$ may be estimated from the dual frequency reflectivity ratio (DFR).
    \item However, the DFR may be very noisy.
    \item More robust methods, i.e. Machine-Learning based, may be used 
    to derive to incorporate "in-situ" information and provide
    more accurate estimates.
    \end{itemize}
\end{column}
\begin{column}{0.5\textwidth}
\begin{figure}
\begin{center}
\includegraphics[width=0.9\textwidth]{Figures/fig7.png}
\includegraphics[width=0.9\textwidth]{Figures/Fig8.png}
\end{center}
\end{figure}
\end{column}
\end{columns}
\end{frame}    
\end{document}

### 2023-02-28 ucIrvine/presentation.tex
updates Feb 28

@@
 \institute{NASA Goddard Space Flight Center and Morgan State University}
-\date{March 2nd,2023}
+\date{March 2,2023}
+%%\date{\today}
 
@@
 \end{frame}    
+\begin{frame}
+\frametitle{Ground Clutter}
+\begin{columns}
+\begin{column}{0.55\textwidth}
+    \begin{itemize}
+    \item Ground clutter is strong echo in the radar observation caused 
+    by the ground.
+    \item It can extend up to 2.0 km above the surface and completely 
+    obscure the precipitation echo.
+    \item Nadir observations are minimally affected 
+    and can be used in the development of statistical clutter mitigation 
+    methodologies, e.g.
+    $pRate(z<z_{cf})=pRate(z_{cf}) \times$
+    $pRate_{mean}(z)/pRate_{mean}(z_{cf})$
+    \end{itemize}
+\end{column}
+\begin{column}{0.5\textwidth}
+    \begin{figure}
+        \begin{center}
+        \includegraphics[width=0.9\textwidth]{Figures/fig9.png}
+        \includegraphics[width=0.9\textwidth]{Figures/fig10.png}
+        \end{center}
+    \end{figure}
+\end{column}
+\end{columns}
+\end{frame}
+\begin{frame}
+\frametitle{Impact of Ground Clutter Correction}
+\begin{figure}
+        \begin{center}
+        \includegraphics[width=0.9\textwidth]{Figures/fig11.png}
+        \end{center}
+    \end{figure}
+\end{frame}
+\begin{frame}
+    \frametitle{Light Precipitation}
+    \begin{columns}
+        \begin{column}{0.55\textwidth}
+            \begin{itemize}
+            \item The DPR does not miss all the light precipitation associated with a given point in the Tb-space
+            \item An empirical algorithm may be derived from collocated Tbs and DPR retrievals
+            \item The empirical algorithm is to be applied only when the DPR does not detect precipitation. When estimates are greater than 0, a decision is required.
+
+            \end{itemize}
+        \end{column}
+        \begin{column}{0.5\textwidth}
+            \begin{figure}
+                \begin{center}
+                \includegraphics[width=0.9\textwidth]{Figures/fig12.png}
+                \includegraphics[width=0.9\textwidth]{Figures/fig13.png}
+                \end{center}
+            \end{figure}
+        \end{column}
+    \end{columns}
+\end{frame} 
+\begin{frame}
+    \frametitle{Summary and Conclusions}
+    \begin{itemize}
+        \item Although the most accurate at the instantaneous level, space-borne radar precipitation 
+        estimates affected by multiple types of uncertainties.
+        \item Even when dual frequency observations are available, the 
+        estimation problem is still posed.
+        \item Parameterizations and "a priori" information derived 
+        from ground observations may be used to mitigate the uncertainties.
+        \item However, the process is not trivial and requires a sustained
+        long-term effort.
+    \end{itemize}
+\end{frame}
 \end{document}

### 2023-05-02 RST/responseReviewer2.tex
revision Reviewer_2

\documentclass[12pt]{article}
\usepackage{amsmath,amssymb,amsthm}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{subcaption}
\author{Mircea Grecu and John E. Yorks}
\title{Response to Reviewers}
\date{}
\begin{document}
\maketitle

\noindent
\textit{ \textbf{General Summary:}
This a well-written paper that fits within the scope of Journal of Atmospheric and 
Oceanic Technology and presents important and exciting results. The authors construct
 profiles of ice clouds based on CloudSat observations and then simulate retrievals 
 based on the instruments of the planned AOS mission. There are, however, 
 a few important ways in which the paper could be strengthened. I recommend 
 publication after the following comments are addressed.}\\
\newline
 We appreciate \\
\newline
\textit{\textbf{Major Comments:} 1) There are several large sources of uncertainty that will be present in real 
retrievals from AOS that are ignored in this paper. While there is certainly still 
value in such analysis, it needs to be made more clear in the paper that this is an 
idealized scenario, and there should be more discussion about the sources of 
uncertainty that will affect real-world retrievals. Several in particular stand 
out to me. For one, it appears that all of the simulated measurements have been 
created at the same resolution (presumably, the resolution of CloudSat). This
should be clarified, and the authors should discuss how the lidar, radar, and 
radiometer will all have different resolutions and how this might affect results. 
Second, the electromagnetic scattering properties assumed in this study are 
unlikely to perfectly match reality. Finally, it appears that no measurement 
uncertainty has been assumed. The last paragraph of the paper does briefly mention 
many of these uncertainties, but they need to be introduced earlier in the 
manuscript, and more details are needed about the specifics of how IMPACTS
will help mitigate them.}

2) You say in your abstract (and conclusion) that “the characteristics of 
the instruments e.g.,  frequencies, sensitivities, etc. are set based on the 
expected characteristics of the AOS mission.” However, the AOS mission 
specifications are still very much in flux. Therefore, for reproducibility, 
I suggest adding a table explicitly laying out the frequencies and 
sensitivities you assume for each instrument (radar, lidar, radiometer). 
Some of these (e.g., the radiometer frequencies) are mentioned in the text, 
but it would be nice to see it all in one place.

Minor Comments

Line 38: Where is 8.0 dBZ coming from? There should be a citation here, or it should be stated that this is an estimate of the radar’s expected sensitivity.

Lines 50-51: While your approach is justifiable and worthwhile, it is worth noting that your CloudSat IWC retrieval also makes numerous assumptions about ice cloud microphysical properties and structures (as laid out in Section 2a). So the problem is not entirely avoided by relying on “observations” instead of models, as the observations were of radar reflectivity, not of ice water contents or particle sizes.

Lines 65-67: This is fine, but it should be noted that in doing this, you are underestimating the true retrieval uncertainty. The particle distribution assumptions and electromagnetic scattering properties that you use in training the retrieval algorithm will almost certainly not perfectly reflect the actual size and shape distributions of ice particles around the globe.

Lines 83-90 and Eq. (1): I looked over the Testud et al. (2001) paper, and could not find anything that directly supported this statement, or any equations that resembled Eq. (1). Please add further explanation, either here or in an appendix, to help the reader follow your logic.

Figure 2: Neither the caption nor the text adequately explains what is being plotted in this figure. What does each individual point represent? The text says, “Figure 2 is a scatter plot representation of the class multiplicative coefficient ai as a function of relative height,” but if that were the case, I would expect there to be many more points on the scatter plot (i.e. 18 classes times a certain number of relative height levels, to yield hundreds of points).

Lines 120-121: How is the data in Fig. 2 used to parameterize Nw as a function of height? Do you fit a line/curve to the data in Fig. 2? This needs to be clarified.

Figure 3: It would be helpful to add vertical lines to these plots to indicate the detection thresholds you are assuming for the radar (8 dBZ?) and lidar.

Line 220: It is my understanding that an ensemble Kalman smoother assumes normal distributions. How much does it matter that IWC is not normally distributed in nature?

Typos

Line 276: “wells” should be “well”

Line 308: Figure 8 should be Figure 9

Line 315: IWP should be IWC

Line 361: eliminate “out

\pagebreak
\noindent \textbf{Reviewer 2:}\textit{The study investigates the synergy between observations from elastic backscatter lidar, Ku-band radar,
and sub-millimeter radiometer observations to retrieve ice concentrations. The topic is undoubtedly timely,
considering that AOS, NASA's upcoming satellite mission targeting remote sensing of the atmosphere,
is currently in its design phase. The methodology is sound, albeit slightly contrived. Methods and
results are presented in an accessible way.
Nonetheless, there are significant conceptual flaws in the study.}\\
\newline
We thank the reviewer for their perspective and constructive suggestions.\\

\noindent \textit{First of all, I can't entirely agree with the study's framing. Currently, 
the study investigates the synergies
achieved by adding lidar and radar observations to radiometer observations. 
However, since the active sensors will likely have a much narrower swath, 
combining these observations will only be possible at
a small part of the radiometer's swath. Given the significantly higher information 
content of the active observations, the more relevant question is: What is the 
benefit of including radiometer observations
when radar and lidar observations are available? Simply showing an added value 
from adding radar
and/or lidar observations to radiometer observations, which is the current main 
conclusion of the paper, is a trivial result.}\\
\newline
Although active observations are characterized by much narrower swaths and 
significantly worse temporal sampling, they are nevertheless crucial in process 
studies and the development and validation of products from radiometer-only
observations (\cite{stephens2008},\cite{sano2022},\cite{stubenrauch2021}).  The
particular combination of active and passive observations considered in this
study is justified by the current design of the AOS mission. Specifically,
the Ku-band radar is intended to enable accurate estimates of 
the convective precipitation.  Although it was originally envisioned that AOS would
deploy two Ku-band radars, one in an inclined orbit and one in a polar orbit, only
the inclined orbit Ku-band radar is currently planned for the AOS mission.  
The polar orbit radar will operate at higher-frequencies (Ka- or W-and) and 
have cloud-detecting capabilities (i.e. high-sensitivity), but
will not be able to provide direct observations of intense convective precipitation
due to attenuation. Moreover, the polar radar will have nadir-only capabilities,
while the inclined orbit radar (which will be contributed by JAXA) will have a
swath of about 200 km.  Both the polar and inclined orbit radars will operate
in tandem with a lidar and a high-frequency radiometer.  We are interested
in the inclined orbit despite the significantly lower sensitivity of the 
Ku-band radar because the synergistic Ku-band radar, elastic lidar and radiometer
estimates of ice in anvil clouds can be studied jointly with various parameters
(i.e. precipitation intensity, vertical velocities, etc.) characterizing the
convective processes responsible for the creation of the anvil clouds.  Although 
the polar orbit instruments are likely to enable the derivation of more accurate 
synergistic ice estimates, their severely limited ability to quantify convection
makes synergistic ice retrievals from the combination of instruments studied in
this manuscript relevant to the problem of relating the strength of convection to
the horizontal extent of anvil clouds and their impact on climate \cite{hartmann2016}.
Although the improvement in the ice estimates due to the inclusion of radiometer 
observations may seem trivial, it is rather significant given that neither the
lidar nor the Ku-band radar can provide by themselves accurate and complete 
ice estimates, the lidar
due to attenuation and the radar due to its limited sensitivity. \\

\noindent \textit{Secondly, since the dependency of the multiplicative factor a is parametrized by relative height,
the authors essentially use a single-moment retrieval to determine the profile database upon
which they base their whole study. This approach will significantly underestimate the variability
in the retrieved PSDs. Pfreundschuh et al. 2020 show that this variability causes considerable
uncertainty in radar-only retrievals of ice water path and is one of the dimensions in which
radiometer observations can help to improve radar-only retrievals. Neglecting this variability
likely significantly overestimates the accuracy of the proposed retrievals.
Finally, I do not consider that the omission of seven ($!$) of the ten radiometer channels is justified,
given that the authors use the inability of the radiometer to reproduce the vertical distribution of
ice particles as a principal motivation for studying the synergy with active observations.}\\
\newline
We agree with the reviewer. The parametrization of the multiplicative factor $a$ is equivalent to a single-moment retrieval. In the revised
manuscript, to investigate the impact of variability in $N_w$, we randomly perturbed its parameterized value by a log-normally distributed
random variable with a standard deviation of 0.5. This was achieved by generating one-dimensional vectors of independent random variables
with a standard normal distribution and applying a Gaussian smoothing filter \cite{nixon2019}.  The filter size was chosen to be two CloudSat radar bins.
The initial standard deviation was scaled to result a standard deviation of 0.5 after smoothing, which is roughly consistent with the variability
in $N_w$ observed in real observations \cite{grecu2018}. \\
\newline 
Moreover, the revised manuscript includes all radiometer channels in the analysis. It should be noted though that the 183.31 and 321.15-GHz radiometer channels
are centered on water vapor absorption lines. The differences in the brightness temperatures observed by these channels are driven by the vertical distribution
of water vapor and have only indirect impact on the retrieval of ice.  This is particularly true for the simulated observations considered in this study, which
are not characterized by large variability in the vertical distribution of water vapor. Consequently, the inability of the radiometer to reproduce the vertical
distribution of ice particles is not likely related to the omission of these channels. We nevertheless included all channels in the retrievals, as the vertical
distribution of moisture is likely to be important in future versions of the methodology when more diverse environmental contexts are going to be considered.\\
\newline
\textit{\textbf{Major comments:}}
\textit{1. To address the first-mentioned issue above, the authors should include radar-only, lidar-only, and radar-lidar 
    retrieval in their analysis.}\\
\newline
We agree with the reviewer. The revised manuscript includes radar-only, lidar-only, and radar-lidar retrievals in the analysis.  Shown 
in Fig. \ref{fig:IWP_IWC} are density plots of single-instrument retrievals of IWP and IWC as a function of their reference values.
As apparent in the figure, retrievals from single active instruments exhibit significant deficiencies.  This is not surprising given that
the lidar observations are subject to significant attenuation, while a Ku-radar with the AOS specifications misses a significant fraction of
the ice clouds. A W-band cloud radar would do much better by itself, but as already stated in our response, a Ku-band radar in conjunction
with a backscatter lidar and a high-frequency radiometer is a very exciting prospect, as it enables the study of the relation between 
the strength of convection and the characteristics of associated anvil clouds. \\
%include figure
\begin{figure}
\centering
\includegraphics[width=.31\linewidth]{IWP_LidarOnly.png}
\includegraphics[width=.31\linewidth]{IWP_RadarOnly.png}
\includegraphics[width=.31\linewidth]{IWP_RadiometerOnly.png}
\includegraphics[width=.31\linewidth]{IWC_LidarOnly.png}
\includegraphics[width=.31\linewidth]{IWC_RadarOnly.png}
\includegraphics[width=.31\linewidth]{IWC_RadiometerOnly.png}
\caption{Top Row: Density plots of IWP retrievals from lidar-only (left), radar-only (middle), and radiometer-only (right) observations as a function
of true IWP. Bottom Row: Same as in the top row but for IWC.}
\label{fig:IWP_IWC}
\end{figure}

\noindent
\textit{2. To improve the representation of microphysical variability in the retrieval, I suggest that the authors at 
    least randomize the parametrization of the multiplicative factor a in a way that covers the spread of the distribution 
    shown in Fig. 2.}\\
\newline
We agree with the reviewer. The revised manuscript includes a randomization of the parametrization of the multiplicative factor $a$.  Specifically,
we randomly perturb $N_w$ by multiplying its parameterized value with log-normally distributed
random variable with a standard deviation of 0.5.  As explained above, we use a Gaussian smoothing filter to impose a vertical auto-correlation in
$N_w$ similar to that observed in field experiments.\\
\newline
\textit{3. The authors should include all radiometer channels in their analysis.}\\
\newline
As explained above, we included all the radiometer channels in the revised version of the manuscript.\\
\newline
\textit{4. The introduction's title and first sentence mention "high ice clouds". However, the authors do not provide any 
    information on how the 'high ice clouds' are defined or what distinguishes them from regular ice clouds.}\\
\newline
We dropped the qualifier "high" from the title and the first sentence of the introduction.\\
\newline
\textit{5. Although the authors state that the sensor characteristics are set based on the expected characteristics of the 
    AOS mission instruments, it is unclear whether thermal noise is included in the retrieval experiments. 
    The authors should mention this in the description of their methodology.}\\
\newline
No, we don't include thermal noise in the retrieval experiments.  We acknowledge the lack of an uncertainty component in 
the lidar model in section 2a.\\ 
\newline
\textit{\textbf{Minor comments}
    1. l. 29 - 31: Please provide more information on the planned or possible observation configuration of AOS, such as swath
     widths of the active instruments and the type of radiometer. The provided reference (Braun 2022) is missing, and I 
     could not find conclusive information on this online.}\\
\newline
We apologize for the missing reference.  We included a reference to an IEEE IGARS conference paper \cite{braun2022} that provides a brief description of the AOS mission.
The IEEE document is publicly available and provides a link to an additional AOS document (https://aos.gsfc.nasa.gov/docs/ACCP\_Science\_Narrative-2021.07.19.pdf). 
However, the swath widths of the active instruments are not explicitly specified, as architectures and instruments have been not finalized.  Nevertheless,
the Ku-band radar will be based on the GPM DPR radar and is expected to have a swath width of 245 km.  The elastic lidar will be in many respects 
similar to the CALIPSO/CALIOP lidar (https://calipso.cnes.fr/en/CALIPSO/lidar.htm) and will provide nadir observations,  while the radiometer is expected 
to have a swath of around 1000 km.\\
\newline
\textit{2. l. 83 -90: The results by Testud 2001 are for raindrops assuming Rayleigh scattering. It is not evident that this 
    is also generally true for ice particles.}\\
\newline
The impact of normalization on Z-IWC relationship is shown in \cite{delanoe2014}. Soft spheres were assumed and the Mie solution was used in \cite{delanoe2014}
to derive reflecitivities from observed Particle Size Distributions, but \cite{chase2021}  used more advanced backscattering models to calculate Z from
observed PSDs.  Their calculations are publicly available at https://github.com/dopplerchase/Chase\_et\_al\_2021\_NN and may be used to confirm the contraction
of the ($Z/N_w$,$IWC/N_w$) distribution to almost one-to-one relationships.  As explained in the manuscript, we use the self-similar Rayleigh-Gans approximation
to model the backscattering property of ice particles but get a normalization behavior similar to that anticipated by Testud et al. (2001),
shown in \cite{delanoe2014} and present in the \cite{chase2021} database.\\
\newline
\textit{3. l. 96: Please clarify: What do the authors mean by 'similarity'? Euclidean distance?}\\
\newline
Yes, the similarity is measured by the Euclidean distance.  We added a sentence to clarify this point.\\
\newline    
\textit{4. l. 112 and Fig. 2: Please clarify: Why are there 36 scatter points in Fig. 2.? How is the relative height value 
    of the scatter points derived?}\\
\newline
We initially started with 36 similarity classes but then reduced the number of classes to 18 to make the classification and 
estimation process more robust. However, we kept the results from the 36 classes analysis in the derivation of the $N_w$ parameterization,
as they enable the derivation of a more accurate parameterization than 18 classes. The relative height is the distance
between the peak of the mean IWC peak and the freezing level.\\
\newline
\textit{5. Fig. 2: Please add a line showing the parametrization used to derive the database.}\\
\newline
Thanks for the suggestion.  We added a line showing the parametrization used to derive the database.\\
\newline
\textit{6. l. 130: Please clarify: What is the assumed shape of the ice particles?}
The ice particles are aggregates of bullet rosettes, columnar crystals, and plates.  The aggregation model and 
other details are comprehensively described in \cite{hogan2014}. We added a sentence in the manuscript to clarify this point.\\
\newline
\textit{7. Fig. 3: Please clarify: Why is the distribution of radar reflectivities going to zero towards the freezing level? 
    What about precipitating clouds?}\\
\newline
We selected only CloudSat reflectivity profiles with no echo below the freezing level and the freezing level above ground.  
Profiles that satisfy this requirement include anvil clouds, whose quantification in relation to the convective process with which they are associated
is an AOS objection. We added a sentence in the manuscript to clarify this point.\\
\newline
\textit{8. l. 181 - 182: Please clarify: Do the authors mean surface emissivities?}\\
\newline
Yes, we mean surface emissivities.  We clarified this in the manuscript.\\
\newline
\textit{9. l. 198: Please provide more information on how the IWC and PSD parameters are estimated. If more than one moment
is estimated, the authors must use a priori assumptions to resolve the ambiguity in the retrievals.}\\
\newline
We derive the IWC as a function of CloudSat $Z$ and $N_w$.  The $N_w$ is given by the vertical parameterization and perturbed using the random model
described above and in the revised manuscript.  We added a sentence in the manuscript to clarify this point.\\
\newline
\textit{10. l. 203: I don't think that 200 000 profiles can be considered a large dataset by today's standards.}\\
\newline
We agree that 200,000 profiles is not a large dataset by today's standards.  However, these profiles are derived from a month of CloudSat observations.
Specifically, all CloudSat and 2ICE product files from June 2019 were processed, and only profiles with no echo below the freezing level
and the freezing level above ground were selected. The selection process significantly reduced the number of profiles. Nevertheless, there is sufficient
variability in the selected profiles to derive meaningful conclusions. We added a sentence in the manuscript to clarify this point.\\
\newline
\textit{11. Fig. 4: Please add lines marking the diagonal and conditional mean}\\
\newline
Thank you for the suggestion. We added the lines to the figure.\\
\newline   
\textit{12. Fig. 6: I think it would be valuable to check whether the classification step causes the multi-modal distribution 
    in the retrieval results. Please produce a version of Fig. 6 in which the 18 centers of gravity of the clusters 
    are marked using horizontal lines. The is, however, no need to include the figure in the manuscript if that doesn't lead
     to any conclusive results.}\\
\newline
Thank you for the suggestion.  We added the lines to the figure.\\
\newline  
\textit{\textbf{Typos}
    1. l. 323: Missing word}\\
\newline
Thank you for pointing this out.  We added the missing word.\\
\newline

\begin{thebibliography}{999}

  \bibitem{stephens2008}
  Stephens GL, Vane DG, Tanelli S, Im E, Durden S, Rokey M, Reinke D, Partain P, 
  Mace GG, Austin R, L'Ecuyer T. CloudSat mission: Performance and early science 
  after the first year of operation. 
  Journal of Geophysical Research: Atmospheres. 2008 Apr 27;113(D8).

  \bibitem{sano2022}
  Sanò P, Casella D, Camplani A, D’Adderio LP, Panegrossi G. A Machine Learning
  Snowfall Retrieval Algorithm for ATMS. Remote Sensing. 
  2022 Mar 18;14(6):1467.

  \bibitem{stubenrauch2021}
  Stubenrauch CJ, Caria G, Protopapadaki SE, Hemmer F. 3D radiative heating of 
  tropical upper tropospheric cloud systems derived from synergistic A-Train 
  observations and machine learning. 
  Atmospheric Chemistry and Physics. 2021 Jan 26;21(2):1015-34.

  \bibitem{hartmann2016}
  Hartmann DL. Tropical anvil clouds and climate sensitivity. 
  Proceedings of the National Academy of Sciences. 2016 Aug 9;113(32):8897-9.

  \bibitem{braun2022}
  S. A. Braun, J. Yorks, T. Thorsen, D. Cecil and D. Kirschbaum, 
  "NASA'S Earth System Observatory-Atmosphere Observing System," 
  IGARSS 2022 - 2022 IEEE International Geoscience and Remote Sensing Symposium, 
  Kuala Lumpur, 
  Malaysia, 2022, pp. 7391-7393, doi: 10.1109/IGARSS46834.2022.9884029.

  \bibitem{nixon2019}
  Nixon, M. and Aguado, A., Feature extraction and image processing for computer vision. 2019, Academic press.

  \bibitem{grecu2018}
  Grecu M, Tian L, Heymsfield GM, Tokay A, Olson WS, Heymsfield AJ, Bansemer A. Nonparametric methodology to estimate precipitating ice from multiple-frequency radar reflectivity observations. Journal of Applied Meteorology and Climatology. 2018 Nov;57(11):2605-22.

  \bibitem{delanoe2014}
  Delanoë, J. M. E., Heymsfield, A. J., Protat, A., Bansemer, A., and Hogan, R. J. , Normalized particle size distribution for remote sensing application, J. Geophys. Res. Atmos. 2014, 119, 4204– 4227, doi:10.1002/2013JD020700.

  \bibitem{chase2021}
  Chase RJ, Nesbitt SW, McFarquhar GM. A dual-frequency radar retrieval of two parameters of the snowfall particle size distribution 
  using a neural network. Journal of Applied Meteorology and Climatology. 2021 Mar;60(3):341-59.

  \bibitem{hogan2014}
  Hogan, R. J., and C. D. Westbrook, 2014: Equation for the Microwave Backscatter Cross Section of Aggregate Snowflakes Using the Self-Similar Rayleigh–Gans Approximation. J. Atmos. Sci., 71, 3292–3301, https://doi.org/10.1175/JAS-D-13-0347.1.

\end{thebibliography}

\end{document}

### 2023-05-02 RST/RSTManuscript.tex
revision Reviewer_2

%% Version 6.1, 1 September 2021
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% TemplateV6.1.tex --  LaTeX-based blank template for submissions to the 
% American Meteorological Society
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% PREAMBLE
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% Start with one of the following:
% 1.5-SPACED VERSION FOR SUBMISSION TO THE AMS
\documentclass{ametsocV6.1}
\usepackage{slashbox}
% TWO-COLUMN JOURNAL PAGE LAYOUT---FOR AUTHOR USE ONLY
% \documentclass[twocol]{ametsocV6.1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%% To be entered by author:

%% May use \\ to break lines in title:

\title{Synergistic retrievals of ice in clouds from elastic backscatter lidar, 
Ku-band radar and submillimeter wave radiometer observations}

%% Enter authors' names
% and affiliations as you see in the examples below.
%
%% Use \correspondingauthor{} and \thanks{} (\thanks command to be used for affiliations footnotes, 
%% such as current affiliation, additional affiliation, deceased, co-first authors, etc.)
%% immediately following the appropriate author.
%
%% Note that the \correspondingauthor{} command is NECESSARY.
%% The \thanks{} commands are OPTIONAL.
%
%% Enter affiliations within the \affiliation{} field. Use \aff{#} to indicate the affiliation letter at both the
%% affiliation and at each author's name. Use \\ to insert line breaks to place each affiliation on its own line.

%\authors{Author One,\aff{a}\correspondingauthor{Author One, email@email.com} 
%Author Two,\aff{a} 
%Author Three,\aff{b} 
%Author Four,\aff{a} 
%Author Five\thanks{Author Five's current affiliation: NCAR, Boulder, Colorado},\aff{c} 
%Author Six,\aff{c} 
%Author Seven,\aff{d}
% and Author Eight\aff{a,d}
%}
%
%\affiliation{\aff{a}{First Affiliation}\\
%\aff{b}{Second Affiliation}\\
%\aff{c}{Third Affiliation}\\
%\aff{d}{Fourth Affiliation}
%}

\authors{Mircea Grecu\aff{a,b}\correspondingauthor{Mircea Grecu, mircea.grecu-1@nasa.gov}
and John E. Yorks \aff{a}}
\affiliation{\aff{a}{NASA GSFC}\\
\aff{b}{Morgan State University}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ABSTRACT
%
% Enter your abstract here
% Abstracts should not exceed 250 words in length!
%

\abstract{ In this study, we investigate the synergy of elastic backscatter lidar, Ku-band radar, and sub-millimeter-wave radiometer measurements
in the retrieval of ice from satellite observations. The synergy is analyzed through the generation of a large dataset 
of Ice Water Content (IWC) profiles and simulated lidar, radar and radiometer observations. 
The characteristics of the instruments e.g. frequencies, sensitivities, etc. are set based on the expected characteristics 
of instruments of the Atmosphere Observing System (AOS) mission. A cross-validation methodology is used to assess the
accuracy of the IWC profiles retrieved from various combinations of observations from the three instruments. Specifically, the IWC and
associated observations are randomly divided into two datasets, one for training and the other for evaluation.
The training dataset is used to train the retrieval algorithm, while the evaluation dataset is used to assess the 
retrieval performance. The dataset of IWC profiles is derived from CloudSat 
reflectivity and CALIOP lidar observations.
The retrieval of the ice water content IWC profiles from the computed observations is achieved in two steps. 
In the first step, a class, out of 18 potential classes characterized by different vertical distribution of IWC, 
is estimated from the observations. The 18 classes are predetermined based on the k-Means clustering algorithm. 
In the second step, the IWC profile is estimated using an Ensemble Kalman Smoother (EKS) algorithm that uses the 
estimated class as a priori information.\\
The results of the study show that the synergy of lidar, radar, and radiometer observations is significant
in the retrieval of the IWC profiles.  The inclusion of the lidar backscatter observations in the retrieval process has a
larger impact on the retrieval performance than the inclusion of the radar observations. 
As ice clouds have a significant impact on atmospheric radiative processes, this
work is relevant to ongoing efforts to reduce uncertainties in climate analyses and projections.}

\begin{document}

%% Necessary!
\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% SIGNIFICANCE STATEMENT/CAPSULE SUMMARY
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% If you are including an optional significance statement for a journal article or a required capsule summary for BAMS 
% (see www.ametsoc.org/ams/index.cfm/publications/authors/journal-and-bams-authors/formatting-and-manuscript-components for details), 
% please apply the necessary command as shown below:
%
% Significance Statement (all journals except BAMS)
%

%\statement
%This paper is significant.
%	 Enter significance statement here, no more than 120 words. See \url{www.ametsoc.org/index.cfm/ams/publications/author-information/significance-statements/} for details.
%
%% Capsule (BAMS only)
%%
%\capsule
%       Enter BAMS capsule here, no more than 30 words. See \url{www.ametsoc.org/index.cfm/ams/publications/author-information/formatting-and-manuscript-components/#capsule} for details.
%
%% * * If using twocol mode, you will need to use the commands "twocolsig" and "twocolcapsule" in place of "sig" and "capsule"
%%      to ensure that the text box correctly spans across both columns.
%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% MAIN BODY OF PAPER
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%

%% In all cases, if there is only one entry of this type within
%% the higher level heading, use the star form: 
%%
\section{Introduction}
The future NASA Atmospheric Observing System (AOS) mission (Braun 2022) is expected to feature new 
combinations of observations that may be used to quantify the amounts of ice in high clouds and 
characterize the microphysical properties of ice particles. These observations include backscatter
from an elastic backscatter lidar \citep{weitkamp2006},
Ku-band radar reflectivity, and submillimeter wave radiometer brightness temperature measurements. 
The AOS mission design, while not necessarily optimal for cloud ice estimation, enables characterization 
of the broader spectrum of cloud and precipitation processes. Nevertheless, these measurements are 
complimentary for synergistic characterization of ice clouds. 
%While not optimal for cloud ice estimation, but for the characterization of a broader spectrum of 
%cloud and precipitation processes, these observations are nevertheless synergistic from the 
%characterization of ice clouds perspective. 
That is, despite the fact that lidar observations 
attenuate quickly in thick ice clouds and the Ku-band radar will not be able to detect clouds 
characterized by an echo weaker than 8.0 dBZ, the active observations are expected to provide context 
that may be incorporated into the radiometer retrievals. Herein, the term retrieval is defined as the 
process of estimating geophysical variables from remote sensing observations.
%, but, as customary in the scientific
%literature, estimates are occasionally referred to as retrievals as well.  
In this study, we investigate the impact of 
incorporating lidar and radar observations into the radiometer retrieval of ice clouds.
Because the existing amount of coincident backscatter lidar, Ku-band radar, and submillimeter-wave 
radiometer observations is rather insufficient to derive conclusive results, we employ accurate 
physical models to simulate lidar, radar and radiometer observations and use a cross-validation 
methodology to characterize the retrieval accuracy. As estimates from passive instrument 
observations strongly depend on "a priori" information \citep{rodgers2000inverse}, for results to be 
relevant in real applications, it is necessary to base them on realistic vertical distributions of 
ice properties.  Such distributions may be derived from cloud-resolving-model (CRM) simulations 
\citep{pfreundschuh2020synergistic,liu2022assessing} or directly from observations.  In this study, we employ the latter 
approach, as CRMs may still be deficient in properly reproducing the vertical distribution of ice 
clouds and their associated microphysical properties.  Specifically, we use observations and
products from the CloudSat (CS) mission (Stephens et al. 2002) to derive a database of ice microphysical
properties and associated simulated lidar, radar, and radiometer observations.  This database is used to
investigate the accuracy of estimated ice cloud properties from the simulated observations. Another major
difference relative to previous studies is the unique combination of instruments investigated herein.
The article is organized as follows.  In Section 2, we describe the approach used to derive the ice properties and
the associated simulated observations, the retrieval and the evaluation methodology.  In Section 3, we 
present the results of the evaluation methodology. We conclude in Section 4.

% \subsection*{subsection}
% text...
\section{Methodology}
As previously mentioned, we use CloudSat (CS) observations \citep{stephens2002} to derive the vertical 
distributions of ice properties needed in the investigation.  Although research quality CS cloud ice 
products exist, to maximize the physical consistency of the approach, we do not use them but derive ice amounts
and associated properties directly from CS reflectivity observations.  This ensures the consistency between 
the particle
distribution assumptions and the electromagnetic scattering properties used in the CS reflectivity processing 
and those used the simulation of the lidar, Ku-band radar and radiometer observations.  
Our CS-based ice product is optimized to be consistent with the synergistic 
Cloudsat and CALIPSO Ice Cloud Property Product (2C-ICE) of \cite{deng2015}.  When the CALIPSO lidar detects echo
associated with clouds but the CS radar signal is below the noise level, we use the 2C-ICE product to extend our 
CS-based estimates. Lidar, Ku-band radar, and submillimeter-wave radiometer observations are simulated 
from CS observations using accurate physical 
models and realistic assumptions consistent with the most recent knowledge in the field of ice cloud 
microphysics, and a non-parametric estimation methodology based on the k-Means clustering algorithm 
\cite{mackay2003information} is used to 
investigate the instrument synergy.  Details of the methodology are presented below.

\subsection{Assumptions and forward models}
To quantify the number of ice particles in an elementary atmospheric volume as a function of their size, we use
normalized gamma functions \citep{bringi2003}.  The benefit of normalized gamma functions is that they 
encapsulate the variability of Ice Water Content (IWC) - reflectivity relationship into a single parameter,
i.e. the 
normalized Particle Size Distribution (PSD) intercept \citep{testud2001,bringi2003}. The normalized
PSD intercept is defined as $N_w=\frac {4^4} {\pi \rho_w} \frac {IWC} {D_m^4}$, where $IWC$ is the ice water 
content associated with the PSD, and $D_m$ is the mass weighted mean diameter.  \cite{testud2001} showed 
that the variability in IWC reflectivity (Z) relationships may be fully explained by variability in $N_w$, and 
that a formula of the type
\begin{equation}
IWC=N_w^{1-b}aZ^b \label{eq:1}
\end{equation}
perfectly explains the relationships between IWC and Z calculated from observed PSDs. Equation (\ref{eq:1}) is not 
sufficient to derive accurate, unbiased estimates of ice water contents, because $N_w$ varies considerably in 
time and space. Nevertheless, multiple studies showed that it is beneficial to parameterize $N_w$ as a function 
of various variables, such as temperature \citep{hogan2006retrieval, delanoe2008variational,deng2010tropical}, 
rather than using $N_w$ independent reflectivity ice relations.  In this study, we parameterize
$N_w$ as a function of temperature based on the CloudSat 2C-ICE product \citep{deng2010tropical,deng2013evaluation}.  
Specifically, we cluster, based on similarity, a large set 2C-ICE profiles into 18 classes using a k-Means 
procedure. The mean IWC profiles associated with the 18 classes are shown in continuous lines in Figure 1.  
Our estimates, derived using PSD assumptions and electromagnetic scattering calculations that enable 
accurate and physically consistent simulations of radar observations at Ku-band and radiometer observations of 
submillimeter-wave frequencies are also shown in Figure 1. These estimates are based on the self-similarity 
Rayleigh-Gans approximation (SSRGA) of \cite{hogan2017ssrga}. Details regarding the
estimation process are provided in the following paragraphs.  As apparent in Figure 1, the CS and SSRGA estimates are in good 
agreement.  Some discrepancies due to differences between the SSRGA $N_w$ parameterization and the CS 2C-ICE 
"a priori assumptions" are also apparent, but they are not deemed critical in this study, whose objective is 
the investigation of synergistic lidar, Ku-band radar and submillimeter-wave radiometer retrievals, because the 
outcome is not likely to be sensitive to such details.
\begin{figure}[t]
    \centering
    \includegraphics[width=0.75\textwidth,angle=0]{./Figs/fig01.png}\\
    \caption{Mean CS IWC profiles for 18 classes derived using the k-Means clustering algorithm. For a 
    compressed but intelligible representation, three classes are shown in different colors in each panel.
    The associated 
    mean profiles derived from CS reflectivity observations using the SSRGA scattering calculations 
    and $N_w$ parameterization developed in this study are shown using symbol *. The vertical coordinate 
    is defined relative to the freezing level.}\label{f1}
\end{figure}
One may notice that the average IWC profiles in Figure 1 are characterized by different peak values and heights. 
This facilitates a simple way to approximately reverse-engineer the "a priori" assumptions used in the CS 
2C-ICE product and use them in formulation of the type described in Equation (1).  Specifically, the  
derivation of relationships of the type $IWC=a_i Z^{b_i}$ for every class i may be used to study $a_i$ as a
function of height. 
Shown in Figure 2 is a  scatter plot representation of the class multiplicative coefficient $a_i$ as a function of
relative height (defined as the absolute height relative to the height of the freezing
level).  As 
apparent in the figure, and as expected, $a_i$ exhibits a strong variation with the relative height. 
Coefficient $b_i$ exhibits a height dependency as well (not shown), but the range of variation is significantly
smaller, 
almost zero relative to the mean value of $b$. Given that any deviation of the multiplicative coefficient in 
an IWC-Z relation from an average is equivalent to a deviation of the associated $N_w$ from its mean value 
(Testud et al. 2001), the variation of $a$ as a function of relative-height may be converted into a $N_w$ 
as a function of relative-height relationship.  We, therefore, use the data in Fig. 2 to parameterize $N_w$ as a 
function of the relative height.

\begin{figure}[t]
    \centering
    \includegraphics[width=0.75\textwidth,angle=0]{./Figs/fig02.png}\\
    \caption{Multiplicative factor in the 2ICE IWC-relation as a function of the IWC profile
    peak height.}\label{f2}
\end{figure}

For the determination of reference $a$ and $b$ values to be used with Equation (\ref{eq:1}), we assume that 
PSDs are normalized gamma  distributions with $N_w=0.08cm^{-4}$ and $\mu=2$ and calculate 
\begin{equation}
 Z=\frac {\lambda ^4} {\pi ^5 |K_w|^2} \int_0^{\infty} N(D,D_m) \sigma _b(D) dD
\end{equation}
where $\lambda$ is the radar frequency, $|K_w|$ is the dielectric factor of water, $N(D,D_m) dD$ is the number 
of ice particles of diameter with D and D+dD per unit volume, $D_m$ is the mass weighted mean diameter of the 
distribution, and $\sigma _b(D)$ is the backscattering cross-section of ice particle of diameter $D$. The mass 
weighted mean diameter is equidistantly sampled to span the entire range of IWC values in the CS 2C-ICE dataset.
The assumed mass-size relation is that of \cite{brown1995improved} because it works well with the 
SSRGA scattering calculations \citep{heymsfield2022impacts}.  The open source software scatter-1.1 of \citep{scatter-1.1}
is used to provide the actual scattering properties. The SSRGA theory was developed for millimeter and 
submillimeter-wave calculations and may not be applicable at lidar's wavelength.

The lidar observations are computed using the Mie solution included in the scatter-1.1 
package for a backscatter lidar, given that such a lidar is expected to be flown with a 
Ku-band radar and microwave radiometer as part of the NASA Earth System Observatory 
(ESO) Atmosphere Observing System (AOS) in an inclined and/or polar orbit \citep{yorks2022}.
The lidar simulations are idealized in that they do not include random errors due to daytime 
solar background noise or systematic errors such as calibration accuracy or lidar ratio
assumptions. There are three main reasons this approach was taken: (1) it is important to
understand the limitations of combining these three datasets before factoring in 
individual sensor limitations, (2) while more accurate calculations based on more 
realistic ice particle shapes exist, they are rather incomplete and not readily available,
and (3) Wagner and Delene (2022) compared lidar backscatter observations with backscatter 
calculations based on coincident PSD observations and the Mie solution and found good 
agreement, which suggests that electromagnetic properties derived from Mie calculations 
are adequate for practical applications. The lidar molecular backscatter and extinction
are calculated using the lidar module of the CFMIP Observation Simulator Package (COSP; 
Bodas-Salcedo et al. (2011)). COSP simulates the lidar total attenuated backscatter 
signal and scattering ratios at 532 nm for scenes with and without clouds. COSP also 
assumes cloud particles are spherical so that the backscattering phase function is 
estimated based on the effective radius (Mie theory). Despite this assumption, the 
COSP simulations agree well with CALIOP observations. To account for multiple-scattering 
in the lidar observations, we are using the multiscatter-1.2.11 model (Hogan 2015) of 
Hogan and Battaglia (2008).

Shown in Figure 3 are the distributions of simulated Ku-band radar reflectivity and 
lidar backscatter as function of height above the freezing level.
As apparent in the figure, the IWCs associated with detectable Ku-band reflectivity
signal are likely to
occur near mostly around 3.0 km above the freezing level, while the lidar backscatter
distribution exhibits a peak at around 6.0-7.0 km
above the freezing level, which is consistent with the fact the lidar observations
are strongly attenuated in the bottom part of the cloud.

\begin{figure}[t]
    \centering
    \includegraphics[width=0.75\textwidth,angle=0]{./Figs/fig03.png}\\
    \caption{Simulated distributions of Ku-band radar reflectivity (left) and lidar backscatter (right)
    as function of height above 
    the freezing level.}\label{f3}
\end{figure}

The radiometer observations are calculated using a one-dimensional efficient, but accurate, radiative transfer 
solver based on Eddington's approximation \citep{kummerow1993edd}. The Eddington's approximation has been found to work 
well in cloud and precipitation retrieval application despite its simplicity relative to more general (but also 
computationally intensive) approaches such as Monte Carlo radiative transfer solvers \citep{liu1996three} 
It should be noted tough that the phase functions of ice particles tend to be highly asymmetric at 
sub-millimeter wave frequencies. For radiative transfer solutions based on the Eddington's approximation to be 
accurate, it is necessary that the delta-scaling approach \citep{joseph1976delta} be employed. The delta-scaling 
approach transforms the initial radiative transfer equation into an equivalent one characterized by a less 
asymmetric scattering function and more extinction, which makes the solution Eddington approximation more stable 
and accurate. The absorption due to water vapor and other gases is quantified using the Rosenkranz model 
\citep{rosenkranz1998}.  The water vapor, temperature, and pressure distributions are derived based on a WRF 
simulation of summer convection over the United States.  Specifically, the water vapor, temperature, and 
pressure profiles associated with times and areas where the model produces anvils are selected and clustered 
into 40 classes using the k-Means approach.  The mean extinction profiles at the radiometer frequencies are 
calculated for every class 
and used in process of calculating the brightness temperatures from the estimated ice profiles using a simple 
Monte Carlo procedure.  That is, given a retrieved ice profile and its scattering property, an anvil class 
and its associated absorption, temperature, and pressure profiles are randomly selected and attached to the 
ice scattering properties. To make the procedure physically meaningful, temperature rather than height is 
used in the ice scattering-gas absorption collocation process.  The emissivities are randomly chosen between 
0.8 and 1.0 and assumed constant for all radiometer frequencies.  Brightness temperatures are calculated at 
89-, 183.31 ± 1.1, and 325.15 ± 1.5 GHz, which correspond to three of the 10 channels of the SAPHIR-NG 
radiometer envisioned to be deployed in the AOS mission \citep{brogniez2022time}.  The other channels are 
centered on the same water vapor absorption lines and are not likely to offer additional information in 
this rather controlled experiment. Nevertheless, the other channels are expected to be useful in reducing the 
uncertainties caused by variability in the vertical distribution of water vapor, which may be greater in 
real life than in the simulated environment.

The processing steps used to process the CS reflectivity observations and calculate the lidar, Ku-band and 
submillimeter-wave radiometer observations may be summarized as follows:
\begin{enumerate}
\item Derivation of physically consistent radar and radiometer lookup tables to relate basic radar and 
radiometer properties (e.g. reflectivity, attenuation, extinction, scattering-albedo, etc.) to PSD parameters 
such as $IWC$ and $D_m$. The tables are derived for a single of $N_w$, but are usable with any value of 
$N_w$ using the "normalization" operations described in \citep{grecu2011}.
\item Derivation of $N_w$-relative height parameterization using the 2C-ICE product.
\item Estimation of IWC and related PSD parameters from CS W-band radar observations, using the tables 
constructed
in Step 1 and parameterization derived in Step 2.
\item Calculation of lidar, Ku-band radar and radiometer observations from the estimates derived in Step 3 
and the tables obtained in Step 1, and extended with non-zero 2C-ICE estimates for the radar bins characterized by no echo
above the noise level.
\end{enumerate}

The application of these steps produces a large dataset of approximately 200,000 cloud ice profiles and 
associated lidar, radar and radiometer observations that may be used to investigate the synergy of the three 
sensors. Details are provided in the next section.

\subsection{Estimation and evaluation}
Given that the lidar observations may attenuate quickly in thick clouds, while the Ku-band radar will not detect
clouds with an echo weaker than 8.0 dBZ, the radiometer is the instrument likely to provide by itself the most 
complete information about the total amount of ice in its observing volume.  However, the vertical distribution 
of ice is difficult to quantify from radiometer-only observations, because significantly different ice vertical 
distributions may lead to very similar radiometer observations.  This makes radiometer-only retrievals highly 
dependent on the "a priori" information on the distribution of ice clouds in the atmosphere. As previously 
mentioned, this is the reason why CS-based IWC retrievals were preferred to CRM simulations, as retrievals are 
expected to result in more natural and less biased distributions.

For retrievals, we employ a two-step estimation methodology similar to that of \cite{grecu2018}.  In the first step, we 
estimate an IWC class, out of the 18 classes of shown in Figure \ref{f1},
to which the estimated IWC profile is most likely to belong. The class estimation procedures is 
trained using the synthetic observations.
In the second step, we estimate the IWC profile,
using a class specific ensemble Kalman Smoother (EKS) methodology similar to that of \cite{grecu2018}.  
The EKS algorithm updates the estimated IWC relative to the mean IWC of the class to which the profile belongs.
The differences between the actual active and passive observations and their mean class values are used
in the update. The 
second step of this procedure is formally identical to the one used in \cite{grecu2018}, but the first step is
different. In \cite{grecu2018}, the first step was based on a simple distance-based evaluation. That strategy
is likely to be suboptimal in this study, because the joint distribution of IWC profiles and associated
observations are significantly more complex. We, therefore, use a more complex classification methodology
based on the TensorFlow library \citep{tensorflow2016}. The class estimation model is defined as a
TensorFlow Model with two dense layers of 30 neurons each, followed by a softmax layer \citep{deepL2016}. The 
class estimation model is trained using the 70\% of the simulated observations and the corresponding IWC 
profiles, while the remaining 30\% of the data being used for evaluation. 
The EKS update is based on the formula%The training is done using the Adam
\begin{equation}
    \mathbf{X}=\mathbf{\bar{X}_i}+\mathbf{Cov(X_i,Y_i)}\mathbf{Cov(Y_i,Y_i)}^{-1}(\mathbf{Y}-\mathbf{\bar{Y}_i}) \label{eq:enks}
\end{equation}
where $\mathbf{X}$ is the state variable describing the IWC profile, $\mathbf{Y}$ is the vector containing
the actual observations, $\mathbf{X_i}$ is the set of state variables for profiles
in class $i$, and $\mathbf{Y_i}$ is the
set of observations associated with profiles in class $i$.  Variables $\mathbf{\bar{X}_i}$ and $\mathbf{\bar{Y}_i}$ are the mean values
of the state variables and observations in class $i$, respectively. The covariance matrices between 
$\mathbf{X_i}$ and $\mathbf{Y_i}$ are denoted by $\mathbf{Cov(X_i,Y_i)}$. In step 1, the class is estimated
using the TensorFlow model, while in step 2, the IWC profile is estimated using the EKS algorithm summarized in 
Equation \ref{eq:enks}.

As already mentioned, a cross-validation methodology is used for evaluation, with 70\% of 
the data used for training and the remaining 30\% of the data used for validation.  The partition of the data into 
training and 
evaluation subsets is done randomly.  Usually, the partition, training and evaluation steps are repeated 
several times.  However, given the fact that differences in the relationships between the ice property and 
their associated simulated observations are functions of the meteorological context, and that all regimes are 
well-sampled in both the training and testing subsets (e.g. out of every 10 pixels in a scene, about seven end up 
in the training dataset, while the others in the testing dataset), the repetition of the partition, training, 
and evaluation steps multiple times is not necessary. Therefore, in our evaluation, we partition the data into 
training and evaluation only once and perform all the evaluation for a single partition. The evaluation 
criteria include the correlation coefficient, the bias, and visual inspections of graphical representations of 
the estimated properties relative to their references.
%vs

\section{Results}
\subsection{Radiometer-only retrievals}
As previously mentioned, submillimeter-wave radiometers are likely to provide by themselves more complete 
information about the total amount of ice in their observing volumes than lidars or Ku-band radars with 
limited sensitivity. However, radiometers observations are an integrated measure of radiative processes in 
clouds that provide little information about the vertical distribution of ice. From this perspective, an 
evaluation in terms of the ice water path (IWP) defined as the vertical integral of the IWC, i.e. 
$IWP=\int_0^{Z_{top}}IWC(z)dz$ is insightful.  Shown in Figure \ref{f4} is the distribution of IWP 
estimated from radiometer-only observations as a function of its true value. As apparent in the figure, there is good 
correlation between the retrieved and the true IWP values. The numerical value of the correlation coefficient 
is 0.92, and there is no-overall bias. That is, the mean values of retrieved IWP and true IWP values are 
equal. However, conditional biases are apparent, with overestimation of IWP for values smaller than 
100 g/$m^2$ and some underestimation for values larger than 1000 g/$m^2$.  The biases at the low end of the 
IWP range are not surprising, given that the impact caused by ice scattering on the total radiometric signal 
is small for low values of IWP and hard to distinguish from other sources of variability in radiometer 
observations. Saturation effects are most likely responsible for underestimation at the high end.  It should 
be noted that in this evaluation, only atmospheric profiles that exhibit ice detectable by the CS radar or CALIOP lidar are used. 
Therefore,
a radiometer-only estimation procedure derived from this training dataset is likely to result in significant 
overestimation if not used in conjunction with a discrimination procedure. However, such procedure is not critical 
in this study, as, in a synergistic application, the lidar observations may be used to discriminate between clear skies and ice clouds.
\begin{figure}[t]
    \centering
    \includegraphics[width=0.75\textwidth,angle=0]{./Figs/fig04.png}\\
    \caption{Frequency plot of estimated IWP derived radiometer-observations as a function of the 
    true IWP used in observations synthesis.}\label{f4}
\end{figure}
However, although the radiometer-only estimation procedure is able to estimate the integrated amount of ice in clouds fairly
wells, its ability to characterize the vertical distribution of ice in clouds is limited.  Figure \ref{f5} shows the
conditional vertical distributions of the estimated and true IWC for the 18 classes described in Section 2a 
and shown in Figure \ref{f1}. As apparent in the figure, there are significant differences between the
estimated and true IWC profiles. 
\begin{figure}[t]
    \centering
    \includegraphics[width=0.75\textwidth,angle=0]{./Figs/fig05.png}\\
    \caption{True and radiometer-only retrieved conditional mean IWC for the 18 classes described in 
    Figure \ref{f1}.}\label{f5}
\end{figure}

Further insight into the radiometer-only estimation performance may be derived by defining the ice profile gravity center 
(GC) as $z_{GC}=\frac {\int_0^{Z_{top}}zIWC(z)dz} {\int_0^{Z_{top}}IWC(z)dz}$, where $z$ is the distance relative to 
the freezing level, the $Z_{top}$ is the distance from the top of the atmosphere to the freezing level.  
Shown in Figure \ref{f6} is the frequency of IWC gravity center
estimated from radiometer-only observations as a function of its true value.  It may be observed in the figure that while 
the true IWC gravity center exhibits quite a broad distribution, the one retrieved from the radiometer-only 
observations exhibits a multimodal narrow distribution. Moreover, there is almost no correlation between the retrieved 
and the true IWC gravity center.  This is another indication that, while the total amount of ice may be reasonably 
estimated from radiometer-only observations, its vertical distribution can not be 
accurately determined from 
radiometer-only observations.

\begin{figure}[t]
    \centering
    \includegraphics[width=0.75\textwidth,angle=0]{./Figs/fig06.png}\\
    \caption{Same as in Figure 4, but for the IWC gravity center.}\label{f6}
\end{figure}

\subsection{Synergistic retrievals}

The synergy of the instrument on the estimates may be investigated by simply incorporating lidar 
and radar observations
into the retrieval process and comparing the results with the radiometer-only estimates.  Although
the lidar observations are subject to attenuation, they are able to provide information about the vertical
distribution of ice in clouds, mostly at the top of the clouds. The radar observations, on the other hand,
are able to provide information in the bottom part of the clouds, where the lidar signal is below the noise
level due to attenuation. Therefore, the combined used of lidar and radar observations is expected to 
provide a more complete characterization of the vertical distribution of ice in clouds and enable the
derivation of more specific estimates than those derived from radiometer-only observations. It should be mentioned that, 
although deficiencies 
and potential biases in the simulated observations may distort conclusions to some degree, the forward models
used in this study are state-of-the-art and are expected to enable a realistic characterization of the impact
of individual instruments on the synergistic retrievals.

Shown in Figure \ref{f7} is the distribution of the synergistic IWP estimates as a function of their true
values.  As apparent in the figure, the synergistic IWP estimates are more accurate than the radiometer-only
estimates. At the same time, as apparent in Figure \ref{f8}, the retrieved conditional mean IWC for the 
18 classes described section 2a and shown in Figure \ref{f1} are in significantly better (almost perfect)
agreement with the true IWC profiles than those derived from radiometer-only observations.  Furthermore,
as seen in Figure \ref{f8} the synergistic IWC gravity center estimates are in very good agreement with 
the true IWC gravity center.

\begin{figure}[t]
    \centering
    \includegraphics[width=0.75\textwidth,angle=0]{./Figs/fig07.png}\\
    \caption{Same as in Figure \ref{f4}, but with the lidar and radar observations incorporated in the
    retrievals.}\label{f7}
\end{figure}

\begin{figure}[t]
    \centering
    \includegraphics[width=0.75\textwidth,angle=0]{./Figs/fig08.png}\\
    \caption{Same as in Figure \ref{f5}, but with the lidar and radar observations incorporated in the
    retrievals.}\label{f8}
\end{figure}

While the estimates based on all instruments are significantly more accurate than those based on radiometer-only
observations, it is useful to investigate how the two active instruments (lidar and radar) impact the
estimates.  For conciseness, we use two statistical scores, namely, the normalized root mean square (NRMS) and
the classification accuracy, to summarize the performance of the estimates.  The NRMS is defined as
\begin{equation}
NRMS=\frac {\sqrt {\frac {\sum_{i=1}^{N} (IWC_{i}-IWC_{true,i})^2} {N}}} {\sqrt {\frac {\sum_{i=1}^{N} (IWC_{true,i}-\overline{IWC})^2} {N}}} 
\end{equation}
where IWC$_{i}$ is the estimated IWP for the $i$-th sample, IWC$_{true,i}$ is the true IWC for the $i$-th sample,
$\overline{IWC}$ is the IWC mean, and $N$ is the size of the estimation dataset.  The classification
accuracy is defined as
\begin{equation}
CA=\frac {\sum_{i=1}^{N} \delta_{i}} {N}
\end{equation}
where $\delta_{i}$ is a binary variable that is equal to 1 if the estimated IWC class for the $i$-th sample is
equal to the true IWC class for the $i$-th sample, and 0 otherwise. The performance summary is shown 
in Table \ref{t1} for several combinations of instruments. It may be observed in the table that the
performance of the estimates based on all instruments is significantly better than those based on
radiometer-only observations.  Furthermore, the inclusion of the lidar observations in the retrieval
process has a larger impact on the retrieval performance than the inclusion of the radar
observations.  This is expected since the lidar observations are able to provide information about the
top of the clouds, where the radar observations are above the noise level only occasionally. Nevertheless,
the inclusion of the radar observations in the retrieval process has a notable impact on the
accuracy of the IWC estimates relative to radiometer-only retrievals.

%Therefore, results in table
%\ref{t1} are in agreement with a qualitative analysis.  Specifically, given the 
%limited sensitivity of the Ku-band radar, the fraction of profiles with radar 
%observations above the noise level is relatively small.  On the other hand, 
%the lidar provides useful information that may be used to more accurately 
%identify the most likely vertical IWC distribution class for all profiles.

\begin{figure}[t]
    \centering
    \includegraphics[width=0.75\textwidth,angle=0]{./Figs/fig09.png}\\
    \caption{Same as in Figure \ref{f6}, but with the lidar and radar observations incorporated in the
    retrievals.}\label{f9}
\end{figure}

\begin{table}[t]
\caption{Performance summary.}\label{t1}
\begin{center}
\begin{tabular}{c|cccc}
\hline\hline
\backslashbox{Score}{Instruments} & Radar-Radiometer & Lidar-Radiometer & Radar-Lidar-Radiometer\\
Score& & & &\\
\hline
% NRMS & 0.73 &  0.59  &  0.32 & 0.32 \\
NRMS & 0.73 &  0.59  &  0.32 & 0.22 \\
Class. Accurracy & 0.39 & 0.48 & 0.92 & 0.94 \\
\hline
\end{tabular}
\end{center}
\end{table}

% text...
% 
\section{Conclusions}
In this study, we investigate the synergy of elastic backscatter lidar, Ku-band radar, and sub-millimeter-wave radiometer
measurements in the retrieval of the ice from satellite observations.  The synergy is analyzed through
the generation of a large dataset of IWC profile and the calculation of lidar, radar and radiometer 
observations using realistic models. The characteristics of the instruments (e.g. frequencies, sensitivities,
etc.) are set based on the expected characteristics of instruments of the AOS mission. 
A cross-validation methodology is used to assess the 
accuracy of the retrieved IWC profiles from various combinations of observations from the three instruments.
Specifically, the IWC and associated observations is randomly divided into two datasets, one for
training and the other for evaluation.  The training dataset is used to train the retrieval algorithm,
while the evaluation dataset is used to assess the retrieval performance.

To ensure the self-consistency of results and their relevance to practical applications, the dataset of IWC 
profiles is derived from CloudSat reflectivity observations and extended with lidar-based estimates from the
2C-ICE product. Although subject to potential biases
and uncertainties due to deficiencies in the retrieval models, these profiles are deemed to be more realistic
than those derived from cloud resolving model simulations. Moreover, they are roughly consistent with
the 2C-ICE CloudSat product \citep{deng2015}, while relying on assumptions and parameterizations that
enable the accurate computation of backscatter lidar, Ku-band radar, and sub-millimiter-wave radiometer
observations.

The retrieval of the ice water content (IWC) profiles from the computed observations is achieved in two
steps.  In the first step, a class, out of 18 potential classes characterized by different 
vertical distribution of IWC, is estimated from the observations. The 18 classes are predetermined based on
k-Means clustering algorithm.  In the second step, the IWC profile is estimated using an Ensemble Kalman
Smoother (EKS) algorithm that uses the estimated class as a priori information.

The results of the study show that the synergy of lidar, radar, and radiometer observations is significant
in the retrieval of the IWC profiles.  The inclusion of the lidar observations in the retrieval process has a
larger impact on the retrieval performance than the inclusion of the radar observations.  Although results are
not directly comparable to those from other studies \citep{pfreundschuh2020synergistic,liu2022assessing}, given the differences
between the instruments in this study relative to those from other studies, it may be concluded that they
are not inconsistent with previous studies. Specifically, previous studies show some skills in radiometer-only
retrievals and significant improvements in the retrieval performance when the active observations are incorporated.
From this perspective, our findings are consistent with previous studies.

Further work is necessary out to assess the impact of sources of uncertainties such as potential biases in the
forward models, variability in the
PSD intercept not captured by the current parameterization, differences in the instruments' footprint sizes,
and non-uniform beam filling on the retrievals of the IWC profiles.  Other sources of uncertainties that
need be considered include the potential existence of supercooled liquid water in the clouds and uncertainties
in the electromagnetic scattering properties used in the in instruments' forward models. These uncertainties
may be best investigated and mitigated through the use of high-quality observations from field campaigns
such as The Investigation of Microphysics and Precipitation for Atlantic Coast-Threatening Snowstorms (IMPACTS)
\citep{mcmurdie2022}.

%%%
% \section{First primary heading}

% \subsection{First secondary heading}

% \subsubsection{First tertiary heading}

% \paragraph{First quaternary heading}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% TABLES---INSERT NEAR IN-TEXT DISCUSSION
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%  Enter tables near where they are discussed within the document. 
%%  Please place tables before/after paragraphs, not within a paragraph.
%%
%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% FIGURES---INSERT NEAR IN-TEXT DISCUSSION
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%  Enter figures near where they are discussed within the document.
%%  Please place figures before/after paragraphs, not within a paragraph.
% %
%
%\begin{figure}[t]
%  \noindent\includegraphics[width=19pc,angle=0]{figure01.pdf}\\
%  \caption{Enter the caption for your figure here.  Repeat as
%  necessary for each of your figures. Figure from \protect\cite{Knutti2008}.}\label{f1}
%\end{figure}

\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ACKNOWLEDGMENTS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\acknowledgments
This work was supported by the NASA Remote Sensing Theory project through Grant 80NSSC20K1729. 
The authors thank Dr. Lucia Tsaoussi (NASA Headquarters) for her support of this effort.
%  Keep acknowledgments (note correct spelling: no ``e'' between the ``g'' and
% ``m'') as brief as possible. In general, acknowledge only direct help in
%  writing or research. Financial support (e.g., grant numbers) for the work done, 
%  for an author, or for the laboratory where the work was performed must be 
%  acknowledged here rather than as footnotes to the title or to an author's name.
%  Contribution numbers (if the work has been published by the author's institution 
%  or organization) should be placed in the acknowledgments rather than as 
%  footnotes to the title or to an author's name.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% DATA AVAILABILITY STATEMENT
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% 
%
\datastatement
The CloudSat data can be accessed at: https://www.cloudsat.cira.colostate.edu/.
%  The data availability statement is where authors should describe how the data underlying 
%  the findings within the article can be accessed and reused. Authors should attempt to 
%  provide unrestricted access to all data and materials underlying reported findings. 
%  If data access is restricted, authors must mention this in the statement. See
%  {http://www.ametsoc.org/PubsDataPolicy} for more info.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% APPENDIXES
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%% If only one appendix, use

%\appendix

%% If more than one appendix, use \appendix[<letter>], e.g.,

%\appendix[A]

%% Appendix title is necessary! For appendix title:

%\appendixtitle{Title of Appendix}

%%% Appendix section numbering (note, skip \section and begin with \subsection)
%
% \subsection{First primary heading}

% \subsubsection{First secondary heading}

% \paragraph{First tertiary heading}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% REFERENCES
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Make your BibTeX bibliography by using these commands:
\bibliographystyle{ametsocV6}
\bibliography{references}

\end{document}

### 2023-05-04 RST/responseReviewer2.tex
Revision

@@
 \newline
- We appreciate \\
+ We thank the reviewer for their perspective and constructive suggestions. Our point by point response is 
+ provided below.\\
 \newline
@@
 manuscript, and more details are needed about the specifics of how IMPACTS
-will help mitigate them.}
-
-2) You say in your abstract (and conclusion) that “the characteristics of 
+will help mitigate them.}\\
+\newline
+We thank the reviewer for the uncertainty discussion. Yes, indeed the resolution of the simulated 
+observations is that of the CloudSat, while the AOS instruments are expected to have different resolutions,
+which might impact results. In the revised manuscript, we included statements to convey this aspect.
+We agree that the assumed electromagnetic scattering properties assumed in the study although
+reasonable may contain significant uncertainties.  We acknowledge this aspect in the revised manuscript 
+as well. Regarding the measurement uncertainties, we account for them in the radar and radiometer simulations,
+but only in a highly idealized way in the lidar simulations.  Specifically, we assume random errors with 0.0 mean
+and 0.5dB standard deviation in the radar observations \cite{takahashi2008}, and a Noise-Equivalent-Delta-T (NEDT)
+of 1K, which is a readily achievable level for modern satellite radiometer \cite{draper2015}. A complex 
+model of space-borne lidar measurement uncertainties was developed in \cite{liu2006}, but its parameters are
+difficult to reliably quantify from theoretical considerations based on instrument expectations alone. Instead, we
+assume a simple multiplicative error model, with the multiplicative factor model as a log-normally distributed random
+variable with 0.0 mean and 1.0 standard deviation, which result in observation uncertainties similar to those in
+the radar observations. We acknowledge the limitations of this model in the manuscript.\\
+\newline
+\textit{2) You say in your abstract (and conclusion) that “the characteristics of 
 the instruments e.g.,  frequencies, sensitivities, etc. are set based on the 
@@
 Some of these (e.g., the radiometer frequencies) are mentioned in the text, 
-but it would be nice to see it all in one place.
-
-Minor Comments
-
-Line 38: Where is 8.0 dBZ coming from? There should be a citation here, or it should be stated that this is an estimate of the radar’s expected sensitivity.
-
-Lines 50-51: While your approach is justifiable and worthwhile, it is worth noting that your CloudSat IWC retrieval also makes numerous assumptions about ice cloud microphysical properties and structures (as laid out in Section 2a).
-So the problem is not entirely avoided by relying on “observations” instead of models, as the observations were of radar reflectivity, not of ice water contents or particle sizes.
-
-Lines 65-67: This is fine, but it should be noted that in doing this, you are underestimating the true retrieval uncertainty.
-The particle distribution assumptions and electromagnetic scattering properties that you use in training the retrieval algorithm will almost certainly not perfectly reflect the actual size and shape distributions of ice particles around the globe.
-
-Lines 83-90 and Eq.
-(1): I looked over the Testud et al.
-(2001) paper, and could not find anything that directly supported this statement, or any equations that resembled Eq.
-(1).
-Please add further explanation, either here or in an appendix, to help the reader follow your logic.
-
-Figure 2: Neither the caption nor the text adequately explains what is being plotted in this figure.
-What does each individual point represent?
-The text says, “Figure 2 is a scatter plot representation of the class multiplicative coefficient ai as a function of relative height,” but if that were the case, I would expect there to be many more points on the scatter plot (i.e.
-18 classes times a certain number of relative height levels, to yield hundreds of points).
-
-Lines 120-121: How is the data in Fig. 2 used to parameterize Nw as a function of height? Do you fit a line/curve to the data in Fig. 2? This needs to be clarified.
-
-Figure 3: It would be helpful to add vertical lines to these plots to indicate the detection thresholds you are assuming for the radar (8 dBZ?) and lidar.
-
-Line 220: It is my understanding that an ensemble Kalman smoother assumes normal distributions. How much does it matter that IWC is not normally distributed in nature?
-
-Typos
-
-Line 276: “wells” should be “well”
-
-Line 308: Figure 8 should be Figure 9
-
-Line 315: IWP should be IWC
-
-Line 361: eliminate “out
-
+but it would be nice to see it all in one place.}\\
+\newline
+We agree with the reviewer that, unfortunately, the AOS mission is still in flux. We added a table in the revised
+manuscript to summarize the assumed instrument characteristics.\\
+\newline
+\textit{\textbf{Minor Comments}
+Line 38: Where is 8.0 dBZ coming from? There should be a citation here, or it should be stated that this is 
+an estimate of the radar’s expected sensitivity.}\\
+\newline
+This is an estimate of the radar’s expected sensitivity provided by Japan Aerospace Exploration Agency (JAXA), which the AOS International
+partner contributing the Ku-band radar in the inclined orbit.\\
+\newline
+\textit{Lines 50-51: While your approach is justifiable and worthwhile, it is worth noting that your CloudSat IWC retrieval 
+also makes numerous assumptions about ice cloud microphysical properties and structures (as laid out in Section 2a). 
+So the problem is not entirely avoided by relying on “observations” instead of models, as the observations were of 
+radar reflectivity, not of ice water contents or particle sizes.}\\
+\newline
+We agree. We do not consider CloudSat retrievals uncertainties or bias free, but we prefer them because
+they are consistent with an observed vertical distribution of reflectivity at W-band. The vertical distributions of radar reflectivity
+derived from satellite or ground-based observations are generally difficult to reproduce by model simulations. We added a statement 
+to this effect in the revised manuscript.\\
+\newline
+\textit{Lines 65-67: This is fine, but it should be noted that in doing this, you are underestimating the true retrieval 
+uncertainty. The particle distribution assumptions and electromagnetic scattering properties that you use in training 
+the retrieval algorithm will almost certainly not perfectly reflect the actual size and shape distributions of ice 
+particles around the globe.}\\
+\newline
+We agree. As explained above, our  objective was to derive IWC distributions consistent with the observed vertical distribution of
+CloudSat reflectivity, without implying that the resulting distributions are uncertainty or bias free.  To make the performance
+assessment more realistic, in the revised version of the manuscript, we perturbed the $N_w$ value (which in the initial version was
+parametrized as a function of height) used in the derivation of IWC as a function of observed CloudSat reflectivity by a random log-normally
+distributed factor. This accounts for uncertainties in the PSDs (at least from the perspective of the normalized PSD framework) and 
+makes the evaluation realistic. Nevertheless, as observed by the reviewer, the electromagnetic scattering properties are 
+an additional source of uncertainty.  We included additional statements in the manuscript to describe $N_w$ perturbation approach and
+discuss uncertainties in the electromagnetic scattering properties that are not quantified in the manuscript.\\ 
+\newline
+\textit{Lines 83-90 and Eq. (1): I looked over the Testud et al. (2001) paper, and could not find anything that directly
+ supported this statement, or any equations that resembled Eq. (1). Please add further explanation, either here 
+ or in an appendix, to help the reader follow your logic.}\\
+\newline
+We apologize for the confusion. Testud et al. (2001) only show that normalization by $N_w$ ($N_0^*$ in their notation) explains the variability
+in Z-R relationships (fig. 9d). Ferreira et al. (2001) \cite{ferreira2001} make the more general statement that any two integrated rainfall parameters ($X$, $Y$) 
+may be related through a power law of the form $X = m N_w^{(1-n)} X^n$, where $m$ and $n$ are constants almost independent of $N_w$. Delanoe et al. (2014)
+ \cite{delanoe2014} show that the same relationship holds for ice PSDs. We added statements to this effect in the revised manuscript.\\
+\newline
+\textit{Figure 2: Neither the caption nor the text adequately explains what is being plotted in this figure. What does each 
+individual point represent? The text says, “Figure 2 is a scatter plot representation of the class multiplicative 
+coefficient ai as a function of relative height,” but if that were the case, I would expect there to be many more 
+points on the scatter plot (i.e. 18 classes times a certain number of relative height levels, to yield hundreds of points).}\\
+\newline
+We apologize for the confusion.  The points in figure 2 were derived from 36 classes.  We initially used 36 classes in the analysis, but
+we reduced the number of classes to 18 to make the classification and the Ensemble Kalman Filter more robust.  However, we kept the results from the
+36 classes in the $N_w$ parameterization because they enable a more accurate representation of the variability of the multiplicative coefficient.
+For each of the 36 classes, we regressed IWC from the 2C-ICE product against the observed CloudSat reflectivity and derived height of the associated
+mean IWC profile peak.  We then plotted the multiplicative coefficient as a function of the height of the mean IWC profile peak.  The height was
+evaluated relative to the freezing level.  We included clarifying statements in the manuscript.\\
+\newline
+\textit{Lines 120-121: How is the data in Fig. 2 used to parameterize Nw as a function of height? Do you fit a line/curve to the 
+data in Fig. 2? This needs to be clarified.}\\
+\newline
+Yes, We fit a line to determine the variation of $ln(a)$ as a function of height and then divide the slope by $(1-b)$ to get the variation of
+$ln(N_w)$ as a function of height.\\
+\newline
+\textit{Figure 3: It would be helpful to add vertical lines to these plots to indicate the detection thresholds you are 
+assuming for the radar (8 dBZ?) and lidar.}\\
+\newline
+We included vertical lines to indicate the detection thresholds.\\
+\newline
+\textit{Line 220: It is my understanding that an ensemble Kalman smoother assumes normal distributions. How much does it matter 
+that IWC is not normally distributed in nature?}\\
+\newline
+Yes, the ensemble Kalman smoother assumes normal distributions.  We derive a Kalman gain, i.e. a matrix of type $\mathbf{Cov(X,Y) Cov(Y,Y)^{-1}}$, 
+for every class, and although IWC is not normally distributed in nature, its conditional class distributions are not likely to be as skewed 
+as the entire distribution.
+This is most likely why the Kalman smoother produces rather accurate results when observations from all instruments is available.  
+Nevertheless, from the theoretical perspective, the Kalman Smoother results are suboptimal given the non-Gaussian distribution of IWC.  This
+does not necessarily mean that more accurate results may be derived in practice given the multitude of uncertainty factors that may
+impact the results.\\
+\newline
+\textbf{Typos}\textit{Line 276: “wells” should be “well”}\\
+\newline
+Thank you for pointing that out. We corrected it in the revised version of the manuscript.\\
+\newline
+\textit{Line 308: Figure 8 should be Figure 9}\\
+\newline
+Thank you for pointing that out. We corrected it in the revised version of the manuscript.\\
+\newline
+\textit{Line 315: IWP should be IWC}\\
+\newline
+Thank you for pointing that out. We corrected it in the revised version of the manuscript.\\
+\newline
+\textit{Line 361: eliminate “out}\\
+\newline
+Thank you for pointing that out. We corrected it in the revised version of the manuscript.\\
+\newline
 \pagebreak
@@
 We agree that 200,000 profiles is not a large dataset by today's standards.  However, these profiles are derived from a month of CloudSat observations.
-Specifically, all CloudSat and 2ICE product files from June 2019 were processed, and only profiles with no echo below the freezing level
+Specifically, all CloudSat and 2C-ICE product files from June 2019 were processed, and only profiles with no echo below the freezing level
 and the freezing level above ground were selected. The selection process significantly reduced the number of profiles. Nevertheless, there is sufficient
@@
 \begin{thebibliography}{999}
+
+\bibitem{takahashi2008}
+Takahashi N, Iguchi T. Characteristics of TRMM/PR system noise and their application to the rain detection algorithm. 
+IEEE transactions on geoscience and remote sensing. 2008 May 16;46(6):1697-704.
+
+\bibitem{draper2015}
+Draper DW, Newell DA, Wentz FJ, Krimchansky S, Skofronick-Jackson GM. The global precipitation measurement (GPM) 
+microwave imager (GMI): Instrument overview and early on-orbit performance. IEEE Journal of Selected Topics in 
+Applied Earth Observations and Remote Sensing. 2015 Mar 2;8(7):3452-62.
+\bibitem{liu2006}  
+Liu Z, Hunt W, Vaughan M, Hostetler C, McGill M, Powell K, Winker D, and  Hu Y, 
+Estimating random errors due to shot noise in backscatter lidar observations. 2006, Appl. Opt. 45, 4437-4447.
+
+\bibitem{ferreira2001}
+Ferreira, F., P. Amayenc, S. Oury, and J. Testud, 2001: Study and Tests of Improved Rain Estimates from the TRMM Precipitation Radar. 
+J. Appl. Meteor. Climatol., 40, 1878–1899, https://doi.org/10.1175
 

### 2023-05-04 RST/RSTManuscript.tex
Revision

@@
 However, although the radiometer-only estimation procedure is able to estimate the integrated amount of ice in clouds fairly
-wells, its ability to characterize the vertical distribution of ice in clouds is limited.  Figure \ref{f5} shows the
+well, its ability to characterize the vertical distribution of ice in clouds is limited.  Figure \ref{f5} shows the
 conditional vertical distributions of the estimated and true IWC for the 18 classes described in Section 2a 
@@
 agreement with the true IWC profiles than those derived from radiometer-only observations.  Furthermore,
-as seen in Figure \ref{f8} the synergistic IWC gravity center estimates are in very good agreement with 
+as seen in Figure \ref{f9} the synergistic IWC gravity center estimates are in very good agreement with 
 the true IWC gravity center.
@@
 \end{equation}
-where IWC$_{i}$ is the estimated IWP for the $i$-th sample, IWC$_{true,i}$ is the true IWC for the $i$-th sample,
+where IWC$_{i}$ is the estimated IWC for the $i$-th sample, IWC$_{true,i}$ is the true IWC for the $i$-th sample,
 $\overline{IWC}$ is the IWC mean, and $N$ is the size of the estimation dataset.  The classification
@@
 
-Further work is necessary out to assess the impact of sources of uncertainties such as potential biases in the
+Further work is necessary to assess the impact of sources of uncertainties such as potential biases in the
 forward models, variability in the

### 2023-05-04 RST/responseReviewer2.tex
Revision

@@
 \usepackage{subcaption}
+\usepackage{longtable}
 \author{Mircea Grecu and John E. Yorks}
@@
 \newline
-\textit{10. l. 203: I don't think that 200 000 profiles can be considered a large dataset by today's standards.}\\
+\textit{10. l. 203: I don't think that 200,000 profiles can be considered a large dataset by today's standards.}\\
 \newline
@@
 \newline
+
+\begin{longtable}[]{@{}llll@{}}
+
+    Instrument & Frequency/Wavelength & Noise & Sensitivity \\
+    \hline
+    \endhead
+    Lidar & 524 nm & LogNormal(0,1) & 1e-6 \\
+    Radar & 13.8 GHz & 0.5dB & 8 dBZ \\
+    Radiometer & 89 GHz & 1.0 K & N/A \\
+    & 183.31+/-0.2 GHz & 1.0 K & N/A \\
+    & 183.31+/-1.1 GHz & 1.0 K & N/A \\
+    & 183.31+/-2.8 GHz & 1.0 K & N/A \\
+    & 183.31+/-4.2 GHz & 1.0 K & N/A \\
+    & 183.31+/-6.8 GHz & 1.0 K & N/A \\
+    & 183.31+/-11 GHz & 1.0 K & N/A \\
+    & 325.15+/-1.5 GHz & 1.0 K & N/A \\
+    & 325.15+/-3.5 GHz & 1.0 K & N/A \\
+    & 183.31+/-9.5 GHz & 1.0 K & N/A \\
+   \hline 
+    \end{longtable}
 

### 2023-05-12 RST/responseReviewers.tex
review May 12

--- 2023-05-04 RST/responseReviewer2.tex
@@
 
-\noindent
-\textit{ \textbf{General Summary:}
+\noindent \textbf{Reviewer 1:}
+\noindent\textit{ \textbf{General Summary:}
 This a well-written paper that fits within the scope of Journal of Atmospheric and 
@@
 \newline
-We agree. We do not consider CloudSat retrievals uncertainties or bias free, but we prefer them because
+We agree. We do not consider CloudSat retrievals uncertainty or bias free, but we prefer them because
 they are consistent with an observed vertical distribution of reflectivity at W-band. The vertical distributions of radar reflectivity
@@
 \newline
-We agree. As explained above, our  objective was to derive IWC distributions consistent with the observed vertical distribution of
+We agree. As described explained above, our  objective was to derive IWC distributions consistent with the observed vertical distribution of
 CloudSat reflectivity, without implying that the resulting distributions are uncertainty or bias free.  To make the performance
@@
 \newline
-We apologize for the confusion.  The points in figure 2 were derived from 36 classes.  We initially used 36 classes in the analysis, but
-we reduced the number of classes to 18 to make the classification and the Ensemble Kalman Filter more robust.  However, we kept the results from the
-36 classes in the $N_w$ parameterization because they enable a more accurate representation of the variability of the multiplicative coefficient.
-For each of the 36 classes, we regressed IWC from the 2C-ICE product against the observed CloudSat reflectivity and derived height of the associated
-mean IWC profile peak.  We then plotted the multiplicative coefficient as a function of the height of the mean IWC profile peak.  The height was
-evaluated relative to the freezing level.  We included clarifying statements in the manuscript.\\
+We apologize for the confusion.  The points in figure 2 were derived from a classification of the IWC profiles into 36 classes.  
+We initially used 36 classes in the analysis, but
+we reduced the number of classes to 18 to make the classification and the Ensemble Kalman Filter more robust.  
+For consistency and simplicity in the revised manuscript we derive IWC-Z relationships 
+for each CloudSat radar bin, with the bins indexed relative to the freezing level bin.  The relative height is now defined as the distance
+between the height of the associated bin relative to the freezing level.  Results are quantitatively similar to those in the initial
+submission, especially given that the differences between the current $a$/$N_w$ parameterization and the initial one are smaller than 
+the random perturbations in $N_w$ that we introduced in the revised analysis to make the evaluation of the radar retrievals more realistic.\\
 \newline
@@
 \newline
-Yes, We fit a line to determine the variation of $ln(a)$ as a function of height and then divide the slope by $(1-b)$ to get the variation of
+Yes, we fit a line to determine the variation of $ln(a)$ as a function of height and then divide the slope by $(1-b)$ to get the variation of
 $ln(N_w)$ as a function of height.\\
@@
 \newline
-\pagebreak
-\noindent \textbf{Reviewer 2:}\textit{The study investigates the synergy between observations from elastic backscatter lidar, Ku-band radar,
+\newline
+%\newpage
+\clearpage
+\noindent \textbf{Reviewer 2:}
+\newline \textit{The study investigates the synergy between observations from elastic backscatter lidar, Ku-band radar,
 and sub-millimeter radiometer observations to retrieve ice concentrations. The topic is undoubtedly timely,
@@
 We thank the reviewer for their perspective and constructive suggestions.\\
-
+\newline
 \noindent \textit{First of all, I can't entirely agree with the study's framing. Currently, 
@@
 the convective precipitation.  Although it was originally envisioned that AOS would
-deploy two Ku-band radars, one in an inclined orbit and one in a polar orbit, only
-the inclined orbit Ku-band radar is currently planned for the AOS mission.  
+deploy two Ku-band radars, one in an inclined orbit and one in a polar orbit, currently, only
+the inclined orbit is expected to feature a Ku-band radar contributed by 
+Japan Aerospace Exploration Agency - JAXA.  
 The polar orbit radar will operate at higher-frequencies (Ka- or W-and) and 
@@
 due to attenuation. Moreover, the polar radar will have nadir-only capabilities,
-while the inclined orbit radar (which will be contributed by JAXA) will have a
+while the inclined orbit radar will have a
 swath of about 200 km.  Both the polar and inclined orbit radars will operate
@@
 (i.e. precipitation intensity, vertical velocities, etc.) characterizing the
-convective processes responsible for the creation of the anvil clouds.  Although 
+convective processes responsible for their creation.  Although 
 the polar orbit instruments are likely to enable the derivation of more accurate 
@@
 ice estimates, the lidar
-due to attenuation and the radar due to its limited sensitivity. \\
+due to attenuation and the radar due to its limited sensitivity and complex Particle
+Size Distribution (PSD) variability. \\
 
@@
 \centering
-\includegraphics[width=.31\linewidth]{IWP_LidarOnly.png}
-\includegraphics[width=.31\linewidth]{IWP_RadarOnly.png}
-\includegraphics[width=.31\linewidth]{IWP_RadiometerOnly.png}
-\includegraphics[width=.31\linewidth]{IWC_LidarOnly.png}
-\includegraphics[width=.31\linewidth]{IWC_RadarOnly.png}
-\includegraphics[width=.31\linewidth]{IWC_RadiometerOnly.png}
+\includegraphics[width=.31\linewidth]{ResponseFigs/lidarOnly_IWP.png}
+\includegraphics[width=.31\linewidth]{ResponseFigs/radarOnly_IWP.png}
+\includegraphics[width=.31\linewidth]{ResponseFigs/radiometer_IWP.png}
+\includegraphics[width=.31\linewidth]{ResponseFigs/lidarOnly_GC.png}
+\includegraphics[width=.31\linewidth]{ResponseFigs/radarOnly_GC.png}
+\includegraphics[width=.31\linewidth]{ResponseFigs/radiometer_GC.png}
 \caption{Top Row: Density plots of IWP retrievals from lidar-only (left), radar-only (middle), and radiometer-only (right) observations as a function
@@
 \newline
-We dropped the qualifier "high" from the title and the first sentence of the introduction.\\
+In AOS terminology, high clouds are convectively generated high clouds \cite{braun2022}. Such clouds are the result of strong vertical mass transport 
+from the lower troposphere and horizontal transport in the upper troposphere.  While preponderantly consisting of ice, high clouds may exhibit 
+liquid phased hydrometeors. To make it clear upfront that we are not considering mixed-phase clouds in this study, we included ice in the title.
+We included clarifying statements on the definition of high-clouds in the manuscript. \\
 \newline
@@
 \newline
-We initially started with 36 similarity classes but then reduced the number of classes to 18 to make the classification and 
-estimation process more robust. However, we kept the results from the 36 classes analysis in the derivation of the $N_w$ parameterization,
-as they enable the derivation of a more accurate parameterization than 18 classes. The relative height is the distance
-between the peak of the mean IWC peak and the freezing level.\\
+We initially started with 36 similarity classes and derived (through linear regressions in the log-log space) 
+an IWC-Z relationship for each class. The points in Fig. 2 represent 
+the coefficients $a$ as a function of the height of the class-average IWC peak for each class. We subsequently reduced the number of classes 
+to 18 to make the classification and 
+estimation process more robust. However, for consistency and simplicity in the revised manuscript we derive IWC-Z relationships 
+for each CloudSat radar bins, with the bins indexed relative to the freezing level bin.  The relative height is now defined as the distance
+between the height of the associated bin relative to the freezing level.  Results are quantitatively similar, especially given that the
+differences between the current $a$/$N_w$ parameterization and the one used in the initial submission are smaller than the random
+$N_w$ perturbations used in the analysis.\\
 \newline
@@
 Profiles that satisfy this requirement include anvil clouds, whose quantification in relation to the convective process with which they are associated
-is an AOS objection. We added a sentence in the manuscript to clarify this point.\\
+is an AOS objective. We added a sentence in the manuscript to clarify this point.\\
 \newline
@@
 We derive the IWC as a function of CloudSat $Z$ and $N_w$.  The $N_w$ is given by the vertical parameterization and perturbed using the random model
-described above and in the revised manuscript.  We added a sentence in the manuscript to clarify this point.\\
+described above and in the revised manuscript.  We added statements in the manuscript to clarify this point.\\
 \newline
@@
 and the freezing level above ground were selected. The selection process significantly reduced the number of profiles. Nevertheless, there is sufficient
-variability in the selected profiles to derive meaningful conclusions. We added a sentence in the manuscript to clarify this point.\\
+variability in the selected profiles to derive meaningful conclusions. We added statements in the manuscript to clarify this point.\\
 \newline
@@
   S. A. Braun, J. Yorks, T. Thorsen, D. Cecil and D. Kirschbaum, 
-  "NASA'S Earth System Observatory-Atmosphere Observing System," 
+  NASA'S Earth System Observatory-Atmosphere Observing System, 
   IGARSS 2022 - 2022 IEEE International Geoscience and Remote Sensing Symposium, 
   Kuala Lumpur, 
-  Malaysia, 2022, pp. 7391-7393, doi: 10.1109/IGARSS46834.2022.9884029.
+  Malaysia. 2022, pp. 7391-7393, doi: 10.1109/IGARSS46834.2022.9884029.
 

### 2023-05-12 RST/RSTManuscript.tex
review May 12

@@
 \usepackage{slashbox}
+\usepackage{longtable}
 % TWO-COLUMN JOURNAL PAGE LAYOUT---FOR AUTHOR USE ONLY
@@
 \section{Introduction}
-The future NASA Atmospheric Observing System (AOS) mission (Braun 2022) is expected to feature new 
+The future NASA Atmospheric Observing System (AOS) mission \citep{braun2022} is expected to feature new 
 combinations of observations that may be used to quantify the amounts of ice in high clouds and 
-characterize the microphysical properties of ice particles. These observations include backscatter
+characterize the microphysical properties of ice particles. In the AOS terminology, high clouds are convectively
+generated clouds \citep{braun2022}. They are the result of strong vertical mass transport accompanied by
+horizontal transport of hydrometeors in the upper troposphere.  As high clouds are of paramount importance in
+understanding the quantifying radiative processes, they constitute one of the AOS major objectives \citep{braun2022}.
+To achieve its objective, AOS will rely on a combination of active and passive observations onboard multiple spacecrafts
+in two different orbits.  One of the orbits is expected to be polar, while the other is expected to be inclined, which
+would allow the study of atmopsheric processes at sub-daily time scales, with an emphasis on deep
+convection, high clouds, and aerosols \citep{braun2022,yorks2022}.
+The AOS inclined observations will include backscatter
 from an elastic backscatter lidar \citep{weitkamp2006},
 Ku-band radar reflectivity, and submillimeter wave radiometer brightness temperature measurements. 
-The AOS mission design, while not necessarily optimal for cloud ice estimation, enables characterization 
-of the broader spectrum of cloud and precipitation processes. Nevertheless, these measurements are 
+While not necessarily optimal for cloud ice estimation, these measurements are 
 complimentary for synergistic characterization of ice clouds. 
-%While not optimal for cloud ice estimation, but for the characterization of a broader spectrum of 
-%cloud and precipitation processes, these observations are nevertheless synergistic from the 
-%characterization of ice clouds perspective. 
 That is, despite the fact that lidar observations 
 attenuate quickly in thick ice clouds and the Ku-band radar will not be able to detect clouds 
-characterized by an echo weaker than 8.0 dBZ, the active observations are expected to provide context 
+characterized by an echo weaker than 8.0 dBZ (which is the expected sensitivity of the radar in the inclined
+orbit), the active observations are expected to provide context 
 that may be incorporated into the radiometer retrievals. Herein, the term retrieval is defined as the 
@@
 products from the CloudSat (CS) mission (Stephens et al. 2002) to derive a database of ice microphysical
-properties and associated simulated lidar, radar, and radiometer observations.  This database is used to
+properties and associated simulated lidar, radar, and radiometer observations.  To account for variability in 
+the Particle Size Distributions (PSD) that may not be well represented in the 2C-ICE product, we use a simple
+but effective approach to perturb the PSD generalized intercepts from their nominal values.
+The resulting database is used to
 investigate the accuracy of estimated ice cloud properties from the simulated observations. Another major
-difference relative to previous studies is the unique combination of instruments investigated herein.
+difference relative to previous studies is the unique combination of instruments investigated herein. It should be
+mentioned that although based on observations rather than CRM simulations, the approach used in this study is
+idealized in many respects. As a consequence, the results presented may not unbiasedly extrapolate to practice.
+Nonetheless, the results are expected to provide useful insights into the potential of the AOS inclined observations and
+good first step towards the development of an operational synergistic retrieval of ice clouds from the AOS inclined observations.
 The article is organized as follows.  In Section 2, we describe the approach used to derive the ice properties and
@@
 \begin{equation}
-IWC=N_w^{1-b}aZ^b \label{eq:1}
+IWC=N_w^{1-b}a_{norm} Z^b \label{eq:1}
 \end{equation}
-perfectly explains the relationships between IWC and Z calculated from observed PSDs. Equation (\ref{eq:1}) is not 
+perfectly explains the relationships between IWC and Z calculated from observed PSDs. 
+Although the initial work of \cite{testud2001} was limited to rain and only hinted about the applicability
+of the normalization concept to ice observations, subsequent studies showed that parameterizations
+of type (\ref{eq:1}) extended to ice \citep{ferreira2001,delanoe2014}.
+Equation (\ref{eq:1}) is not 
 sufficient to derive accurate, unbiased estimates of ice water contents, because $N_w$ varies considerably in 
@@
 \end{figure}
-One may notice that the average IWC profiles in Figure 1 are characterized by different peak values and heights. 
-This facilitates a simple way to approximately reverse-engineer the "a priori" assumptions used in the CS 
-2C-ICE product and use them in formulation of the type described in Equation (1).  Specifically, the  
-derivation of relationships of the type $IWC=a_i Z^{b_i}$ for every class i may be used to study $a_i$ as a
-function of height. 
-Shown in Figure 2 is a  scatter plot representation of the class multiplicative coefficient $a_i$ as a function of
-relative height (defined as the absolute height relative to the height of the freezing
-level).  As 
-apparent in the figure, and as expected, $a_i$ exhibits a strong variation with the relative height. 
-Coefficient $b_i$ exhibits a height dependency as well (not shown), but the range of variation is significantly
-smaller, 
-almost zero relative to the mean value of $b$. Given that any deviation of the multiplicative coefficient in 
+
+A scatter-plot analysis (not shown) of relationships between the 2C-ICE IWC and the associated reflectivity suggests that
+$a$ is parameterized as a function of temperature (or equivalently, as a function of height relative to the freezing level) in 
+the default 2C-ICE retrievals. The default $a$ (the value the provides the IWC estimate prior to the ingestion of the lidar
+observations) may be simply estimated by regressing the 2C-ICE IWC on the associated CloudSat reflectivity as a function of 
+height relative the freezing level. Specifically, a regression of the type $ln(IWC)=ln(a)+b*ln(Z)$ is performed for 
+each bin relative to the zero degree bin is performed, and variation of $ln(a)$ is parameterized as a function of the
+relative height.  The results are shown in Figure 2. As apparent in the figure, and as expected, $a$ exhibits a strong variation 
+with the relative height.  Coefficient $b_i$ exhibits a height dependency as well (not shown), but the range of variation is 
+significantly smaller, almost zero relative to the mean value of $b$. Given that any deviation of the multiplicative coefficient in 
 an IWC-Z relation from an average is equivalent to a deviation of the associated $N_w$ from its mean value 
@@
 as a function of relative-height relationship.  We, therefore, use the data in Fig. 2 to parameterize $N_w$ as a 
-function of the relative height.
+function of the relative height.  Specifically, given that $(1-b)ln(N_w)+ln(a_{norm})=ln(a)$ and $ln(a)=ln(a_0)+s*H$, we 
+can express $N_w$ as a function of the relative height, slope $s$, and parameters $a_0$ and $a_{norm}$.
 
@@
     \centering
-    \includegraphics[width=0.75\textwidth,angle=0]{./Figs/fig02.png}\\
-    \caption{Multiplicative factor in the 2ICE IWC-relation as a function of the IWC profile
-    peak height.}\label{f2}
+    \includegraphics[width=0.75\textwidth,angle=0]{./Figs/fig02.rev.png}\\
+    \caption{Multiplicative factor in the 2ICE IWC-relation as a function of the height relative to the freezing level.}\label{f2}
 \end{figure}
@@
 submillimeter-wave calculations and may not be applicable at lidar's wavelength.  
+To improve the representation of microphysical variability in the study, we don't assume the values of $N_w$ given
+by the 2C-ICE based parametrization described above as the true values. Instead, we perturb them by multiplication 
+with a log-normally distributed random variable with 0.0 mean a standard deviation of 0.5.  The 
+perturbations are vertically correlated. Specifically, normal random variables with zero mean and a standard deviation
+of 1.0 are generated and then passed through a Gaussian smoothing filter \citep{nixon2019} with a size of two radar bins. The perturbation
+vector is then rescaled to have a standard deviation of 0.5.  The perturbations are then exponentiated and multiplied
+with the 2C-ICE based $N_w$ values to produce the final $N_w$ values.  The filter size and noise magnitude are chosen
+to roughly mimic the observed variability in $N_w$ derived from in situ observations of PSDs, such as those described
+in \cite{heymsfield2022impacts}. In addition, we are adding a random noise of
+0.0 mean and 0.5 dB standard deviation to the calculated reflectivity values.  Nevertheless, it should be noted that the
+uncertainties in the reflectivity calculations are likely to be even larger, given that the SSRGA theory although 
+quite accurate cannot possibly capture the entire range of uncertainties in the scattering properties of ice particles.
 
@@
     \centering
-    \includegraphics[width=0.75\textwidth,angle=0]{./Figs/fig03.png}\\
+    \includegraphics[width=0.75\textwidth,angle=0]{./Figs/fig03.rev.png}\\
     \caption{Simulated distributions of Ku-band radar reflectivity (left) and lidar backscatter (right)
     as function of height above 
-    the freezing level.}\label{f3}
+    the freezing level. Vertical lines indicated the assumed instrument sensitivities (8dBZ for the 
+    radar and 5e-7 m$^{-1}$ sr$^{-1}$ for the lidar, respectively.) }\label{f3}
 \end{figure}
@@
 ice scattering properties. To make the procedure physically meaningful, temperature rather than height is 
-used in the ice scattering-gas absorption collocation process.  The emissivities are randomly chosen between 
+used in the ice scattering-gas absorption collocation process.  The surface emissivities are randomly chosen between 
 0.8 and 1.0 and assumed constant for all radiometer frequencies.  Brightness temperatures are calculated at 
-89-, 183.31 ± 1.1, and 325.15 ± 1.5 GHz, which correspond to three of the 10 channels of the SAPHIR-NG 
-radiometer envisioned to be deployed in the AOS mission \citep{brogniez2022time}.  The other channels are 
-centered on the same water vapor absorption lines and are not likely to offer additional information in 
-this rather controlled experiment. Nevertheless, the other channels are expected to be useful in reducing the 
-uncertainties caused by variability in the vertical distribution of water vapor, which may be greater in 
-real life than in the simulated environment.
+%89-, 183.31 ± 1.1, and 325.15 ± 1.5 GHz,
+all frequencies of the 10 channels of the SAPHIR-NG 
+radiometer envisioned to be deployed in the AOS mission \citep{brogniez2022time}.  One of the frequencies is 
+89-GHz, while the others are centered on the 183.31, and 325.15 GHz water vapor absorption lines. Details are
+provided in Table \ref{t_sensors}.
 
@@
 
-The application of these steps produces a large dataset of approximately 200,000 cloud ice profiles and 
+\begin{table}
+    \caption{Assumed instrument characteristics.}\label{t_sensors}
+    \begin{center}
+    \begin{tabular}{llll}
+
+    Instrument & Frequency/Wavelength & Noise & Sensitivity \\
+    \hline
+    Lidar & 524 nm & LogNormal(0,1) & 5e-7 \\
+    \hline
+    Radar & 13.8 GHz & 0.5dB & 8 dBZ \\
+    \hline
+    Radiometer & 89 GHz & 1.0 K & N/A \\
+    & 183.31+/-0.2 GHz & 1.0 K & N/A \\
+    & 183.31+/-1.1 GHz & 1.0 K & N/A \\
+    & 183.31+/-2.8 GHz & 1.0 K & N/A \\
+    & 183.31+/-4.2 GHz & 1.0 K & N/A \\
+    & 183.31+/-6.8 GHz & 1.0 K & N/A \\
+    & 183.31+/-11 GHz & 1.0 K & N/A \\
+    & 325.15+/-1.5 GHz & 1.0 K & N/A \\
+    & 325.15+/-3.5 GHz & 1.0 K & N/A \\
+    & 183.31+/-9.5 GHz & 1.0 K & N/A \\
+   \hline \\
+
+    \end{tabular}
+\end{center}
+\end{table}
+
+The application of these steps produces a large dataset of approximately 800,000 cloud ice profiles and 
 associated lidar, radar and radiometer observations that may be used to investigate the synergy of the three 
@@
     \centering
-    \includegraphics[width=0.75\textwidth,angle=0]{./Figs/fig04.png}\\
+    \includegraphics[width=0.75\textwidth,angle=0]{./Figs/fig04.rev.png}\\
     \caption{Frequency plot of estimated IWP derived radiometer-observations as a function of the 
@@
     \centering
-    \includegraphics[width=0.75\textwidth,angle=0]{./Figs/fig05.png}\\
+    \includegraphics[width=0.75\textwidth,angle=0]{./Figs/fig05.rev.png}\\
     \caption{True and radiometer-only retrieved conditional mean IWC for the 18 classes described in 
@@
     \centering
-    \includegraphics[width=0.75\textwidth,angle=0]{./Figs/fig06.png}\\
+    \includegraphics[width=0.75\textwidth,angle=0]{./Figs/fig06.rev.png}\\
     \caption{Same as in Figure 4, but for the IWC gravity center.}\label{f6}
@@
     \centering
-    \includegraphics[width=0.75\textwidth,angle=0]{./Figs/fig07.png}\\
+    \includegraphics[width=0.75\textwidth,angle=0]{./Figs/fig07.rev.png}\\
     \caption{Same as in Figure \ref{f4}, but with the lidar and radar observations incorporated in the
@@
     \centering
-    \includegraphics[width=0.75\textwidth,angle=0]{./Figs/fig08.png}\\
+    \includegraphics[width=0.75\textwidth,angle=0]{./Figs/fig08.rev.png}\\
     \caption{Same as in Figure \ref{f5}, but with the lidar and radar observations incorporated in the
@@
     \centering
-    \includegraphics[width=0.75\textwidth,angle=0]{./Figs/fig09.png}\\
+    \includegraphics[width=0.75\textwidth,angle=0]{./Figs/fig09.rev.png}\\
     \caption{Same as in Figure \ref{f6}, but with the lidar and radar observations incorporated in the

### 2023-05-14 RST/responseReviewers.tex
review May 14

@@
 as well. Regarding the measurement uncertainties, we account for them in the radar and radiometer simulations,
-but only in a highly idealized way in the lidar simulations.  Specifically, we assume random errors with 0.0 mean
+and in a rather idealized way in the lidar simulations.  Specifically, we assume random errors with 0.0 mean
 and 0.5dB standard deviation in the radar observations \cite{takahashi2008}, and a Noise-Equivalent-Delta-T (NEDT)
-of 1K, which is a readily achievable level for modern satellite radiometer \cite{draper2015}. A complex 
+of 1K, which is a readily achievable level for modern satellite radiometers \cite{draper2015}. A complex 
 model of space-borne lidar measurement uncertainties was developed in \cite{liu2006}, but its parameters are
@@
 assume a simple multiplicative error model, with the multiplicative factor model as a log-normally distributed random
-variable with 0.0 mean and 1.0 standard deviation, which result in observation uncertainties similar to those in
-the radar observations. We acknowledge the limitations of this model in the manuscript.\\
+variable with 0.0 mean and 0.1 standard deviation, which result in observation uncertainties similar to those in
+the in evaluation study of the Cloud-Aerosol Transport System (CATS) lidar-system. 
+We acknowledge the limitations of this approach in the manuscript.\\
 \newline
@@
 \newline
-No, we don't include thermal noise in the retrieval experiments.  We acknowledge the lack of an uncertainty component in 
-the lidar model in section 2a.\\ 
+Regarding the measurement uncertainties, we account for them in the radar and radiometer simulations,
+and in a rather idealized way in the lidar simulations.  Specifically, we assume random errors with 0.0 mean
+and 0.5dB standard deviation in the radar observations \cite{takahashi2008}, and a Noise-Equivalent-Delta-T (NEDT)
+of 1K, which is a readily achievable level for modern satellite radiometers \cite{draper2015}.
+A complex model of space-borne lidar measurement uncertainties was developed in \cite{liu2006}, but its parameters are
+difficult to reliably quantify from theoretical considerations based on instrument expectations alone. Instead, we
+assume a simple multiplicative error model, with the multiplicative factor model as a log-normally distributed random
+variable with 0.0 mean and 0.1 standard deviation, which result in observation uncertainties similar to those in
+the in evaluation study of the Cloud-Aerosol Transport System (CATS) lidar-system. 
+We acknowledge the limitations of this approach in the manuscript.\\ 
 \newline

### 2023-05-14 RST/RSTManuscript.tex
review May 14

@@
 $N_w$ as a function of temperature based on the CloudSat 2C-ICE product \citep{deng2010tropical,deng2013evaluation}.  
-Specifically, we cluster, based on similarity, a large set 2C-ICE profiles into 18 classes using a k-Means 
+Specifically, we cluster, based on similarity (quantified through the evaluation of the Euclidian distance between profiles), 
+a large set 2C-ICE profiles into 18 classes using a k-Means 
 procedure. The mean IWC profiles associated with the 18 classes are shown in continuous lines in Figure 1.  
@@
 SSRGA scattering calculations \citep{heymsfield2022impacts}.  The open source software scatter-1.1 of \citep{scatter-1.1}
-is used to provide the actual scattering properties. The SSRGA theory was developed for millimeter and 
-submillimeter-wave calculations and may not be applicable at lidar's wavelength.  
+is used to provide the actual scattering properties.   
 To improve the representation of microphysical variability in the study, we don't assume the values of $N_w$ given
@@
 in \cite{heymsfield2022impacts}. In addition, we are adding a random noise of
-0.0 mean and 0.5 dB standard deviation to the calculated reflectivity values.  Nevertheless, it should be noted that the
+0.0 mean and 0.5 dB standard deviation to the calculated reflectivity values, which is consistent with the expected
+performance of a space-borne Ku-band radar system \citep{takahashi2008}.  Nevertheless, it should be noted that the
 uncertainties in the reflectivity calculations are likely to be even larger, given that the SSRGA theory although 
 quite accurate cannot possibly capture the entire range of uncertainties in the scattering properties of ice particles.
+Specifically, the SSRGA calculations were carried out assuming ice particles consisting of aggregates of 
+bullet rosettes, columnar crystals, and plates.
 
-The lidar observations are computed using the Mie solution included in the scatter-1.1 
+Moreover, the SSRGA theory was developed for millimeter and 
+submillimeter-wave calculations and may not be applicable at lidar's wavelength. 
+Therefore, lidar observations are computed using the Mie solution included in the scatter-1.1 
 package for a backscatter lidar, given that such a lidar is expected to be flown with a 
@@
 (ESO) Atmosphere Observing System (AOS) in an inclined and/or polar orbit \citep{yorks2022}.
-The lidar simulations are idealized in that they do not include random errors due to daytime 
+While uncertainties in the lidar forward model are rather complex and difficult to quantify,
+uncertainties in the lidar observations may be accounted for to some extent by including a
+multiplicative random noise factor in the calculated lidar backscatter values.  Specifically, the calculated
+lidar backscatter values are multiplied with a log-normally distributed random variable with 0.0 mean and a
+standard deviation of 0.1.  This results in uncertainties of about 10\% in the lidar backscatter values, which is
+consistent the performance of  Cloud-Aerosol Transport System (CATS) lidar system reported in \cite{pauly2019}.  
+Nevertheless, the lidar simulations are rather idealized in that they do not include random errors due to daytime 
 solar background noise or systematic errors such as calibration accuracy or lidar ratio
-assumptions. There are three main reasons this approach was taken: (1) it is important to
+assumptions. More advanced models of the observation errors exist \citep{liu2006}, but are not 
+considered here. There are three main reasons this approach was taken: (1) it is important to
 understand the limitations of combining these three datasets before factoring in 
@@
 agreement, which suggests that electromagnetic properties derived from Mie calculations 
-are adequate for practical applications. The lidar molecular backscatter and extinction
+are adequate for practical applications. 
+The lidar molecular backscatter and extinction
 are calculated using the lidar module of the CFMIP Observation Simulator Package (COSP; 
@@
 COSP simulations agree well with CALIOP observations. To account for multiple-scattering 
-in the lidar observations, we are using the multiscatter-1.2.11 model (Hogan 2015) of 
-Hogan and Battaglia (2008).
+in the lidar observations, we are using the multiscatter-1.2.11 model \citep{multiscatter-1.2.11} of 
+\cite{hogan2008lidar_ms}.
 
@@
 above the freezing level, which is consistent with the fact the lidar observations
-are strongly attenuated in the bottom part of the cloud.
+are strongly attenuated in the bottom part of the cloud.  The vertical resolutions of
+the radar and lidar observations are the same, i.e. 240 m, and the same as the resolution of the
+CS observations upon which they are based. In reality, the radar and lidar observations are likely to have
+different vertical resolutions and footprint sizes, which would need to be accounted for in a more 
+realistic study.  It should be noted though that the AOS radars and radiometers are expected to achieve
+Nyquist sampling, which enables resolution enhancement \citep{early2001image}.  Nevertheless, differences
+in the vertical resolutions of the radar and lidar observations and differences in the footprint sizes
+may deteriorate the performance of the retrieval algorithm.
 
@@
 89-GHz, while the others are centered on the 183.31, and 325.15 GHz water vapor absorption lines. Details are
-provided in Table \ref{t_sensors}.
+provided in Table \ref{t_sensors}.  Errors in the radiometer observations are modeled assuming a 
+Noise-Equivalent-Delta-T (NEDT) of 1K, which is a readily achievable level for modern satellite radiometers \cite{draper2015}.
 
@@
 
-    Instrument & Frequency/Wavelength & Noise & Sensitivity \\
+    Instrument & Frequency/Wavelength & Noise Assumptions& Sensitivity \\
     \hline
-    Lidar & 524 nm & LogNormal(0,1) & 5e-7 \\
+    Lidar & 524 nm & LogNormal(0,0.1) & 5e-7 m$^{-1}$sr$^{-1}$ \\
     \hline
@@
 
+\subsection{Active instrument retrievals}
+Although retrievals from the lidar-only or radar-only observations are not expected to be as accurate as those from
+radiometer-only observations, the lidar observations being subject to severe attenuation while the radar observations
+being limited by its sensitivity, they are nevertheless insightful.
 \subsection{Synergistic retrievals}

### 2023-05-16 RST/responseReviewers.tex
Review May 15

@@
 \author{Mircea Grecu and John E. Yorks}
-\title{Response to Reviewers}
+\title{Synergistic retrievals of ice in high clouds
+Response to Reviewers}
 \date{}
@@
 We thank the reviewer for the uncertainty discussion. Yes, indeed the resolution of the simulated 
-observations is that of the CloudSat, while the AOS instruments are expected to have different resolutions,
+observations is that of the CloudSat radar, while the AOS instruments are expected to have different resolutions,
 which might impact results. In the revised manuscript, we included statements to convey this aspect.
-We agree that the assumed electromagnetic scattering properties assumed in the study although
-reasonable may contain significant uncertainties.  We acknowledge this aspect in the revised manuscript 
-as well. Regarding the measurement uncertainties, we account for them in the radar and radiometer simulations,
+We agree that the assumed electromagnetic scattering properties assumed in the study, although
+reasonable, may contain significant uncertainties.  We acknowledge this aspect in the revised manuscript 
+as well. Regarding the measurement uncertainties, we realistically account for them in the radar and radiometer simulations,
 and in a rather idealized way in the lidar simulations.  Specifically, we assume random errors with 0.0 mean
@@
 assume a simple multiplicative error model, with the multiplicative factor model as a log-normally distributed random
-variable with 0.0 mean and 0.1 standard deviation, which result in observation uncertainties similar to those in
+variable with 0.0 mean and 0.1 standard deviation, which results in observation uncertainties similar to those in
 the in evaluation study of the Cloud-Aerosol Transport System (CATS) lidar-system. 
@@
 \newline
-This is an estimate of the radar’s expected sensitivity provided by Japan Aerospace Exploration Agency (JAXA), which the AOS International
-partner contributing the Ku-band radar in the inclined orbit.\\
+This is an estimate of the radar’s expected sensitivity provided by Japan Aerospace Exploration Agency (JAXA), which is
+the AOS International partner contributing the Ku-band radar in the inclined orbit.\\
 \newline
@@
 they are consistent with an observed vertical distribution of reflectivity at W-band. The vertical distributions of radar reflectivity
-derived from satellite or ground-based observations are generally difficult to reproduce by model simulations. We added a statement 
-to this effect in the revised manuscript.\\
+derived from satellite or ground-based observations are generally difficult to reproduce by model simulations.
+We added statements
+to this effect in the revised manuscript (bottom of page 3 and top part of page 4).\\
 \newline
@@
 CloudSat reflectivity, without implying that the resulting distributions are uncertainty or bias free.  To make the performance
-assessment more realistic, in the revised version of the manuscript, we perturbed the $N_w$ value (which in the initial version was
-parametrized as a function of height) used in the derivation of IWC as a function of observed CloudSat reflectivity by a random log-normally
+assessment more realistic, in the revised version of the manuscript, we perturbed the $N_w$ value (which in the first version 
+of the manuscript was
+parameterized as a function of height) used in the derivation of IWC as a function of observed CloudSat reflectivity by a random log-normally
 distributed factor. This accounts for uncertainties in the PSDs (at least from the perspective of the normalized PSD framework) and 
 makes the evaluation realistic. Nevertheless, as observed by the reviewer, the electromagnetic scattering properties are 
-an additional source of uncertainty.  We included additional statements in the manuscript to describe $N_w$ perturbation approach and
+an additional source of uncertainty.  We included additional statements in the manuscript 
+(bottom of page 8 and top of page 9) to describe $N_w$ perturbation approach and
 discuss uncertainties in the electromagnetic scattering properties that are not quantified in the manuscript.\\ 
@@
 \newline
-We apologize for the confusion. Testud et al. (2001) only show that normalization by $N_w$ ($N_0^*$ in their notation) explains the variability
+We apologize for the confusion. Testud et al. (2001) only show that normalization by $N_w$ ($N_0^*$ in their notation) 
+explains the variability
 in Z-R relationships (fig. 9d). Ferreira et al. (2001) \cite{ferreira2001} make the more general statement that any two integrated rainfall parameters ($X$, $Y$) 
 may be related through a power law of the form $X = m N_w^{(1-n)} X^n$, where $m$ and $n$ are constants almost independent of $N_w$. Delanoe et al. (2014)
- \cite{delanoe2014} show that the same relationship holds for ice PSDs. We added statements to this effect in the revised manuscript.\\
+ \cite{delanoe2014} show that the same relationship holds for ice PSDs. We added statements to this effect in the paragraph following
+ the definition of Equation (1).\\
 \newline
@@
 for each CloudSat radar bin, with the bins indexed relative to the freezing level bin.  The relative height is now defined as the distance
-between the height of the associated bin relative to the freezing level.  Results are quantitatively similar to those in the initial
-submission, especially given that the differences between the current $a$/$N_w$ parameterization and the initial one are smaller than 
-the random perturbations in $N_w$ that we introduced in the revised analysis to make the evaluation of the radar retrievals more realistic.\\
+between the height of the associated bin relative to the freezing level.  Results are rather similar, although the overall performance is
+degraded to some degree.\\
 \newline
@@
 Yes, we fit a line to determine the variation of $ln(a)$ as a function of height and then divide the slope by $(1-b)$ to get the variation of
-$ln(N_w)$ as a function of height.\\
+$ln(N_w)$ as a function of height. We added explanations in the paragraph preceding Equation (2).\\
 \newline
@@
 manuscript, to investigate the impact of variability in $N_w$, we randomly perturbed its parameterized value by a log-normally distributed
-random variable with a standard deviation of 0.5. This was achieved by generating one-dimensional vectors of independent random variables
+random variable with zero mean a standard deviation of 0.5. This was achieved by generating one-dimensional vectors of independent random variables
 with a standard normal distribution and applying a Gaussian smoothing filter \cite{nixon2019}.  The filter size was chosen to be two CloudSat radar bins.
@@
 \newline
-We agree with the reviewer. The revised manuscript includes radar-only, lidar-only, and radar-lidar retrievals in the analysis.  Shown 
+We agree with the reviewer. The revised manuscript includes radar-only, lidar-only, and other combinations of instruments retrievals
+ in the analysis.  Shown 
 in Fig. \ref{fig:IWP_IWC} are density plots of single-instrument retrievals of IWP and IWC as a function of their reference values.
@@
 we randomly perturb $N_w$ by multiplying its parameterized value with log-normally distributed
-random variable with a standard deviation of 0.5.  As explained above, we use a Gaussian smoothing filter to impose a vertical auto-correlation in
-$N_w$ similar to that observed in field experiments.\\
+random variable with zero mean a standard deviation of 0.5.  As explained above, we use a Gaussian smoothing filter to impose a
+vertical auto-correlation in
+$N_w$ similar to that observed in field experiments. This is explained in lines 153 through 162 of the revised manuscript.\\
 \newline
@@
 liquid phased hydrometeors. To make it clear upfront that we are not considering mixed-phase clouds in this study, we included ice in the title.
-We included clarifying statements on the definition of high-clouds in the manuscript. \\
+We included clarifying statements on the definition of high-clouds in the manuscript introduction. \\
 \newline
@@
 the in evaluation study of the Cloud-Aerosol Transport System (CATS) lidar-system. 
-We acknowledge the limitations of this approach in the manuscript.\\ 
+We acknowledge the limitations of this approach in the revised manuscript.  Statements to this effect are included in
+the paragraphs devoted to the descriptions of the forward models. \\ 
 \newline
@@
 The impact of normalization on Z-IWC relationship is shown in \cite{delanoe2014}. Soft spheres were assumed and the Mie solution was used in \cite{delanoe2014}
-to derive reflecitivities from observed Particle Size Distributions, but \cite{chase2021}  used more advanced backscattering models to calculate Z from
+to derive reflectivities from observed Particle Size Distributions, but \cite{chase2021}  used more advanced backscattering models to calculate Z from
 observed PSDs.  Their calculations are publicly available at https://github.com/dopplerchase/Chase\_et\_al\_2021\_NN and may be used to confirm the contraction
@@
 to model the backscattering property of ice particles but get a normalization behavior similar to that anticipated by Testud et al. (2001),
-shown in \cite{delanoe2014} and present in the \cite{chase2021} database.\\
+shown in \cite{delanoe2014} and exhbited in the \cite{chase2021} database.
+We added statements to this effect in the paragraph following the definition of Equation (1).\\
 \newline
@@
 estimation process more robust. However, for consistency and simplicity in the revised manuscript we derive IWC-Z relationships 
-for each CloudSat radar bins, with the bins indexed relative to the freezing level bin.  The relative height is now defined as the distance
-between the height of the associated bin relative to the freezing level.  Results are quantitatively similar, especially given that the
-differences between the current $a$/$N_w$ parameterization and the one used in the initial submission are smaller than the random
-$N_w$ perturbations used in the analysis.\\
+for each CloudSat radar bins, with the bins indexed relative to the freezing level bin.  The relative height is defined as the distance
+between the height of the associated bin relative to the freezing level.  Results are rather similar, although the overall performance is
+degraded to some degree.\\
 \newline
@@
 \newline
-We selected only CloudSat reflectivity profiles with no echo below the freezing level and the freezing level above ground.  
+We selected only CloudSat reflectivity profiles with no echo at and below the freezing level.  
 Profiles that satisfy this requirement include anvil clouds, whose quantification in relation to the convective process with which they are associated
@@
 We derive the IWC as a function of CloudSat $Z$ and $N_w$.  The $N_w$ is given by the vertical parameterization and perturbed using the random model
-described above and in the revised manuscript.  We added statements in the manuscript to clarify this point.\\
+described above and in the revised manuscript.  Specifically, given $N_w$ and the associated $Z$ at W-band, we estimate $IWC/N_w$ as a function of $Z/N_w$ using the
+normalized tables and then derive $IWC$ from $IWC/N_w$ and $N_w$. We added a statement in the manuscript to clarify this point.\\
 \newline
@@
 We agree that 200,000 profiles is not a large dataset by today's standards.  However, these profiles are derived from a month of CloudSat observations.
-Specifically, all CloudSat and 2C-ICE product files from June 2019 were processed, and only profiles with no echo below the freezing level
-and the freezing level above ground were selected. The selection process significantly reduced the number of profiles. Nevertheless, there is sufficient
-variability in the selected profiles to derive meaningful conclusions. We added statements in the manuscript to clarify this point.\\
+Specifically, all CloudSat and 2C-ICE product files from June 2019 were processed, and only profiles with no cloud echo at 
+and below the freezing level
+were selected. Moreover, in the first version of the manuscript, we imposed the rather artificial condition that at least 
+three CloudSat reflectivity pixels above the freezing level be classified as cloud for a profile to be select.
+These two conditions resulted in a rather small number of profiles. 
+We removed the second condition is the selection of profiles used in the revised analysis. This resulted in 
+about 800,000 of profiles. While 75\% of these profiles consist of at most three pixels, there is sufficient
+vertical variability in the selected profiles to derive meaningful conclusions. 
+We added statements in the manuscript to clarify this aspect.\\
 \newline
@@
 \newline
-
-\begin{longtable}[]{@{}llll@{}}
-
-    Instrument & Frequency/Wavelength & Noise & Sensitivity \\
-    \hline
-    \endhead
-    Lidar & 524 nm & LogNormal(0,1) & 1e-6 \\
-    Radar & 13.8 GHz & 0.5dB & 8 dBZ \\
-    Radiometer & 89 GHz & 1.0 K & N/A \\
-    & 183.31+/-0.2 GHz & 1.0 K & N/A \\
-    & 183.31+/-1.1 GHz & 1.0 K & N/A \\
-    & 183.31+/-2.8 GHz & 1.0 K & N/A \\
-    & 183.31+/-4.2 GHz & 1.0 K & N/A \\
-    & 183.31+/-6.8 GHz & 1.0 K & N/A \\
-    & 183.31+/-11 GHz & 1.0 K & N/A \\
-    & 325.15+/-1.5 GHz & 1.0 K & N/A \\
-    & 325.15+/-3.5 GHz & 1.0 K & N/A \\
-    & 183.31+/-9.5 GHz & 1.0 K & N/A \\
-   \hline 
-    \end{longtable}
 

### 2023-05-16 RST/RSTManuscript.tex
Review May 15

@@
 
-\title{Synergistic retrievals of ice in clouds from elastic backscatter lidar, 
+\title{Synergistic retrievals of ice in high clouds from elastic backscatter lidar, 
 Ku-band radar and submillimeter wave radiometer observations}
@@
 in two different orbits.  One of the orbits is expected to be polar, while the other is expected to be inclined, which
-would allow the study of atmopsheric processes at sub-daily time scales, with an emphasis on deep
+would allow the study of atmospheric processes at sub-daily time scales, with an emphasis on deep
 convection, high clouds, and aerosols \citep{braun2022,yorks2022}.
@@
 of type (\ref{eq:1}) extended to ice \citep{ferreira2001,delanoe2014}.
-Equation (\ref{eq:1}) is not 
+However, equation (\ref{eq:1}) is not 
 sufficient to derive accurate, unbiased estimates of ice water contents, because $N_w$ varies considerably in 
@@
 is used to provide the actual scattering properties.   
-To improve the representation of microphysical variability in the study, we don't assume the values of $N_w$ given
+To improve the representation of microphysical variability in the study, we do not assume the values of $N_w$ given
 by the 2C-ICE based parametrization described above as the true values. Instead, we perturb them by multiplication 
@@
 above the freezing level, which is consistent with the fact the lidar observations
-are strongly attenuated in the bottom part of the cloud.  The vertical resolutions of
+are strongly attenuated in the bottom part of the cloud.  It should be noted that, consistently with the objective of
+ this study, only CS reflectivity profiles with no echo at or below the freezing level were selected
+and used in the calculation of Ku-band reflectivity distributions shown in Figure \ref{f3}.
+The vertical resolutions of
 the radar and lidar observations are the same, i.e. 240 m, and the same as the resolution of the
@@
 radiometer envisioned to be deployed in the AOS mission \citep{brogniez2022time}.  One of the frequencies is 
-89-GHz, while the others are centered on the 183.31, and 325.15 GHz water vapor absorption lines. Details are
+89-GHz, while the others are centered on the 183.31, and 325.15 GHz water vapor absorption lines. Details regarding
+the radiometer and the active instruments are
 provided in Table \ref{t_sensors}.  Errors in the radiometer observations are modeled assuming a 
@@
 \item Estimation of IWC and related PSD parameters from CS W-band radar observations, using the tables 
-constructed
-in Step 1 and parameterization derived in Step 2.
+constructed Step 1, and $N_w$ profiles derived through parameterization obtained in Step 2 and perturbed using the random model 
+described above. Specifically, given $N_w$ and the associated $Z$ at W-band, we estimate $IWC/N_w$ as a function of $Z/N_w$ using the
+normalized tables and then derive $IWC$ from $IWC/N_w$ and $N_w$.
 \item Calculation of lidar, Ku-band radar and radiometer observations from the estimates derived in Step 3 
@@
 the true IWC gravity center exhibits quite a broad distribution, the one retrieved from the radiometer-only 
-observations exhibits a multimodal narrow distribution. Moreover, there is almost no correlation between the retrieved 
-and the true IWC gravity center.  This is another indication that, while the total amount of ice may be reasonably 
-estimated from radiometer-only observations, its vertical distribution can not be 
-accurately determined from 
-radiometer-only observations.
+observations exhibits a multimodal narrow distribution. Moreover, there correlation between the retrieved 
+and the true IWC gravity center is rather low.  This is another indication that, while the total amount of ice may be 
+reasonably estimated from radiometer-only observations, its vertical distribution can not be 
+accurately determined from radiometer-only observations.
 
@@
 \subsection{Active instrument retrievals}
+
 Although retrievals from the lidar-only or radar-only observations are not expected to be as accurate as those from
 radiometer-only observations, the lidar observations being subject to severe attenuation while the radar observations
-being limited by its sensitivity, they are nevertheless insightful.
+being limited by its sensitivity, they are nevertheless insightful. This is because quantifying the limitations of
+retrievals from active instruments may be used to better assess  the benefits of synergistic retrievals.  Shown in
+top of row of Figure \ref{f7} are the distributions of IWP estimated from lidar-only and radar-only observations 
+as a function of their true values. As apparent in the figure, and as expected, the lidar-only retrievals tend to
+be accurate for IWP values smaller than 50 g/m$^2$, while the radar-only retrievals tend to be reliable only for large
+IWP values on the order of hundreds of g/m$^2$.  Not that the radar-only retrievals exhibit a bimodal distributions,
+with a broad distribution of real IWP values associated with a single retrieved IWP value.  This is a consequence
+of the fact that there is a large number of atmospheric profiles characterized by not necessarily small IWP values that
+are not associated with detectable radar signals. The results shown in the top row of Figure \ref{f7} are conditional on
+either the CALIOP lidar or CS radar observations detecting ice, and so are the radar-only retrievals. However, the 
+indiscriminate application of a Ku-band radar-only retrieval procedure to the entire dataset may result in significant
+overestimation because clear skies would be associated with the same IWP value as ice clouds with reflectivity signal 
+below the detection threshold. On the other hand, limiting the application of the radar-only retrieval procedure to atmospheric profiles
+with detectable ice clouds would result in a significant underestimation, as a large number of atmospheric profiles
+with ice clouds would be associated with zero IWP.  From this perspective, the derivation of  Ku-band radar-only ice estimates
+is not useful if coincident radiometer observations are available.
+
+Shown in the bottom row of Figure \ref{f7} are the distributions of IWC gravity centers
+estimated from lidar-only and radar-only observations as a function of their true values. Results are similar to those
+obtained for IWP, in the sense that the radar-only retrieval distribution is bimodal and reliable only for ice clouds
+with Ku-band radar observations above the detection threshold.
+The lidar-only retrievals produce a much broader IWC gravity center distribution, with values that exhibit moderate 
+correlation with the true IWC gravity centers.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=.481\linewidth]{Figs/fig07a.rev.png}
+    \includegraphics[width=.481\linewidth]{Figs/fig07b.rev.png}
+    \includegraphics[width=.481\linewidth]{Figs/fig07c.rev.png}
+    \includegraphics[width=.481\linewidth]{Figs/fig07d.rev.png}
+    \caption{Top Row: Density plots of IWP retrievals from lidar-only (left), and radar-only (right) observations as a function
+    of true IWP. Bottom Row: Same as in the top row but for IWC gravity center.}\label{f7}
+\end{figure}
+
 \subsection{Synergistic retrievals}
@@
 
-Shown in Figure \ref{f7} is the distribution of the synergistic IWP estimates as a function of their true
+Shown in Figure \ref{f8} is the distribution of the synergistic IWP estimates as a function of their true
 values.  As apparent in the figure, the synergistic IWP estimates are more accurate than the radiometer-only
-estimates. At the same time, as apparent in Figure \ref{f8}, the retrieved conditional mean IWC for the 
-18 classes described section 2a and shown in Figure \ref{f1} are in significantly better (almost perfect)
-agreement with the true IWC profiles than those derived from radiometer-only observations.  Furthermore,
-as seen in Figure \ref{f9} the synergistic IWC gravity center estimates are in very good agreement with 
-the true IWC gravity center.
+estimates. At the same time, as apparent in Figure \ref{f9}, the retrieved conditional mean IWC for the 
+18 classes described section 2a and shown in Figure \ref{f1} are in significantly better
+agreement with the true IWC profiles than those derived from radiometer-only observations.  Moreover,
+as seen in Figure \ref{f10} the synergistic IWC gravity center estimates are in much better agreement with 
+the true IWC gravity center than those derived from single-instrument observations.
 
@@
     \centering
-    \includegraphics[width=0.75\textwidth,angle=0]{./Figs/fig07.rev.png}\\
+    \includegraphics[width=0.75\textwidth,angle=0]{./Figs/fig08.rev.png}\\
     \caption{Same as in Figure \ref{f4}, but with the lidar and radar observations incorporated in the
-    retrievals.}\label{f7}
+    retrievals.}\label{f8}
 \end{figure}
@@
     \centering
-    \includegraphics[width=0.75\textwidth,angle=0]{./Figs/fig08.rev.png}\\
+    \includegraphics[width=0.75\textwidth,angle=0]{./Figs/fig09.rev.png}\\
     \caption{Same as in Figure \ref{f5}, but with the lidar and radar observations incorporated in the
-    retrievals.}\label{f8}
+    retrievals.}\label{f9}
 \end{figure}
@@
 
+Additional insight may be gained by examining the distribution of the retrieved of IWP and IWC gravity center as
+a function of their true values.  These distributions are show in Figure \ref{f11} for the lidar-radiometer
+and radar-radiometer retrievals.  As apparent in the figure, the inclusion of radiometer observations improves both
+the lidar and radar retrievals. However, neither of the two instrument combinations is able to provide estimates
+that are as accurate as those derived from all instruments.  Moreover, the lidar-radiometer IWP retrievals appear to be
+less accurate than lidar-only retrievals. This may be an artifact of statistical inversion methodology, which may 
+make suboptimal use of correlations in the observations and improve the accuracy of the estimates for some type of 
+profiles at the expense of others.  While other statistical inversion methodologies may be able to eliminate this
+artifact, given the complexity of the problem and the multiple sources of uncertainty, no better methodology is
+obvious yet.
+
+\begin{figure}[t]
+    \centering
+    \includegraphics[width=0.75\textwidth,angle=0]{./Figs/fig10.rev.png}\\
+    \caption{Same as in Figure \ref{f6}, but with the lidar and radar observations incorporated in the
+    retrievals.}\label{f10}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=.481\linewidth]{Figs/fig11a.rev.png}
+    \includegraphics[width=.481\linewidth]{Figs/fig11b.rev.png}
+    \includegraphics[width=.481\linewidth]{Figs/fig11c.rev.png}
+    \includegraphics[width=.481\linewidth]{Figs/fig11d.rev.png}
+    \caption{Top Row: Density plots of IWP retrievals from lidar-radiometer (left), and radar-radiometer (right) observations as a function
+    of true IWP. Bottom Row: Same as in the top row but for IWC gravity center.}\label{f11}
+\end{figure}
+
 %Therefore, results in table
@@
 
-\begin{figure}[t]
-    \centering
-    \includegraphics[width=0.75\textwidth,angle=0]{./Figs/fig09.rev.png}\\
-    \caption{Same as in Figure \ref{f6}, but with the lidar and radar observations incorporated in the
-    retrievals.}\label{f9}
-\end{figure}
-
 \begin{table}[t]
@@
 \begin{center}
-\begin{tabular}{c|cccc}
+\begin{tabular}{c|ccccccc}
 \hline\hline
-\backslashbox{Score}{Instruments} & Radar-Radiometer & Lidar-Radiometer & Radar-Lidar-Radiometer\\
-Score& & & &\\
+\backslashbox{Score}{Instruments} & Lidar-& Radar- & Radiometer- & Lidar- & 
+Lidar- & Radar-	& Lidar-\\
+ & Only & Only & Only & Radar & Radiometer& Radiometer & Radar-Radiometer\\
 \hline
 % NRMS & 0.73 &  0.59  &  0.32 & 0.32 \\
-NRMS & 0.73 &  0.59  &  0.32 & 0.22 \\
-Class. Accurracy & 0.39 & 0.48 & 0.92 & 0.94 \\
+NRMS & 0.84	& 0.67 & 0.64 & 0.61 &	0.56 &	0.54 &	0.48 \\
+%0.73 &  0.59  &  0.32 & 0.22 \\
+Class. Accurracy & 0.40	& 0.39 &	0.35 &	0.49 &	0.62& 0.53	& 0.64 \\
 \hline

### 2023-05-16 RST/responseReviewers.tex
Review May 16

@@
 We added statements
-to this effect in the revised manuscript (bottom of page 3 and top part of page 4).\\
+to this effect in the revised manuscript (lines 67-69 in the Introduction).\\
 \newline
@@
 For consistency and simplicity in the revised manuscript we derive IWC-Z relationships 
-for each CloudSat radar bin, with the bins indexed relative to the freezing level bin.  The relative height is now defined as the distance
+for each CloudSat radar bin, with the bins indexed relative to the freezing level bin.  The relative height is defined as the distance
 between the height of the associated bin relative to the freezing level.  Results are rather similar, although the overall performance is
-degraded to some degree.\\
+degraded to some degree due to the inclusion of the $N_w$ perturbation component.\\
 \newline
@@
 Yes, we fit a line to determine the variation of $ln(a)$ as a function of height and then divide the slope by $(1-b)$ to get the variation of
-$ln(N_w)$ as a function of height. We added explanations in the paragraph preceding Equation (2).\\
+$ln(N_w)$ as a function of height. We added explanations to this effect in the revised manuscript (lines 114-124).\\
 \newline
@@
 vertical auto-correlation in
-$N_w$ similar to that observed in field experiments. This is explained in lines 153 through 162 of the revised manuscript.\\
+$N_w$ similar to that observed in field experiments. This is explained in lines 155-165 of the revised manuscript.\\
 \newline
@@
 shown in \cite{delanoe2014} and exhbited in the \cite{chase2021} database.
-We added statements to this effect in the paragraph following the definition of Equation (1).\\
+We added statements to this effect in the paragraphs describing Equation (1).\\
 \newline
@@
 between the height of the associated bin relative to the freezing level.  Results are rather similar, although the overall performance is
-degraded to some degree.\\
+degraded to some degree due to the inclusion of the $N_w$ perturbation.\\
 \newline

### 2023-05-16 RST/RSTManuscript.tex
Review May 16

@@
 understanding the quantifying radiative processes, they constitute one of the AOS major objectives \citep{braun2022}.
-To achieve its objective, AOS will rely on a combination of active and passive observations onboard multiple spacecrafts
+To achieve its objectives, AOS will rely on a combination of active and passive observations onboard multiple spacecrafts
 in two different orbits.  One of the orbits is expected to be polar, while the other is expected to be inclined, which
@@
 While not necessarily optimal for cloud ice estimation, these measurements are 
-complimentary for synergistic characterization of ice clouds. 
+complimentary and enable the synergistic characterization of ice clouds. 
 That is, despite the fact that lidar observations 
@@
 Nonetheless, the results are expected to provide useful insights into the potential of the AOS inclined observations and
-good first step towards the development of an operational synergistic retrieval of ice clouds from the AOS inclined observations.
+good first step towards the development of an operational synergistic retrieval of ice clouds from AOS inclined observations.
 The article is organized as follows.  In Section 2, we describe the approach used to derive the ice properties and
@@
 PSD intercept is defined as $N_w=\frac {4^4} {\pi \rho_w} \frac {IWC} {D_m^4}$, where $IWC$ is the ice water 
-content associated with the PSD, and $D_m$ is the mass weighted mean diameter.  \cite{testud2001} showed 
+content associated with the PSD, and $D_m$ is the mass weighted mean diameter.  Based on the work of \cite{testud2001},
+\cite{ferreira2001} and \cite{delanoe2014} showed 
 that the variability in IWC reflectivity (Z) relationships may be fully explained by variability in $N_w$, and 
@@
 \end{equation}
-perfectly explains the relationships between IWC and Z calculated from observed PSDs. 
-Although the initial work of \cite{testud2001} was limited to rain and only hinted about the applicability
-of the normalization concept to ice observations, subsequent studies showed that parameterizations
-of type (\ref{eq:1}) extended to ice \citep{ferreira2001,delanoe2014}.
+(where $a_{norm}$ and $b$ are constants) explains almost perfectly the relationships between IWC and Z calculated from observed PSDs.
 However, equation (\ref{eq:1}) is not 
@@
 $N_w$ as a function of temperature based on the CloudSat 2C-ICE product \citep{deng2010tropical,deng2013evaluation}.  
-Specifically, we cluster, based on similarity (quantified through the evaluation of the Euclidian distance between profiles), 
+A scatter-plot analysis of relationships between the 2C-ICE IWC and the associated reflectivity suggests that
+the multiplicative coefficient in a power low ice-reflectivity relationship is parameterized as a function of temperature 
+(or equivalently, as a function of height relative to the freezing level) in 
+the default 2C-ICE retrievals. The default multiplicative coefficient (the value the provides the IWC estimate prior
+to the ingestion of the lidar observations) may be simply estimated by regressing the 2C-ICE IWC on the associated CloudSat reflectivity 
+as a function of height relative the freezing level. Specifically, a regression of the type $ln(IWC)=ln(a)+b*ln(Z)$ is performed for 
+each bin referenced relative to the zero degree bin is performed, and the variation of $ln(a)$ is parameterized as a function of the
+relative height.  The results are shown in Figure \ref{f2}. As apparent in the figure, and as expected, $a$ exhibits a strong variation 
+with the relative height. Given that any deviation of the multiplicative coefficient in 
+an IWC-Z relation from an average is equivalent to a deviation of the associated $N_w$ from its mean value 
+\citep{ferreira2001,delanoe2014}, the variation of $a$ as a function of relative-height may be converted into an $N_w$ 
+as a function of relative-height relationship.  We, therefore, use the data in Figure \ref{f2} to parameterize $N_w$ as a 
+function of the relative height.  Specifically, given that $(1-b)ln(N_w)+ln(a_{norm})=ln(a)$ and $ln(a)=ln(a_0)+s*H$, with 
+$a_0$ and $s$ the intercept and the slope of a regression for the data in Figure \ref{f2}, we 
+can express $N_w$ as a function of the relative height, slope $s$, and parameters $a_0$ and $a_{norm}$.
+
+\begin{figure}[t]
+    \centering
+    \includegraphics[width=0.75\textwidth,angle=0]{./Figs/fig02.rev.png}\\
+    \caption{Multiplicative factor in the 2ICE IWC-relation as a function of the height relative to the freezing level.}\label{f2}
+\end{figure}
+
+To investigate the variability in the vertical distribution of the 2C-ICE estimates and their consistency with our estimates, 
+we cluster, based on similarity  (quantified through the evaluation of the Euclidian distance between profiles), 
 a large set 2C-ICE profiles into 18 classes using a k-Means 
-procedure. The mean IWC profiles associated with the 18 classes are shown in continuous lines in Figure 1.  
+procedure. The mean IWC profiles associated with the 18 classes are shown in continuous lines in Figure \ref{f1}.  
 Our estimates, derived using PSD assumptions and electromagnetic scattering calculations that enable 
 accurate and physically consistent simulations of radar observations at Ku-band and radiometer observations of 
-submillimeter-wave frequencies are also shown in Figure 1. These estimates are based on the self-similarity 
+submillimeter-wave frequencies are also shown in the figure. These estimates are based on the self-similarity 
 Rayleigh-Gans approximation (SSRGA) of \cite{hogan2017ssrga}. Details regarding the
-estimation process are provided in the following paragraphs.  As apparent in Figure 1, the CS and SSRGA estimates are in good 
+estimation process are provided in the following paragraphs.  As apparent in Figure \ref{f1}, the CS and SSRGA estimates are in good 
 agreement.  Some discrepancies due to differences between the SSRGA $N_w$ parameterization and the CS 2C-ICE 
@@
 the investigation of synergistic lidar, Ku-band radar and submillimeter-wave radiometer retrievals, because the 
-outcome is not likely to be sensitive to such details.
+outcome is not likely to be sensitive to such details. Also apparent in Figure \ref{f1} is the fact that there is significant
+variability in the vertical distribution of the 2C-ICE IWC estimates, which makes the estimation of IWC profiles from passive-only 
+observations challenging.
+
 \begin{figure}[t]
@@
 
-A scatter-plot analysis (not shown) of relationships between the 2C-ICE IWC and the associated reflectivity suggests that
-$a$ is parameterized as a function of temperature (or equivalently, as a function of height relative to the freezing level) in 
-the default 2C-ICE retrievals. The default $a$ (the value the provides the IWC estimate prior to the ingestion of the lidar
-observations) may be simply estimated by regressing the 2C-ICE IWC on the associated CloudSat reflectivity as a function of 
-height relative the freezing level. Specifically, a regression of the type $ln(IWC)=ln(a)+b*ln(Z)$ is performed for 
-each bin relative to the zero degree bin is performed, and variation of $ln(a)$ is parameterized as a function of the
-relative height.  The results are shown in Figure 2. As apparent in the figure, and as expected, $a$ exhibits a strong variation 
-with the relative height.  Coefficient $b_i$ exhibits a height dependency as well (not shown), but the range of variation is 
-significantly smaller, almost zero relative to the mean value of $b$. Given that any deviation of the multiplicative coefficient in 
-an IWC-Z relation from an average is equivalent to a deviation of the associated $N_w$ from its mean value 
-(Testud et al. 2001), the variation of $a$ as a function of relative-height may be converted into a $N_w$ 
-as a function of relative-height relationship.  We, therefore, use the data in Fig. 2 to parameterize $N_w$ as a 
-function of the relative height.  Specifically, given that $(1-b)ln(N_w)+ln(a_{norm})=ln(a)$ and $ln(a)=ln(a_0)+s*H$, we 
-can express $N_w$ as a function of the relative height, slope $s$, and parameters $a_0$ and $a_{norm}$.
-
-\begin{figure}[t]
-    \centering
-    \includegraphics[width=0.75\textwidth,angle=0]{./Figs/fig02.rev.png}\\
-    \caption{Multiplicative factor in the 2ICE IWC-relation as a function of the height relative to the freezing level.}\label{f2}
-\end{figure}
-
-For the determination of reference $a$ and $b$ values to be used with Equation (\ref{eq:1}), we assume that 
+For the determination of reference $a_{norm}$ and $b$ values to be used with Equation (\ref{eq:1}), we assume that 
 PSDs are normalized gamma  distributions with $N_w=0.08cm^{-4}$ and $\mu=2$ and calculate 
@@
 performance of a space-borne Ku-band radar system \citep{takahashi2008}.  Nevertheless, it should be noted that the
-uncertainties in the reflectivity calculations are likely to be even larger, given that the SSRGA theory although 
-quite accurate cannot possibly capture the entire range of uncertainties in the scattering properties of ice particles.
+uncertainties in the reflectivity calculations are likely to be even larger, given that the SSRGA theory, although 
+quite accurate, cannot possibly capture the entire range of uncertainties in the scattering properties of ice particles.
 Specifically, the SSRGA calculations were carried out assuming ice particles consisting of aggregates of 
@@
 provided in Table \ref{t_sensors}.  Errors in the radiometer observations are modeled assuming a 
-Noise-Equivalent-Delta-T (NEDT) of 1K, which is a readily achievable level for modern satellite radiometers \cite{draper2015}.
+Noise-Equivalent-Delta-T (NEDT) of 1K, which is a readily achievable level for modern satellite radiometers \citep{draper2015}.
 
@@
     \includegraphics[width=0.75\textwidth,angle=0]{./Figs/fig06.rev.png}\\
-    \caption{Same as in Figure 4, but for the IWC gravity center.}\label{f6}
+    \caption{Same as in Figure \ref{f4}, but for the IWC gravity center.}\label{f6}
 \end{figure}
@@
 radiometer-only observations, the lidar observations being subject to severe attenuation while the radar observations
-being limited by its sensitivity, they are nevertheless insightful. This is because quantifying the limitations of
+being limited by sensitivity, they are nevertheless insightful. This is because quantifying the limitations of
 retrievals from active instruments may be used to better assess  the benefits of synergistic retrievals.  Shown in

### 2023-05-22 RST/Submission/RSTManuscript.tex
Revised Manuscript

--- 2023-05-16 RST/RSTManuscript.tex
@@
     \centering
-    \includegraphics[width=0.75\textwidth,angle=0]{./Figs/fig02.rev.png}\\
+    \includegraphics[width=0.75\textwidth,angle=0]{fig02.rev.png}\\
     \caption{Multiplicative factor in the 2ICE IWC-relation as a function of the height relative to the freezing level.}\label{f2}
@@
     \centering
-    \includegraphics[width=0.75\textwidth,angle=0]{./Figs/fig01.png}\\
+    \includegraphics[width=0.75\textwidth,angle=0]{fig01.png}\\
     \caption{Mean CS IWC profiles for 18 classes derived using the k-Means clustering algorithm. For a 
@@
 uncertainties in the lidar observations may be accounted for to some extent by including a
-multiplicative random noise factor in the calculated lidar backscatter values.  Specifically, the calculated
-lidar backscatter values are multiplied with a log-normally distributed random variable with 0.0 mean and a
-standard deviation of 0.1.  This results in uncertainties of about 10\% in the lidar backscatter values, which is
-consistent the performance of  Cloud-Aerosol Transport System (CATS) lidar system reported in \cite{pauly2019}.  
-Nevertheless, the lidar simulations are rather idealized in that they do not include random errors due to daytime 
-solar background noise or systematic errors such as calibration accuracy or lidar ratio
-assumptions. More advanced models of the observation errors exist \citep{liu2006}, but are not 
+multiplicative random noise factor in the calculated lidar backscatter values.  
+Specifically, the calculated lidar backscatter values are multiplied with a log-normally distributed random variable 
+with 0.0 mean and a standard deviation of 0.1. This results in uncertainties of about 10\% in the lidar backscatter values, 
+conservative compared to the expected performance of the AOS lidars at 532 nm but consistent the performance of 
+Cloud-Aerosol Transport System (CATS) lidar system at 1064 nm reported in \citep{pauly2019}. While the AOS lidars 
+will have expected backscatter accuracies of 2-5\%, the lidar simulations in this paper are rather idealized in that 
+they do not include random errors due to daytime solar background noise or systematic errors such as calibration accuracy 
+or lidar ratio assumptions. More advanced models of the observation errors exist \citep{liu2006}, but are not 
 considered here. There are three main reasons this approach was taken: (1) it is important to
@@
     \centering
-    \includegraphics[width=0.75\textwidth,angle=0]{./Figs/fig03.rev.png}\\
+    \includegraphics[width=0.75\textwidth,angle=0]{fig03.rev.png}\\
     \caption{Simulated distributions of Ku-band radar reflectivity (left) and lidar backscatter (right)
@@
     \hline
-    Lidar & 524 nm & LogNormal(0,0.1) & 5e-7 m$^{-1}$sr$^{-1}$ \\
+    Lidar & 532 nm & LogNormal(0,0.1) & 5e-7 m$^{-1}$sr$^{-1}$ \\
     \hline
@@
     \centering
-    \includegraphics[width=0.75\textwidth,angle=0]{./Figs/fig04.rev.png}\\
+    \includegraphics[width=0.75\textwidth,angle=0]{fig04.rev.png}\\
     \caption{Frequency plot of estimated IWP derived radiometer-observations as a function of the 
@@
     \centering
-    \includegraphics[width=0.75\textwidth,angle=0]{./Figs/fig05.rev.png}\\
+    \includegraphics[width=0.75\textwidth,angle=0]{fig05.rev.png}\\
     \caption{True and radiometer-only retrieved conditional mean IWC for the 18 classes described in 
@@
     \centering
-    \includegraphics[width=0.75\textwidth,angle=0]{./Figs/fig06.rev.png}\\
+    \includegraphics[width=0.75\textwidth,angle=0]{fig06.rev.png}\\
     \caption{Same as in Figure \ref{f4}, but for the IWC gravity center.}\label{f6}
@@
     \centering
-    \includegraphics[width=.481\linewidth]{Figs/fig07a.rev.png}
-    \includegraphics[width=.481\linewidth]{Figs/fig07b.rev.png}
-    \includegraphics[width=.481\linewidth]{Figs/fig07c.rev.png}
-    \includegraphics[width=.481\linewidth]{Figs/fig07d.rev.png}
+    \includegraphics[width=.481\linewidth]{fig07a.rev.png}
+    \includegraphics[width=.481\linewidth]{fig07b.rev.png}
+    \includegraphics[width=.481\linewidth]{fig07c.rev.png}
+    \includegraphics[width=.481\linewidth]{fig07d.rev.png}
     \caption{Top Row: Density plots of IWP retrievals from lidar-only (left), and radar-only (right) observations as a function
@@
     \centering
-    \includegraphics[width=0.75\textwidth,angle=0]{./Figs/fig08.rev.png}\\
+    \includegraphics[width=0.75\textwidth,angle=0]{fig08.rev.png}\\
     \caption{Same as in Figure \ref{f4}, but with the lidar and radar observations incorporated in the
@@
     \centering
-    \includegraphics[width=0.75\textwidth,angle=0]{./Figs/fig09.rev.png}\\
+    \includegraphics[width=0.75\textwidth,angle=0]{fig09.rev.png}\\
     \caption{Same as in Figure \ref{f5}, but with the lidar and radar observations incorporated in the
@@
     \centering
-    \includegraphics[width=0.75\textwidth,angle=0]{./Figs/fig10.rev.png}\\
+    \includegraphics[width=0.75\textwidth,angle=0]{fig10.rev.png}\\
     \caption{Same as in Figure \ref{f6}, but with the lidar and radar observations incorporated in the
@@
     \centering
-    \includegraphics[width=.481\linewidth]{Figs/fig11a.rev.png}
-    \includegraphics[width=.481\linewidth]{Figs/fig11b.rev.png}
-    \includegraphics[width=.481\linewidth]{Figs/fig11c.rev.png}
-    \includegraphics[width=.481\linewidth]{Figs/fig11d.rev.png}
+    \includegraphics[width=.481\linewidth]{fig11a.rev.png}
+    \includegraphics[width=.481\linewidth]{fig11b.rev.png}
+    \includegraphics[width=.481\linewidth]{fig11c.rev.png}
+    \includegraphics[width=.481\linewidth]{fig11d.rev.png}
     \caption{Top Row: Density plots of IWP retrievals from lidar-radiometer (left), and radar-radiometer (right) observations as a function
@@
 such as The Investigation of Microphysics and Precipitation for Atlantic Coast-Threatening Snowstorms (IMPACTS)
-\citep{mcmurdie2022}.
+\citep{mcmurdie2022}. To achieve its objectives, which were driven by the need to improve the understanding of snowfall processes, 
+remote sensing of snow, and the prediction of banded snow structures \citep{mcmurdie2022}, IMPACTS relied on a suite 
+of active and passive instruments deployed via a satellite-simulating aircraft. These included multiple radars, one of them
+operating at Ku-band, a 532 nm elastic lidar, and a sub-millimeter-wave radiometer similar to the one considered in this study.
+While the objectives of IMPACTS were snowstorms, the instruments used in the campaign sampled a wide range of clouds, 
+including high ice clouds.  The IMPACTS observations associated with high ice clouds may be used to derive IWC estimates that may be
+directly validating using "in-situ" measurements, as the high-altitude aircraft flew in coordination with a cloud penetrating
+aircraft that carried cloud and ice particle probes. These data, although not fully available yet, are expected to
+provide valuable information on the accuracy of the IWC retrievals from synergistic lidar, Ku-band radar and 
+sub-millimeter-wave radiometer observations and enable the refinement of the retrieval algorithm formulated in this study.
+As such, they are considered a priority for future work.
 

### 2023-08-22 RST/Submission/responseReviewers_Rev2.tex
rev 2

\documentclass[12pt]{article}
\usepackage{amsmath,amssymb,amsthm}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{longtable}
\usepackage[margin=1in]{geometry}
\author{Mircea Grecu and John E. Yorks}
\title{Synergistic retrievals of ice in high clouds
Response to Reviewers}
\date{}
\begin{document}
\maketitle

\noindent \textbf{Reviewer 2}

\noindent \textbf{General Summary}

\noindent\textit{ 
For the sake of reproducibility, the authors should make all relevant source code publicly available. Furthermore, the data availability statement should clearly state which CloudSat products were used in the study.}\\

\noindent\textit{l. 23 - 24: Considering that the study neglects substantial uncertainties that will affect the accuracy of an actual synergistic retrieval, the sentence should mention that the 'significant improvements' were found under idealized conditions.}\\

\noindent Yes, we agree with the reviewer that the synergy was found under idealized conditions. We added a statement to this effect in the manuscript. Nevertheless, although additional work is likely needed to make optimal use of this synergy in practice, it is hard to envision how the tree instruments considered together cannot provide a more complete and quantitatively more accurate pictures than any one of them considered independently.\\

\noindent\textit{l. 85: You state that you use the 2C-ICE product to extend your CS-based estimates. However, you do not explain how.} \\

\noindent We apologize for the omission. When the CS-reflectivity is below the noise level, we use the IWC estimates in the 2C\-ICE product and $N_w$ derived by our statistical model. We added a sentence to the manuscript to clarify this point.\\

\noindent\textit{l. 113: Please introduce the variable H as relative height above freezing level.}\\

\noindent\textit{l. 125: I am confused by this paragraph. The preceding paragraph described how you parametrized $N_w$. However, $N_w$ still depends on $a_{norm}$, the calculation of which is described first in the following paragraph.}\\

\noindent We apologize for the confusion. No, $N_w$ does not depend on $a_{norm}$. $a_{norm}$ derived by regressing $log_{10}(IWC/N_w)$ against $log_{10}(Z/_Nw)$ with $Z$ in natural units. The slope of the regression is $b$ in Eq. (1), while $a_{norm}$ is $10^{intercept}$ where $intercept$ is the intercept of the regression.  This is explained in Testud et al. (2001) and several other papers using the concept of normalized PSDs. \\

\noindent\textit{l. 147: Here, you assume a fixed value for $N_w$. How is this consistent with the $N_w$ parametrization described earlier?}\\

\noindent We apologize for the confusion. As explained above, $IWC$ and $Z$ are normalized by $N_w$ in the derivation of $a_{norm}$. Consequently, the actual value of $N_w$ does not matter. One may set $N_w$ to the unit value. However, that would result in very unusual $IWC$ and $Z$, which would make sanity checks impossible.  We therefore set $N_w$ to a frequently used, but arbitrary values. This however does not have any impact on the actual value of $N_w$.\\

\noindent\textit{l. 258: How are absorption and extinction handled in your retrieval of IWC and PSD parameters from CS reflectivities?}\\

The gaseous absorption is modeled using the Rosenkranz (1993) model assuming the same water vapor, temperature and pressure profiles used in the lidar model and the radiative transfer calculations. The extinction is computed using the SSRGA theory, while the attenuation correction is handled through a forward attenuation correction procedure.
\\
\noindent\textit{l. 276 - 278: While generally true, this does not apply to this study since the test data stems from the same distribution as the a priori in either case.}\\

\noindent We just explain our preference for an observation-based dataset as opposed to a purely synthetic (CRM-based) dataset. However, we do not imply that our results are bias-free or that CRM can not eventually produce more complete and less-biased datasets. But until such datasets emerge, we consider the framework described .\\

\noindent\textit{l. 291 - 292: Are you using a two-layer neural network without activations functions, or is this a typo? In the former case, I suggest you use a deeper neural network with non-linearities for the classification as you model may be underfitting the data.}\\

\noindent We apologize for the confusion and omissions. The neural network has two hidden layers with ReLU activations. We considered additional layers and/or different numbers of neurons per layer, but the results did not improve. We added a sentence to the manuscript to clarify this point. \\

\noindent\textit{l. 347: The new distribution of retrieved gravity centers cannot be described as narrow.}

\noindent\textit{Section 3. b.: Please add results for the lidar-radar retrievals.}\\

\noindent\textit{Table 2: Please include NRMS also for IWP and gravity centers. Since you mention it in the text, I suggest to also include the correlation coefficient in the table.}\\

\noindent\textit{l. 440: The term cross-validation implies several iterations of evaluations. This is different from what you are doing here.}\\

\noindent We agree with the reviewer that our particular form of validation, which more generally known as hold-out validation, is not considered cross-validation by some authors. Nevertheless, other authors (e.g. Arlot 2010) consider hold-out the most basic form of cross-validation. To avoid controversy, 
we use term hold-out validation instead of cross-validation in the revised manuscript.\\

\begin{thebibliography}{999}

\bibitem{arlot2010survey}
Arlot, S., A survey of cross-validation procedures for model selection. Statistics Surveys. 2010, 4, 40-79.

\bibitem{ferreira2001}
Ferreira, F., P. Amayenc, S. Oury, and J. Testud, 2001: Study and Tests of Improved Rain Estimates from the TRMM Precipitation Radar. 
J. Appl. Meteor. Climatol., 40, 1878–1899, https://doi.org/10.

\end{thebibliography}

\end{document}

### 2023-08-22 RST/Submission/RSTManuscript_rev2.tex
rev 2

--- 2023-05-22 RST/Submission/RSTManuscript.tex
@@
 estimated class as a priori information.\\
-The results of the study show that the synergy of lidar, radar, and radiometer observations is significant
-in the retrieval of the IWC profiles.  The inclusion of the lidar backscatter observations in the retrieval process has a
-larger impact on the retrieval performance than the inclusion of the radar observations. 
-As ice clouds have a significant impact on atmospheric radiative processes, this
-work is relevant to ongoing efforts to reduce uncertainties in climate analyses and projections.}
+The results of the study show that the synergy of lidar, radar, and radiometer observations is significant in the retrieval of the IWC profiles.
+Nevertheless, it should be mentioned that this synergy was found under idealized conditions, and additional work might be required to materialize it in practice.
+The inclusion of the lidar backscatter observations in the retrieval process has a larger impact on the retrieval performance than the inclusion of the radar observations.
+As ice clouds have a significant impact on atmospheric radiative processes, this work is relevant to ongoing efforts to reduce uncertainties in climate analyses and projections.}
 
@@
 \section{Introduction}
-The future NASA Atmospheric Observing System (AOS) mission \citep{braun2022} is expected to feature new 
-combinations of observations that may be used to quantify the amounts of ice in high clouds and 
-characterize the microphysical properties of ice particles. In the AOS terminology, high clouds are convectively
-generated clouds \citep{braun2022}. They are the result of strong vertical mass transport accompanied by
-horizontal transport of hydrometeors in the upper troposphere.  As high clouds are of paramount importance in
-understanding the quantifying radiative processes, they constitute one of the AOS major objectives \citep{braun2022}.
-To achieve its objectives, AOS will rely on a combination of active and passive observations onboard multiple spacecrafts
-in two different orbits.  One of the orbits is expected to be polar, while the other is expected to be inclined, which
-would allow the study of atmospheric processes at sub-daily time scales, with an emphasis on deep
-convection, high clouds, and aerosols \citep{braun2022,yorks2022}.
-The AOS inclined observations will include backscatter
-from an elastic backscatter lidar \citep{weitkamp2006},
-Ku-band radar reflectivity, and submillimeter wave radiometer brightness temperature measurements. 
-While not necessarily optimal for cloud ice estimation, these measurements are 
-complimentary and enable the synergistic characterization of ice clouds. 
-That is, despite the fact that lidar observations 
-attenuate quickly in thick ice clouds and the Ku-band radar will not be able to detect clouds 
+The future NASA Atmospheric Observing System (AOS) mission \citep{braun2022} is expected to feature new combinations of observations that may be used to quantify the amounts of ice in high clouds and characterize the microphysical properties of ice particles.
+In the AOS terminology, high clouds are convectively generated clouds \citep{braun2022}.
+They are the result of strong vertical mass transport accompanied by
+horizontal transport of hydrometeors in the upper troposphere.
+As high clouds are of paramount importance in understanding the quantifying radiative processes, they constitute one of the AOS major objectives \citep{braun2022}.
+To achieve its objectives, AOS will rely on a combination of active and passive observations onboard multiple spacecrafts in two different orbits.
+One of the orbits is expected to be polar, while the other is expected to be inclined, which would allow the study of atmospheric processes at sub-daily time scales, with an emphasis on deep convection, high clouds, and aerosols \citep{braun2022,yorks2022}.
+The AOS inclined observations will include backscatter from an elastic backscatter lidar \citep{weitkamp2006}, Ku-band radar reflectivity, and submillimeter wave radiometer brightness temperature measurements.
+While not necessarily optimal for cloud ice estimation, these measurements are complimentary and enable the synergistic characterization of ice clouds.
+That is, despite the fact that lidar observations attenuate quickly in thick ice clouds and the Ku-band radar will not be able to detect clouds
 characterized by an echo weaker than 8.0 dBZ (which is the expected sensitivity of the radar in the inclined
 orbit), the active observations are expected to provide context 
-that may be incorporated into the radiometer retrievals. Herein, the term retrieval is defined as the 
-process of estimating geophysical variables from remote sensing observations.
+that may be incorporated into the radiometer retrievals. Herein, the term retrieval is defined as the process of estimating geophysical variables from remote sensing observations.
 %, but, as customary in the scientific
@@
 incorporating lidar and radar observations into the radiometer retrieval of ice clouds.
-Because the existing amount of coincident backscatter lidar, Ku-band radar, and submillimeter-wave 
-radiometer observations is rather insufficient to derive conclusive results, we employ accurate 
-physical models to simulate lidar, radar and radiometer observations and use a cross-validation 
-methodology to characterize the retrieval accuracy. As estimates from passive instrument 
-observations strongly depend on "a priori" information \citep{rodgers2000inverse}, for results to be 
-relevant in real applications, it is necessary to base them on realistic vertical distributions of 
-ice properties.  Such distributions may be derived from cloud-resolving-model (CRM) simulations 
-\citep{pfreundschuh2020synergistic,liu2022assessing} or directly from observations.  In this study, we employ the latter 
-approach, as CRMs may still be deficient in properly reproducing the vertical distribution of ice 
-clouds and their associated microphysical properties.  Specifically, we use observations and
-products from the CloudSat (CS) mission (Stephens et al. 2002) to derive a database of ice microphysical
-properties and associated simulated lidar, radar, and radiometer observations.  To account for variability in 
-the Particle Size Distributions (PSD) that may not be well represented in the 2C-ICE product, we use a simple
-but effective approach to perturb the PSD generalized intercepts from their nominal values.
-The resulting database is used to
-investigate the accuracy of estimated ice cloud properties from the simulated observations. Another major
-difference relative to previous studies is the unique combination of instruments investigated herein. It should be
-mentioned that although based on observations rather than CRM simulations, the approach used in this study is
+Because the existing amount of coincident backscatter lidar, Ku-band radar, and submillimeter-wave  radiometer observations is rather insufficient to derive conclusive results, we employ accurate
+physical models to simulate lidar, radar and radiometer observations and use a cross-validation methodology to characterize the retrieval accuracy.
+As estimates from passive instrument observations strongly depend on "a priori" information \citep{rodgers2000inverse}, for results to be
+relevant in real applications, it is necessary to base them on realistic vertical distributions of ice properties.
+Such distributions may be derived from cloud-resolving-model (CRM) simulations \citep{pfreundschuh2020synergistic,liu2022assessing} or directly from observations.
+In this study, we employ the latter approach, as CRMs may still be deficient in properly reproducing the vertical distribution of ice clouds and their associated microphysical properties.
+Specifically, we use observations and products from the CloudSat (CS) mission (Stephens et al.
+2002) to derive a database of ice microphysical properties and associated simulated lidar, radar, and radiometer observations.
+To account for variability in the Particle Size Distributions (PSD) that may not be well represented in the 2C-ICE product, we use a simple but effective approach to perturb the PSD generalized intercepts from their nominal values.
+The resulting database is used to investigate the accuracy of estimated ice cloud properties from the simulated observations.
+Another major difference relative to previous studies is the unique combination of instruments investigated herein.
+It should be mentioned that although based on observations rather than CRM simulations, the approach used in this study is
 idealized in many respects. As a consequence, the results presented may not unbiasedly extrapolate to practice.
@@
 good first step towards the development of an operational synergistic retrieval of ice clouds from AOS inclined observations.
-The article is organized as follows.  In Section 2, we describe the approach used to derive the ice properties and
-the associated simulated observations, the retrieval and the evaluation methodology.  In Section 3, we 
-present the results of the evaluation methodology. We conclude in Section 4.
+The article is organized as follows.
+In Section 2, we describe the approach used to derive the ice properties and the associated simulated observations, the retrieval and the evaluation methodology.
+In Section 3, we present the results of the evaluation methodology.
+We conclude in Section 4.
 
@@
     & 325.15+/-3.5 GHz & 1.0 K & N/A \\
-    & 183.31+/-9.5 GHz & 1.0 K & N/A \\
+    & 325.15+/-9.5 GHz & 1.0 K & N/A \\
    \hline \\

### 2023-08-22 RST/Submission/responseReviewers_Rev2.tex
revision 2

@@
 \author{Mircea Grecu and John E. Yorks}
-\title{Synergistic retrievals of ice in high clouds
-Response to Reviewers}
+\title{Synergistic retrievals of ice in high clouds\\
+Response to Reviewers\\Second Revision}
 \date{}
@@
 
-\noindent \textbf{Reviewer 2}
-
+\noindent \textbf{Reviewer 2}\\
 \noindent \textbf{General Summary}
@@
 We added a statement to this effect in the manuscript.
-Nevertheless, although additional work is likely needed to make optimal use of this synergy in practice, it is hard to envision how the tree instruments considered together cannot provide a more complete and quantitatively more accurate pictures than any one of them considered independently.\\
+Nevertheless, although additional work is likely needed to make optimal use of this synergy in practice, it is hard to envision how the tree instruments considered together cannot provide a more complete and quantitatively more accurate picture than any one of them considered independently.\\
 
@@
 \noindent We apologize for the omission.
-When the CS-reflectivity is below the noise level, we use the IWC estimates in the 2C\-ICE product and $N_w$ derived by our statistical model.
+When the CS-reflectivity is below the noise level, we use the IWC estimates in the 2C\-ICE product in conjunction with an $N_w$ provided by our statistical model.
 We added a sentence to the manuscript to clarify this point.\\
@@
 \noindent\textit{l. 113: Please introduce the variable H as relative height above freezing level.}\\
+
+Thank you for the suggestion. We introduced $H$ as the relative height above freezing level as suggested.\\
 
@@
 
-The gaseous absorption is modeled using the Rosenkranz (1993) model assuming the same water vapor, temperature and pressure profiles used in the lidar model and the radiative transfer calculations.
+\noindent The gaseous absorption is modeled using the Rosenkranz (1998) model assuming the same water vapor, temperature and pressure profiles used in the lidar model and the radiative transfer calculations.
 The extinction is computed using the SSRGA theory, while the attenuation correction is handled through a forward attenuation correction procedure.
@@
 However, we do not imply that our results are bias-free or that CRM can not eventually produce more complete and less-biased datasets.
-But until such datasets emerge, we consider the framework described .\\
+But until such datasets emerge, we consider the framework described.\\
 
@@
 
-\noindent\textit{l. 347: The new distribution of retrieved gravity centers cannot be described as narrow.}
+\noindent\textit{l. 347: The new distribution of retrieved gravity centers cannot be described as narrow.}\\
+
+\noindent We agree with the reviewer that the range of the retrieved gravity centers is not narrow.
+However, a significant fraction of retrieved IWC GC are within narrower range spanning from about 2.5 km above the freezing level to about 5.5 km above the freezing level.
+We added a sentence to the manuscript to clarify this point.\\
 
 \noindent\textit{Section 3. b.: Please add results for the lidar-radar retrievals.}\\
+
+\noindent Thank you for the suggestion. We added the results for the lidar-radar retrievals in the revised manuscript.\\
 
@@
 
-\noindent We agree with the reviewer that our particular form of validation, which more generally known as hold-out validation, is not considered cross-validation by some authors.
+\noindent We agree with the reviewer that our particular form of validation, which is more generally known as hold-out validation, is not considered cross-validation by some authors.
 Nevertheless, other authors (e.g.

### 2023-08-22 RST/Submission/RSTManuscript_rev2.tex
revision 2

@@
 
-\abstract{ In this study, we investigate the synergy of elastic backscatter lidar, Ku-band radar, and sub-millimeter-wave radiometer measurements
-in the retrieval of ice from satellite observations. The synergy is analyzed through the generation of a large dataset 
-of Ice Water Content (IWC) profiles and simulated lidar, radar and radiometer observations. 
-The characteristics of the instruments e.g. frequencies, sensitivities, etc. are set based on the expected characteristics 
-of instruments of the Atmosphere Observing System (AOS) mission. A cross-validation methodology is used to assess the
-accuracy of the IWC profiles retrieved from various combinations of observations from the three instruments. Specifically, the IWC and
-associated observations are randomly divided into two datasets, one for training and the other for evaluation.
-The training dataset is used to train the retrieval algorithm, while the evaluation dataset is used to assess the 
-retrieval performance. The dataset of IWC profiles is derived from CloudSat 
-reflectivity and CALIOP lidar observations.
+\abstract{ In this study, we investigate the synergy of elastic backscatter lidar, Ku-band radar, and sub-millimeter-wave radiometer measurements in the retrieval of ice from satellite observations.
+The synergy is analyzed through the generation of a large dataset of Ice Water Content (IWC) profiles and simulated lidar, radar and radiometer observations.
+The characteristics of the instruments e.g.
+frequencies, sensitivities, etc.
+are set based on the expected characteristics of instruments of the Atmosphere Observing System (AOS) mission.
+A hold-out validation methodology is used to assess the accuracy of the IWC profiles retrieved from various combinations of observations from the three instruments.
+Specifically, the IWC and associated observations are randomly divided into two datasets, one for training and the other for evaluation.
+The training dataset is used to train the retrieval algorithm, while the evaluation dataset is used to assess the retrieval performance.
+The dataset of IWC profiles is derived from CloudSat reflectivity and CALIOP lidar observations.
 The retrieval of the ice water content IWC profiles from the computed observations is achieved in two steps.
-In the first step, a class, out of 18 potential classes characterized by different vertical distribution of IWC, 
-is estimated from the observations. The 18 classes are predetermined based on the k-Means clustering algorithm. 
-In the second step, the IWC profile is estimated using an Ensemble Kalman Smoother (EKS) algorithm that uses the 
-estimated class as a priori information.\\
+In the first step, a class, out of 18 potential classes characterized by different vertical distribution of IWC, is estimated from the observations.
+The 18 classes are predetermined based on the k-Means clustering algorithm.
+In the second step, the IWC profile is estimated using an Ensemble Kalman Smoother (EKS) algorithm that uses the estimated class as a priori information.\\
 The results of the study show that the synergy of lidar, radar, and radiometer observations is significant in the retrieval of the IWC profiles.
@@
 In the AOS terminology, high clouds are convectively generated clouds \citep{braun2022}.
-They are the result of strong vertical mass transport accompanied by
-horizontal transport of hydrometeors in the upper troposphere.
+They are the result of strong vertical mass transport accompanied by horizontal transport of hydrometeors in the upper troposphere.
 As high clouds are of paramount importance in understanding the quantifying radiative processes, they constitute one of the AOS major objectives \citep{braun2022}.
@@
 While not necessarily optimal for cloud ice estimation, these measurements are complimentary and enable the synergistic characterization of ice clouds.
-That is, despite the fact that lidar observations attenuate quickly in thick ice clouds and the Ku-band radar will not be able to detect clouds
-characterized by an echo weaker than 8.0 dBZ (which is the expected sensitivity of the radar in the inclined
-orbit), the active observations are expected to provide context 
-that may be incorporated into the radiometer retrievals. Herein, the term retrieval is defined as the process of estimating geophysical variables from remote sensing observations.
+That is, despite the fact that lidar observations attenuate quickly in thick ice clouds and the Ku-band radar will not be able to detect clouds characterized by an echo weaker than 8.0 dBZ (which is the expected sensitivity of the radar in the inclined orbit), the active observations are expected to provide context that may be incorporated into the radiometer retrievals.
+Herein, the term retrieval is defined as the process of estimating geophysical variables from remote sensing observations.
 %, but, as customary in the scientific
 %literature, estimates are occasionally referred to as retrievals as well.  
-In this study, we investigate the impact of 
-incorporating lidar and radar observations into the radiometer retrieval of ice clouds.
-Because the existing amount of coincident backscatter lidar, Ku-band radar, and submillimeter-wave  radiometer observations is rather insufficient to derive conclusive results, we employ accurate
-physical models to simulate lidar, radar and radiometer observations and use a cross-validation methodology to characterize the retrieval accuracy.
-As estimates from passive instrument observations strongly depend on "a priori" information \citep{rodgers2000inverse}, for results to be
-relevant in real applications, it is necessary to base them on realistic vertical distributions of ice properties.
+In this study, we investigate the impact of incorporating lidar and radar observations into the radiometer retrieval of ice clouds.
+Because the existing amount of coincident backscatter lidar, Ku-band radar, and submillimeter-wave  radiometer observations is rather insufficient to derive conclusive results, we employ accurate physical models to simulate lidar, radar and radiometer observations and use a hold-out validation methodology to characterize the retrieval accuracy.
+As estimates from passive instrument observations strongly depend on "a priori" information \citep{rodgers2000inverse}, for results to be relevant in real applications, it is necessary to base them on realistic vertical distributions of ice properties.
 Such distributions may be derived from cloud-resolving-model (CRM) simulations \citep{pfreundschuh2020synergistic,liu2022assessing} or directly from observations.
@@
 Another major difference relative to previous studies is the unique combination of instruments investigated herein.
-It should be mentioned that although based on observations rather than CRM simulations, the approach used in this study is
-idealized in many respects. As a consequence, the results presented may not unbiasedly extrapolate to practice.
-Nonetheless, the results are expected to provide useful insights into the potential of the AOS inclined observations and
-good first step towards the development of an operational synergistic retrieval of ice clouds from AOS inclined observations.
+It should be mentioned that although based on observations rather than CRM simulations, the approach used in this study is idealized in many respects.
+As a consequence, the results presented may not unbiasedly extrapolate to practice.
+Nonetheless, the results are expected to provide useful insights into the potential of the AOS inclined observations and good first step towards the development of an operational synergistic retrieval of ice clouds from AOS inclined observations.
 The article is organized as follows.
@@
 \section{Methodology}
-As previously mentioned, we use CloudSat (CS) observations \citep{stephens2002} to derive the vertical 
-distributions of ice properties needed in the investigation.  Although research quality CS cloud ice 
-products exist, to maximize the physical consistency of the approach, we do not use them but derive ice amounts
-and associated properties directly from CS reflectivity observations.  This ensures the consistency between 
-the particle
-distribution assumptions and the electromagnetic scattering properties used in the CS reflectivity processing 
-and those used the simulation of the lidar, Ku-band radar and radiometer observations.  
-Our CS-based ice product is optimized to be consistent with the synergistic 
-Cloudsat and CALIPSO Ice Cloud Property Product (2C-ICE) of \cite{deng2015}.  When the CALIPSO lidar detects echo
-associated with clouds but the CS radar signal is below the noise level, we use the 2C-ICE product to extend our 
-CS-based estimates. Lidar, Ku-band radar, and submillimeter-wave radiometer observations are simulated 
-from CS observations using accurate physical 
-models and realistic assumptions consistent with the most recent knowledge in the field of ice cloud 
-microphysics, and a non-parametric estimation methodology based on the k-Means clustering algorithm 
-\cite{mackay2003information} is used to 
-investigate the instrument synergy.  Details of the methodology are presented below.
+As previously mentioned, we use CloudSat (CS) observations \citep{stephens2002} to derive the vertical distributions of ice properties needed in the investigation.
+Although research quality CS cloud ice products exist, to maximize the physical consistency of the approach, we do not use them but derive ice amounts and associated properties directly from CS reflectivity observations.
+This ensures the consistency between the particle distribution assumptions and the electromagnetic scattering properties used in the CS reflectivity processing and those used the simulation of the lidar, Ku-band radar and radiometer observations.
+Our CS-based ice product is optimized to be consistent with the synergistic Cloudsat and CALIPSO Ice Cloud Property Product (2C-ICE) of \cite{deng2015}.
+When the CALIPSO lidar detects echo associated with clouds but the CS radar signal is below the noise level, we use the 2C-ICE product to extend our CS-based estimates.
+Specifically, we use the IWC estimates from the 2C-ICE product in conjunction with a Particle Size Distribution (PSD) generalized intercept \citep{testud2001} provided by statistical model described in the next section to characterize the ice PSDs of clouds with radar echo below the noise level of the CloudSat radar.
+Lidar, Ku-band radar, and submillimeter-wave radiometer observations are simulated from CS observations using accurate physical models and realistic assumptions consistent with the most recent knowledge in the field of ice cloud microphysics, and a non-parametric estimation methodology based on the k-Means clustering algorithm \cite{mackay2003information} is used to investigate the instrument synergy.
+Details of the methodology are presented below.
 
 \subsection{Assumptions and forward models}
-To quantify the number of ice particles in an elementary atmospheric volume as a function of their size, we use
-normalized gamma functions \citep{bringi2003}.  The benefit of normalized gamma functions is that they 
-encapsulate the variability of Ice Water Content (IWC) - reflectivity relationship into a single parameter,
-i.e. the 
-normalized Particle Size Distribution (PSD) intercept \citep{testud2001,bringi2003}. The normalized
-PSD intercept is defined as $N_w=\frac {4^4} {\pi \rho_w} \frac {IWC} {D_m^4}$, where $IWC$ is the ice water 
-content associated with the PSD, and $D_m$ is the mass weighted mean diameter.  Based on the work of \cite{testud2001},
-\cite{ferreira2001} and \cite{delanoe2014} showed 
-that the variability in IWC reflectivity (Z) relationships may be fully explained by variability in $N_w$, and 
-that a formula of the type
+To quantify the number of ice particles in an elementary atmospheric volume as a function of their size, we use normalized gamma functions \citep{bringi2003}.
+The benefit of normalized gamma functions is that they encapsulate the variability of Ice Water Content (IWC) - reflectivity relationship into a single parameter, i.e.
+the normalized Particle Size Distribution (PSD) intercept \citep{testud2001,bringi2003}.
+The normalized PSD intercept is defined as $N_w=\frac {4^4} {\pi \rho_w} \frac {IWC} {D_m^4}$, where $IWC$ is the ice water content associated with the PSD, and $D_m$ is the mass weighted mean diameter.
+Based on the work of \cite{testud2001}, \cite{ferreira2001} and \cite{delanoe2014} showed that the variability in IWC reflectivity (Z) relationships may be fully explained by variability in $N_w$, and that a formula of the type
 \begin{equation}
@@
 \end{equation}
-(where $a_{norm}$ and $b$ are constants) explains almost perfectly the relationships between IWC and Z calculated from observed PSDs.
-However, equation (\ref{eq:1}) is not 
-sufficient to derive accurate, unbiased estimates of ice water contents, because $N_w$ varies considerably in 
-time and space. Nevertheless, multiple studies showed that it is beneficial to parameterize $N_w$ as a function 
-of various variables, such as temperature \citep{hogan2006retrieval, delanoe2008variational,deng2010tropical}, 
-rather than using $N_w$ independent reflectivity ice relations.  In this study, we parameterize
-$N_w$ as a function of temperature based on the CloudSat 2C-ICE product \citep{deng2010tropical,deng2013evaluation}.  
-A scatter-plot analysis of relationships between the 2C-ICE IWC and the associated reflectivity suggests that
-the multiplicative coefficient in a power low ice-reflectivity relationship is parameterized as a function of temperature 
-(or equivalently, as a function of height relative to the freezing level) in 
-the default 2C-ICE retrievals. The default multiplicative coefficient (the value the provides the IWC estimate prior
-to the ingestion of the lidar observations) may be simply estimated by regressing the 2C-ICE IWC on the associated CloudSat reflectivity 
-as a function of height relative the freezing level. Specifically, a regression of the type $ln(IWC)=ln(a)+b*ln(Z)$ is performed for 
-each bin referenced relative to the zero degree bin is performed, and the variation of $ln(a)$ is parameterized as a function of the
-relative height.  The results are shown in Figure \ref{f2}. As apparent in the figure, and as expected, $a$ exhibits a strong variation 
-with the relative height. Given that any deviation of the multiplicative coefficient in 
-an IWC-Z relation from an average is equivalent to a deviation of the associated $N_w$ from its mean value 
-\citep{ferreira2001,delanoe2014}, the variation of $a$ as a function of relative-height may be converted into an $N_w$ 
-as a function of relative-height relationship.  We, therefore, use the data in Figure \ref{f2} to parameterize $N_w$ as a 
-function of the relative height.  Specifically, given that $(1-b)ln(N_w)+ln(a_{norm})=ln(a)$ and $ln(a)=ln(a_0)+s*H$, with 
-$a_0$ and $s$ the intercept and the slope of a regression for the data in Figure \ref{f2}, we 
-can express $N_w$ as a function of the relative height, slope $s$, and parameters $a_0$ and $a_{norm}$.
+(where $a_{norm}$ and $b$ are constants) explains almost perfectly the relationships between IWC and Z calculated from observed PSDs.However, equation (\ref{eq:1}) is not  sufficient to derive accurate, unbiased estimates of ice water contents, because $N_w$ varies considerably in time and space.
+Nevertheless, multiple studies showed that it is beneficial to parameterize $N_w$ as a function  of various variables, such as temperature \citep{hogan2006retrieval, delanoe2008variational,deng2010tropical}, rather than using $N_w$ independent reflectivity ice relations.
+In this study, we parameterize $N_w$ as a function of temperature based on the CloudSat 2C-ICE product \citep{deng2010tropical,deng2013evaluation}.
+A scatter-plot analysis of relationships between the 2C-ICE IWC and the associated reflectivity suggests that the multiplicative coefficient in a power low ice-reflectivity relationship is parameterized as a function of temperature (or equivalently, as a function of height relative to the freezing level) in the default 2C-ICE retrievals.
+The default multiplicative coefficient (the value the provides the IWC estimate prior to the ingestion of the lidar observations) may be simply estimated by regressing the 2C-ICE IWC on the associated CloudSat reflectivity as a function of $H$, defined as height relative the freezing level.
+Specifically, a regression of the type $ln(IWC)=ln(a)+b*ln(Z)$ is performed for each bin referenced relative to the zero degree bin is performed, and the variation of $ln(a)$ is parameterized as a function of the relative height.
+The results are shown in Figure \ref{f2}.
+As apparent in the figure, and as expected, $a$ exhibits a strong variation with the relative height.
+Given that any deviation of the multiplicative coefficient in an IWC-Z relation from an average is equivalent to a deviation of the associated $N_w$ from its mean value \citep{ferreira2001,delanoe2014}, the variation of $a$ as a function of relative-height may be converted into an $N_w$ as a function of relative-height relationship.
+We, therefore, use the data in Figure \ref{f2} to parameterize $N_w$ as a function of the relative height.
+Specifically, given that $(1-b)ln(N_w)+ln(a_{norm})=ln(a)$ and $ln(a)=ln(a_0)+s*H$, with $a_0$ and $s$ the intercept and the slope of a regression for the data in Figure \ref{f2}, we can express $N_w$ as a function of the relative height, slope $s$, and parameters $a_0$ and $a_{norm}$.
 
@@
 To investigate the variability in the vertical distribution of the 2C-ICE estimates and their consistency with our estimates, 
-we cluster, based on similarity  (quantified through the evaluation of the Euclidian distance between profiles), 
-a large set 2C-ICE profiles into 18 classes using a k-Means 
-procedure. The mean IWC profiles associated with the 18 classes are shown in continuous lines in Figure \ref{f1}.  
-Our estimates, derived using PSD assumptions and electromagnetic scattering calculations that enable 
-accurate and physically consistent simulations of radar observations at Ku-band and radiometer observations of 
-submillimeter-wave frequencies are also shown in the figure. These estimates are based on the self-similarity 
-Rayleigh-Gans approximation (SSRGA) of \cite{hogan2017ssrga}. Details regarding the
-estimation process are provided in the following paragraphs.  As apparent in Figure \ref{f1}, the CS and SSRGA estimates are in good 
-agreement.  Some discrepancies due to differences between the SSRGA $N_w$ parameterization and the CS 2C-ICE 
-"a priori assumptions" are also apparent, but they are not deemed critical in this study, whose objective is 
-the investigation of synergistic lidar, Ku-band radar and submillimeter-wave radiometer retrievals, because the 
-outcome is not likely to be sensitive to such details. Also apparent in Figure \ref{f1} is the fact that there is significant
-variability in the vertical distribution of the 2C-ICE IWC estimates, which makes the estimation of IWC profiles from passive-only 
-observations challenging.
+we cluster, based on similarity  (quantified through the evaluation of the Euclidian distance between profiles), a large set 2C-ICE profiles into 18 classes using a k-Means procedure.
+The mean IWC profiles associated with the 18 classes are shown in continuous lines in Figure \ref{f1}.
+Our estimates, derived using PSD assumptions and electromagnetic scattering calculations that enable accurate and physically consistent simulations of radar observations at Ku-band and radiometer observations of submillimeter-wave frequencies are also shown in the figure.
+These estimates are based on the self-similarity Rayleigh-Gans approximation (SSRGA) of \cite{hogan2017ssrga}.
+Details regarding the estimation process are provided in the following paragraphs.
+As apparent in Figure \ref{f1}, the CS and SSRGA estimates are in good agreement.
+Some discrepancies due to differences between the SSRGA $N_w$ parameterization and the CS 2C-ICE "a priori assumptions" are also apparent, but they are not deemed critical in this study, whose objective is the investigation of synergistic lidar, Ku-band radar and submillimeter-wave radiometer retrievals, because the outcome is not likely to be sensitive to such details.
+Also apparent in Figure \ref{f1} is the fact that there is significant variability in the vertical distribution of the 2C-ICE IWC estimates, which makes the estimation of IWC profiles from passive-only observations challenging.
 
@@
 
-For the determination of reference $a_{norm}$ and $b$ values to be used with Equation (\ref{eq:1}), we assume that 
-PSDs are normalized gamma  distributions with $N_w=0.08cm^{-4}$ and $\mu=2$ and calculate 
+For the determination of reference $a_{norm}$ and $b$ values to be used with Equation (\ref{eq:1}), we assume that PSDs are normalized gamma  distributions with $N_w=0.08cm^{-4}$ and $\mu=2$ and calculate
 \begin{equation}
@@
 \end{equation}
-where $\lambda$ is the radar frequency, $|K_w|$ is the dielectric factor of water, $N(D,D_m) dD$ is the number 
-of ice particles of diameter with D and D+dD per unit volume, $D_m$ is the mass weighted mean diameter of the 
-distribution, and $\sigma _b(D)$ is the backscattering cross-section of ice particle of diameter $D$. The mass 
-weighted mean diameter is equidistantly sampled to span the entire range of IWC values in the CS 2C-ICE dataset.
-The assumed mass-size relation is that of \cite{brown1995improved} because it works well with the 
-SSRGA scattering calculations \citep{heymsfield2022impacts}.  The open source software scatter-1.1 of \citep{scatter-1.1}
-is used to provide the actual scattering properties.   
-To improve the representation of microphysical variability in the study, we do not assume the values of $N_w$ given
-by the 2C-ICE based parametrization described above as the true values. Instead, we perturb them by multiplication 
-with a log-normally distributed random variable with 0.0 mean a standard deviation of 0.5.  The 
-perturbations are vertically correlated. Specifically, normal random variables with zero mean and a standard deviation
-of 1.0 are generated and then passed through a Gaussian smoothing filter \citep{nixon2019} with a size of two radar bins. The perturbation
-vector is then rescaled to have a standard deviation of 0.5.  The perturbations are then exponentiated and multiplied
-with the 2C-ICE based $N_w$ values to produce the final $N_w$ values.  The filter size and noise magnitude are chosen
-to roughly mimic the observed variability in $N_w$ derived from in situ observations of PSDs, such as those described
-in \cite{heymsfield2022impacts}. In addition, we are adding a random noise of
-0.0 mean and 0.5 dB standard deviation to the calculated reflectivity values, which is consistent with the expected
-performance of a space-borne Ku-band radar system \citep{takahashi2008}.  Nevertheless, it should be noted that the
-uncertainties in the reflectivity calculations are likely to be even larger, given that the SSRGA theory, although 
-quite accurate, cannot possibly capture the entire range of uncertainties in the scattering properties of ice particles.
-Specifically, the SSRGA calculations were carried out assuming ice particles consisting of aggregates of 
-bullet rosettes, columnar crystals, and plates.
+where $\lambda$ is the radar frequency, $|K_w|$ is the dielectric factor of water, $N(D,D_m) dD$ is the number of ice particles of diameter with D and D+dD per unit volume, $D_m$ is the mass weighted mean diameter of the distribution, and $\sigma _b(D)$ is the backscattering cross-section of ice particle of diameter $D$.
+The mass weighted mean diameter is equidistantly sampled to span the entire range of IWC values in the CS 2C-ICE dataset.
+The assumed mass-size relation is that of \cite{brown1995improved} because it works well with the SSRGA scattering calculations \citep{heymsfield2022impacts}.
+The open source software scatter-1.1 of \citep{scatter-1.1} is used to provide the actual scattering properties.
+To improve the representation of microphysical variability in the study, we do not assume the values of $N_w$ given by the 2C-ICE based parametrization described above as the true values.
+Instead, we perturb them by multiplication with a log-normally distributed random variable with 0.0 mean a standard deviation of 0.5.
+The perturbations are vertically correlated.
+Specifically, normal random variables with zero mean and a standard deviation of 1.0 are generated and then passed through a Gaussian smoothing filter \citep{nixon2019} with a size of two radar bins.
+The perturbation vector is then rescaled to have a standard deviation of 0.5.
+The perturbations are then exponentiated and multiplied with the 2C-ICE based $N_w$ values to produce the final $N_w$ values.
+The filter size and noise magnitude are chosen to roughly mimic the observed variability in $N_w$ derived from in situ observations of PSDs, such as those described in \cite{heymsfield2022impacts}.
+In addition, we are adding a random noise of 0.0 mean and 0.5 dB standard deviation to the calculated reflectivity values, which is consistent with the expected performance of a space-borne Ku-band radar system \citep{takahashi2008}.
+Nevertheless, it should be noted that the uncertainties in the reflectivity calculations are likely to be even larger, given that the SSRGA theory, although quite accurate, cannot possibly capture the entire range of uncertainties in the scattering properties of ice particles.
+Specifically, the SSRGA calculations were carried out assuming ice particles consisting of aggregates of bullet rosettes, columnar crystals, and plates.
 
-Moreover, the SSRGA theory was developed for millimeter and 
-submillimeter-wave calculations and may not be applicable at lidar's wavelength. 
-Therefore, lidar observations are computed using the Mie solution included in the scatter-1.1 
-package for a backscatter lidar, given that such a lidar is expected to be flown with a 
-Ku-band radar and microwave radiometer as part of the NASA Earth System Observatory 
-(ESO) Atmosphere Observing System (AOS) in an inclined and/or polar orbit \citep{yorks2022}.
-While uncertainties in the lidar forward model are rather complex and difficult to quantify,
-uncertainties in the lidar observations may be accounted for to some extent by including a
-multiplicative random noise factor in the calculated lidar backscatter values.  
-Specifically, the calculated lidar backscatter values are multiplied with a log-normally distributed random variable 
-with 0.0 mean and a standard deviation of 0.1. This results in uncertainties of about 10\% in the lidar backscatter values, 
-conservative compared to the expected performance of the AOS lidars at 532 nm but consistent the performance of 
-Cloud-Aerosol Transport System (CATS) lidar system at 1064 nm reported in \citep{pauly2019}. While the AOS lidars 
-will have expected backscatter accuracies of 2-5\%, the lidar simulations in this paper are rather idealized in that 
-they do not include random errors due to daytime solar background noise or systematic errors such as calibration accuracy 
-or lidar ratio assumptions. More advanced models of the observation errors exist \citep{liu2006}, but are not 
-considered here. There are three main reasons this approach was taken: (1) it is important to
-understand the limitations of combining these three datasets before factoring in 
-individual sensor limitations, (2) while more accurate calculations based on more 
-realistic ice particle shapes exist, they are rather incomplete and not readily available,
-and (3) Wagner and Delene (2022) compared lidar backscatter observations with backscatter 
-calculations based on coincident PSD observations and the Mie solution and found good 
-agreement, which suggests that electromagnetic properties derived from Mie calculations 
-are adequate for practical applications. 
-The lidar molecular backscatter and extinction
-are calculated using the lidar module of the CFMIP Observation Simulator Package (COSP; 
-Bodas-Salcedo et al. (2011)). COSP simulates the lidar total attenuated backscatter 
-signal and scattering ratios at 532 nm for scenes with and without clouds. COSP also 
-assumes cloud particles are spherical so that the backscattering phase function is 
-estimated based on the effective radius (Mie theory). Despite this assumption, the 
-COSP simulations agree well with CALIOP observations. To account for multiple-scattering 
-in the lidar observations, we are using the multiscatter-1.2.11 model \citep{multiscatter-1.2.11} of 
-\cite{hogan2008lidar_ms}.
+Moreover, the SSRGA theory was developed for millimeter and submillimeter-wave calculations and may not be applicable at lidar's wavelength.
+Therefore, lidar observations are computed using the Mie solution included in the scatter-1.1 package for a backscatter lidar, given that such a lidar is expected to be flown with a Ku-band radar and microwave radiometer as part of the NASA Earth System Observatory (ESO) Atmosphere Observing System (AOS) in an inclined and/or polar orbit \citep{yorks2022}.While uncertainties in the lidar forward model are rather complex and difficult to quantify, uncertainties in the lidar observations may be accounted for to some extent by including a multiplicative random noise factor in the calculated lidar backscatter values.
+Specifically, the calculated lidar backscatter values are multiplied with a log-normally distributed random variable with 0.0 mean and a standard deviation of 0.1.
+This results in uncertainties of about 10\% in the lidar backscatter values, conservative compared to the expected performance of the AOS lidars at 532 nm but consistent the performance of Cloud-Aerosol Transport System (CATS) lidar system at 1064 nm reported in \citep{pauly2019}.
+While the AOS lidars will have expected backscatter accuracies of 2-5\%, the lidar simulations in this paper are rather idealized in that they do not include random errors due to daytime solar background noise or systematic errors such as calibration accuracy or lidar ratio assumptions.
+More advanced models of the observation errors exist \citep{liu2006}, but are not considered here.
+There are three main reasons this approach was taken: (1) it is important to understand the limitations of combining these three datasets before factoring in individual sensor limitations, (2) while more accurate calculations based on more realistic ice particle shapes exist, they are rather incomplete and not readily available, and (3) Wagner and Delene (2022) compared lidar backscatter observations with backscatter calculations based on coincident PSD observations and the Mie solution and found good agreement, which suggests that electromagnetic properties derived from Mie calculations are adequate for practical applications.
+The lidar molecular backscatter and extinction are calculated using the lidar module of the CFMIP Observation Simulator Package (COSP; Bodas-Salcedo et al.
+(2011)).
+COSP simulates the lidar total attenuated backscatter signal and scattering ratios at 532 nm for scenes with and without clouds.
+COSP also assumes cloud particles are spherical so that the backscattering phase function is estimated based on the effective radius (Mie theory).
+Despite this assumption, the COSP simulations agree well with CALIOP observations.
+To account for multiple-scattering in the lidar observations, we are using the multiscatter-1.2.11 model \citep{multiscatter-1.2.11} of \cite{hogan2008lidar_ms}.
 
-Shown in Figure 3 are the distributions of simulated Ku-band radar reflectivity and 
-lidar backscatter as function of height above the freezing level.
-As apparent in the figure, the IWCs associated with detectable Ku-band reflectivity
-signal are likely to
-occur near mostly around 3.0 km above the freezing level, while the lidar backscatter
-distribution exhibits a peak at around 6.0-7.0 km
-above the freezing level, which is consistent with the fact the lidar observations
-are strongly attenuated in the bottom part of the cloud.  It should be noted that, consistently with the objective of
- this study, only CS reflectivity profiles with no echo at or below the freezing level were selected
-and used in the calculation of Ku-band reflectivity distributions shown in Figure \ref{f3}.
-The vertical resolutions of
-the radar and lidar observations are the same, i.e. 240 m, and the same as the resolution of the
-CS observations upon which they are based. In reality, the radar and lidar observations are likely to have
-different vertical resolutions and footprint sizes, which would need to be accounted for in a more 
-realistic study.  It should be noted though that the AOS radars and radiometers are expected to achieve
-Nyquist sampling, which enables resolution enhancement \citep{early2001image}.  Nevertheless, differences
-in the vertical resolutions of the radar and lidar observations and differences in the footprint sizes
-may deteriorate the performance of the retrieval algorithm.
+Shown in Figure 3 are the distributions of simulated Ku-band radar reflectivity and lidar backscatter as function of height above the freezing level.
+As apparent in the figure, the IWCs associated with detectable Ku-band reflectivity signal are likely to occur near mostly around 3.0 km above the freezing level, while the lidar backscatter distribution exhibits a peak at around 6.0-7.0 km above the freezing level, which is consistent with the fact the lidar observations are strongly attenuated in the bottom part of the cloud.
+It should be noted that, consistently with the objective of this study, only CS reflectivity profiles with no echo at or below the freezing level were selected and used in the calculation of Ku-band reflectivity distributions shown in Figure \ref{f3}.
+The vertical resolutions of the radar and lidar observations are the same, i.e.
+240 m, and the same as the resolution of the CS observations upon which they are based.
+In reality, the radar and lidar observations are likely to have different vertical resolutions and footprint sizes, which would need to be accounted for in a more realistic study.
+It should be noted though that the AOS radars and radiometers are expected to achieveNyquist sampling, which enables resolution enhancement \citep{early2001image}.
+Nevertheless, differences in the vertical resolutions of the radar and lidar observations and differences in the footprint sizes may deteriorate the performance of the retrieval algorithm.
 
@@
 
-The radiometer observations are calculated using a one-dimensional efficient, but accurate, radiative transfer 
-solver based on Eddington's approximation \citep{kummerow1993edd}. The Eddington's approximation has been found to work 
-well in cloud and precipitation retrieval application despite its simplicity relative to more general (but also 
-computationally intensive) approaches such as Monte Carlo radiative transfer solvers \citep{liu1996three} 
-It should be noted tough that the phase functions of ice particles tend to be highly asymmetric at 
-sub-millimeter wave frequencies. For radiative transfer solutions based on the Eddington's approximation to be 
-accurate, it is necessary that the delta-scaling approach \citep{joseph1976delta} be employed. The delta-scaling 
-approach transforms the initial radiative transfer equation into an equivalent one characterized by a less 
-asymmetric scattering function and more extinction, which makes the solution Eddington approximation more stable 
-and accurate. The absorption due to water vapor and other gases is quantified using the Rosenkranz model 
-\citep{rosenkranz1998}.  The water vapor, temperature, and pressure distributions are derived based on a WRF 
-simulation of summer convection over the United States.  Specifically, the water vapor, temperature, and 
-pressure profiles associated with times and areas where the model produces anvils are selected and clustered 
-into 40 classes using the k-Means approach.  The mean extinction profiles at the radiometer frequencies are 
-calculated for every class 
-and used in process of calculating the brightness temperatures from the estimated ice profiles using a simple 
-Monte Carlo procedure.  That is, given a retrieved ice profile and its scattering property, an anvil class 
-and its associated absorption, temperature, and pressure profiles are randomly selected and attached to the 
-ice scattering properties. To make the procedure physically meaningful, temperature rather than height is 
-used in the ice scattering-gas absorption collocation process.  The surface emissivities are randomly chosen between 
-0.8 and 1.0 and assumed constant for all radiometer frequencies.  Brightness temperatures are calculated at 
+The radiometer observations are calculated using a one-dimensional efficient, but accurate, radiative transfer  solver based on Eddington's approximation \citep{kummerow1993edd}.
+The Eddington's approximation has been found to work well in cloud and precipitation retrieval application despite its simplicity relative to more general (but also computationally intensive) approaches such as Monte Carlo radiative transfer solvers \citep{liu1996three} It should be noted tough that the phase functions of ice particles tend to be highly asymmetric at sub-millimeter wave frequencies.
+For radiative transfer solutions based on the Eddington's approximation to be accurate, it is necessary that the delta-scaling approach \citep{joseph1976delta} be employed.
+The delta-scaling approach transforms the initial radiative transfer equation into an equivalent one characterized by a less asymmetric scattering function and more extinction, which makes the solution Eddington approximation more stable and accurate.
+The absorption due to water vapor and other gases is quantified using the Rosenkranz model \citep{rosenkranz1998}.
+The water vapor, temperature, and pressure distributions are derived based on a WRF simulation of summer convection over the United States.
+Specifically, the water vapor, temperature, and pressure profiles associated with times and areas where the model produces anvils are selected and clustered into 40 classes using the k-Means approach.
+The mean extinction profiles at the radiometer frequencies are calculated for every class and used in process of calculating the brightness temperatures from the estimated ice profiles using a simple Monte Carlo procedure.
+That is, given a retrieved ice profile and its scattering property, an anvil class and its associated absorption, temperature, and pressure profiles are randomly selected and attached to the ice scattering properties.
+To make the procedure physically meaningful, temperature rather than height is used in the ice scattering-gas absorption collocation process.
+The surface emissivities are randomly chosen between 0.8 and 1.0 and assumed constant for all radiometer frequencies.
+Brightness temperatures are calculated at
 %89-, 183.31 ± 1.1, and 325.15 ± 1.5 GHz,
-all frequencies of the 10 channels of the SAPHIR-NG 
-radiometer envisioned to be deployed in the AOS mission \citep{brogniez2022time}.  One of the frequencies is 
-89-GHz, while the others are centered on the 183.31, and 325.15 GHz water vapor absorption lines. Details regarding
-the radiometer and the active instruments are
-provided in Table \ref{t_sensors}.  Errors in the radiometer observations are modeled assuming a 
-Noise-Equivalent-Delta-T (NEDT) of 1K, which is a readily achievable level for modern satellite radiometers \citep{draper2015}.
+all frequencies of the 10 channels of the SAPHIR-NG radiometer envisioned to be deployed in the AOS mission \citep{brogniez2022time}.
+One of the frequencies is 89-GHz, while the others are centered on the 183.31, and 325.15 GHz water vapor absorption lines.
+Details regarding the radiometer and the active instruments are provided in Table \ref{t_sensors}.
+Errors in the radiometer observations are modeled assuming a Noise-Equivalent-Delta-T (NEDT) of 1K, which is a readily achievable level for modern satellite radiometers \citep{draper2015}.
 
-The processing steps used to process the CS reflectivity observations and calculate the lidar, Ku-band and 
-submillimeter-wave radiometer observations may be summarized as follows:
+The processing steps used to process the CS reflectivity observations and calculate the lidar, Ku-band and submillimeter-wave radiometer observations may be summarized as follows:
 \begin{enumerate}
-\item Derivation of physically consistent radar and radiometer lookup tables to relate basic radar and 
-radiometer properties (e.g. reflectivity, attenuation, extinction, scattering-albedo, etc.) to PSD parameters 
-such as $IWC$ and $D_m$. The tables are derived for a single of $N_w$, but are usable with any value of 
+\item Derivation of physically consistent radar and radiometer lookup tables to relate basic radar and radiometer properties (e.g.
+reflectivity, attenuation, extinction, scattering-albedo, etc.) to PSD parameters such as $IWC$ and $D_m$.
+The tables are derived for a single of $N_w$, but are usable with any value of
 $N_w$ using the "normalization" operations described in \citep{grecu2011}.
 \item Derivation of $N_w$-relative height parameterization using the 2C-ICE product.
-\item Estimation of IWC and related PSD parameters from CS W-band radar observations, using the tables 
-constructed Step 1, and $N_w$ profiles derived through parameterization obtained in Step 2 and perturbed using the random model 
-described above. Specifically, given $N_w$ and the associated $Z$ at W-band, we estimate $IWC/N_w$ as a function of $Z/N_w$ using the
-normalized tables and then derive $IWC$ from $IWC/N_w$ and $N_w$.
-\item Calculation of lidar, Ku-band radar and radiometer observations from the estimates derived in Step 3 
-and the tables obtained in Step 1, and extended with non-zero 2C-ICE estimates for the radar bins characterized by no echo
+\item Estimation of IWC and related PSD parameters from CS W-band radar observations, using the tables constructed Step 1, and $N_w$ profiles derived through parameterization obtained in Step 2 and perturbed using the random model described above.
+Specifically, given $N_w$ and the associated $Z$ at W-band, we estimate $IWC/N_w$ as a function of $Z/N_w$ using the normalized tables and then derive $IWC$ from $IWC/N_w$ and $N_w$.
+A simple forward attenuation correction \citep{meneghini1978} is applied to the W-band reflectivity observations prior to the estimation of $IWC$.
+Attenuation due to water vapor and other gases is calculated using the Rosenkranz model \citep{rosenkranz1998} assuming the same temperature, pressure and temperature profiles used in the radiometer simulations.
+
+\item Calculation of lidar, Ku-band radar and radiometer observations from the estimates derived in Step 3 and the tables obtained in Step 1, and extended with non-zero 2C-ICE estimates for the radar bins characterized by no echo
 above the noise level.
@@
 
-The application of these steps produces a large dataset of approximately 800,000 cloud ice profiles and 
-associated lidar, radar and radiometer observations that may be used to investigate the synergy of the three 
-sensors. Details are provided in the next section.
+The application of these steps produces a large dataset of approximately 800,000 cloud ice profiles and associated lidar, radar and radiometer observations that may be used to investigate the synergy of the three sensors.
+Details are provided in the next section.
 
 \subsection{Estimation and evaluation}
-Given that the lidar observations may attenuate quickly in thick clouds, while the Ku-band radar will not detect
-clouds with an echo weaker than 8.0 dBZ, the radiometer is the instrument likely to provide by itself the most 
-complete information about the total amount of ice in its observing volume.  However, the vertical distribution 
-of ice is difficult to quantify from radiometer-only observations, because significantly different ice vertical 
-distributions may lead to very similar radiometer observations.  This makes radiometer-only retrievals highly 
-dependent on the "a priori" information on the distribution of ice clouds in the atmosphere. As previously 
-mentioned, this is the reason why CS-based IWC retrievals were preferred to CRM simulations, as retrievals are 
-expected to result in more natural and less biased distributions.
+Given that the lidar observations may attenuate quickly in thick clouds, while the Ku-band radar will not detect clouds with an echo weaker than 8.0 dBZ, the radiometer is the instrument likely to provide by itself the most complete information about the total amount of ice in its observing volume.
+However, the vertical distribution of ice is difficult to quantify from radiometer-only observations, because significantly different ice vertical distributions may lead to very similar radiometer observations.
+This makes radiometer-only retrievals highly dependent on the "a priori" information on the distribution of ice clouds in the atmosphere.
+As previously mentioned, this is the reason why CS-based IWC retrievals were preferred to CRM simulations, as retrievals are expected to result in more natural and less biased distributions.
 
 For retrievals, we employ a two-step estimation methodology similar to that of \cite{grecu2018}.  In the first step, we 
-estimate an IWC class, out of the 18 classes of shown in Figure \ref{f1},
-to which the estimated IWC profile is most likely to belong. The class estimation procedures is 
-trained using the synthetic observations.
-In the second step, we estimate the IWC profile,
-using a class specific ensemble Kalman Smoother (EKS) methodology similar to that of \cite{grecu2018}.  
+estimate an IWC class, out of the 18 classes of shown in Figure \ref{f1}, to which the estimated IWC profile is most likely to belong.
+The class estimation procedure is trained using the synthetic observations.
+In the second step, we estimate the IWC profile, using a class specific ensemble Kalman Smoother (EKS) methodology similar to that of \cite{grecu2018}.
 The EKS algorithm updates the estimated IWC relative to the mean IWC of the class to which the profile belongs.
-The differences between the actual active and passive observations and their mean class values are used
-in the update. The 
-second step of this procedure is formally identical to the one used in \cite{grecu2018}, but the first step is
-different. In \cite{grecu2018}, the first step was based on a simple distance-based evaluation. That strategy
-is likely to be suboptimal in this study, because the joint distribution of IWC profiles and associated
-observations are significantly more complex. We, therefore, use a more complex classification methodology
-based on the TensorFlow library \citep{tensorflow2016}. The class estimation model is defined as a
-TensorFlow Model with two dense layers of 30 neurons each, followed by a softmax layer \citep{deepL2016}. The 
-class estimation model is trained using the 70\% of the simulated observations and the corresponding IWC 
-profiles, while the remaining 30\% of the data being used for evaluation. 
-The EKS update is based on the formula%The training is done using the Adam
+The differences between the actual active and passive observations and their mean class values are used in the update.
+The second step of this procedure is formally identical to the one used in \cite{grecu2018}, but the first step is different.
+In \cite{grecu2018}, the first step was based on a simple distance-based evaluation.
+That strategy is likely to be suboptimal in this study, because the joint distribution of IWC profiles and associated observations are significantly more complex.
+We, therefore, use a more complex classification methodology based on the TensorFlow library \citep{tensorflow2016}.
+The class estimation model is defined as a
+TensorFlow Model with two dense layers of 30 neurons each, followed by a softmax layer \citep{deepL2016}.
+The activation function used in the dense layer is Rectified Linear Unit (ReLU) \citep{nair2010}.
+The class estimation model is trained using the 70\% of the simulated observations and the corresponding IWC profiles, while the remaining 30\% of the data being used for evaluation.
+The EKS update is based on formula%The training is done using the Adam
 \begin{equation}
@@
 \end{equation}
-where $\mathbf{X}$ is the state variable describing the IWC profile, $\mathbf{Y}$ is the vector containing
-the actual observations, $\mathbf{X_i}$ is the set of state variables for profiles
-in class $i$, and $\mathbf{Y_i}$ is the
-set of observations associated with profiles in class $i$.  Variables $\mathbf{\bar{X}_i}$ and $\mathbf{\bar{Y}_i}$ are the mean values
-of the state variables and observations in class $i$, respectively. The covariance matrices between 
-$\mathbf{X_i}$ and $\mathbf{Y_i}$ are denoted by $\mathbf{Cov(X_i,Y_i)}$. In step 1, the class is estimated
-using the TensorFlow model, while in step 2, the IWC profile is estimated using the EKS algorithm summarized in 
-Equation \ref{eq:enks}.
+where $\mathbf{X}$ is the state variable describing the IWC profile, $\mathbf{Y}$ is the vector containing the actual observations, $\mathbf{X_i}$ is the set of state variables for profiles in class $i$, and $\mathbf{Y_i}$ is the set of observations associated with profiles in class $i$.
+Variables $\mathbf{\bar{X}_i}$ and $\mathbf{\bar{Y}_i}$ are the mean values of the state variables and observations in class $i$, respectively.
+The covariance matrices between  $\mathbf{X_i}$ and $\mathbf{Y_i}$ are denoted by $\mathbf{Cov(X_i,Y_i)}$.
+In step 1, the class is estimated using the TensorFlow model, while in step 2, the IWC profile is estimated using the EKS algorithm summarized in Equation \ref{eq:enks}.
 
-As already mentioned, a cross-validation methodology is used for evaluation, with 70\% of 
-the data used for training and the remaining 30\% of the data used for validation.  The partition of the data into 
-training and 
-evaluation subsets is done randomly.  Usually, the partition, training and evaluation steps are repeated 
-several times.  However, given the fact that differences in the relationships between the ice property and 
-their associated simulated observations are functions of the meteorological context, and that all regimes are 
-well-sampled in both the training and testing subsets (e.g. out of every 10 pixels in a scene, about seven end up 
-in the training dataset, while the others in the testing dataset), the repetition of the partition, training, 
-and evaluation steps multiple times is not necessary. Therefore, in our evaluation, we partition the data into 
-training and evaluation only once and perform all the evaluation for a single partition. The evaluation 
-criteria include the correlation coefficient, the bias, and visual inspections of graphical representations of 
-the estimated properties relative to their references.
+As already mentioned, a hold-out validation methodology is used for evaluation, with 70\% of the data used for training and the remaining 30\% of the data used for validation.
+The partition of the data into training and evaluation subsets is done randomly.
+In a more general cross-validation methodology, the partition, training and evaluation steps are repeated several times.
+However, given the fact that differences in the relationships between the ice property and their associated simulated observations are functions of the meteorological context, and that all regimes are well-sampled in both the training and testing subsets (e.g.
+out of every 10 pixels in a scene, about seven end up in the training dataset, while the others in the testing dataset), the repetition of the partition, training, and evaluation steps multiple times is not necessary.
+Therefore, in our evaluation, we partition the data into training and evaluation only once and perform all the evaluation for a single partition.
+The evaluation criteria include the correlation coefficient, the bias, and visual inspections of graphical representations of the estimated properties relative to their references.
 %vs
@@
 \subsection{Radiometer-only retrievals}
-As previously mentioned, submillimeter-wave radiometers are likely to provide by themselves more complete 
-information about the total amount of ice in their observing volumes than lidars or Ku-band radars with 
-limited sensitivity. However, radiometers observations are an integrated measure of radiative processes in 
-clouds that provide little information about the vertical distribution of ice. From this perspective, an 
-evaluation in terms of the ice water path (IWP) defined as the vertical integral of the IWC, i.e. 
-$IWP=\int_0^{Z_{top}}IWC(z)dz$ is insightful.  Shown in Figure \ref{f4} is the distribution of IWP 
-estimated from radiometer-only observations as a function of its true value. As apparent in the figure, there is good 
-correlation between the retrieved and the true IWP values. The numerical value of the correlation coefficient 
-is 0.92, and there is no-overall bias. That is, the mean values of retrieved IWP and true IWP values are 
-equal. However, conditional biases are apparent, with overestimation of IWP for values smaller than 
-100 g/$m^2$ and some underestimation for values larger than 1000 g/$m^2$.  The biases at the low end of the 
-IWP range are not surprising, given that the impact caused by ice scattering on the total radiometric signal 
-is small for low values of IWP and hard to distinguish from other sources of variability in radiometer 
-observations. Saturation effects are most likely responsible for underestimation at the high end.  It should 
-be noted that in this evaluation, only atmospheric profiles that exhibit ice detectable by the CS radar or CALIOP lidar are used. 
-Therefore,
-a radiometer-only estimation procedure derived from this training dataset is likely to result in significant 
-overestimation if not used in conjunction with a discrimination procedure. However, such procedure is not critical 
-in this study, as, in a synergistic application, the lidar observations may be used to discriminate between clear skies and ice clouds.
+As previously mentioned, submillimeter-wave radiometers are likely to provide by themselves more complete information about the total amount of ice in their observing volumes than lidars or Ku-band radars with limited sensitivity.
+However, radiometers observations are an integrated measure of radiative processes in clouds that provide little information about the vertical distribution of ice.
+From this perspective, an evaluation in terms of the ice water path (IWP) defined as the vertical integral of the IWC, i.e.
+$IWP=\int_0^{Z_{top}}IWC(z)dz$ is insightful.
+Shown in Figure \ref{f4} is the distribution of IWP estimated from radiometer-only observations as a function of its true value.
+As apparent in the figure, there is good correlation between the retrieved and the true IWP values.
+The numerical value of the correlation coefficient is 0.92, and there is no-overall bias.
+That is, the mean values of retrieved IWP and true IWP values are equal.
+However, conditional biases are apparent, with overestimation of IWP for values smaller than 100 g/$m^2$ and some underestimation for values larger than 1000 g/$m^2$.
+The biases at the low end of the IWP range are not surprising, given that the impact caused by ice scattering on the total radiometric signal is small for low values of IWP and hard to distinguish from other sources of variability in radiometer observations.
+Saturation effects are most likely responsible for underestimation at the high end.
+It should be noted that in this evaluation, only atmospheric profiles that exhibit ice detectable by the CS radar or CALIOP lidar are used.
+Therefore, a radiometer-only estimation procedure derived from this training dataset is likely to result in significant overestimation if not used in conjunction with a discrimination procedure.
+However, such procedure is not critical in this study, as, in a synergistic application, the lidar observations may be used to discriminate between clear skies and ice clouds.
 \begin{figure}[t]
@@
 However, although the radiometer-only estimation procedure is able to estimate the integrated amount of ice in clouds fairly
-well, its ability to characterize the vertical distribution of ice in clouds is limited.  Figure \ref{f5} shows the
-conditional vertical distributions of the estimated and true IWC for the 18 classes described in Section 2a 
-and shown in Figure \ref{f1}. As apparent in the figure, there are significant differences between the
-estimated and true IWC profiles. 
+well, its ability to characterize the vertical distribution of ice in clouds is limited.
+Figure \ref{f5} shows the conditional vertical distributions of the estimated and true IWC for the 18 classes described in Section 2a and shown in Figure \ref{f1}.
+As apparent in the figure, there are significant differences between the estimated and true IWC profiles.
 \begin{figure}[t]
@@
 
-Further insight into the radiometer-only estimation performance may be derived by defining the ice profile gravity center 
-(GC) as $z_{GC}=\frac {\int_0^{Z_{top}}zIWC(z)dz} {\int_0^{Z_{top}}IWC(z)dz}$, where $z$ is the distance relative to 
-the freezing level, the $Z_{top}$ is the distance from the top of the atmosphere to the freezing level.  
-Shown in Figure \ref{f6} is the frequency of IWC gravity center
-estimated from radiometer-only observations as a function of its true value.  It may be observed in the figure that while 
-the true IWC gravity center exhibits quite a broad distribution, the one retrieved from the radiometer-only 
-observations exhibits a multimodal narrow distribution. Moreover, there correlation between the retrieved 
-and the true IWC gravity center is rather low.  This is another indication that, while the total amount of ice may be 
-reasonably estimated from radiometer-only observations, its vertical distribution can not be 
-accurately determined from radiometer-only observations.
+Further insight into the radiometer-only estimation performance may be derived by defining the ice profile gravity center (GC) as $z_{GC}=\frac {\int_0^{Z_{top}}zIWC(z)dz} {\int_0^{Z_{top}}IWC(z)dz}$, where $z$ is the distance relative to the freezing level, the $Z_{top}$ is the distance from the top of the atmosphere to the freezing level.
+Shown in Figure \ref{f6} is the frequency of IWC gravity center estimated from radiometer-only observations as a function of its true value.
+It may be observed in the figure that while the true IWC gravity center exhibits quite a broad distribution, the one retrieved from the radiometer-only observations exhibits a multimodal narrow distribution.
+Moreover, there correlation between the retrieved and the true IWC gravity center is rather low.
+This is another indication that, while the total amount of ice may be reasonably estimated from radiometer-only observations, its vertical distribution can not be accurately determined from radiometer-only observations.
 
@@
 
-Although retrievals from the lidar-only or radar-only observations are not expected to be as accurate as those from
-radiometer-only observations, the lidar observations being subject to severe attenuation while the radar observations
-being limited by sensitivity, they are nevertheless insightful. This is because quantifying the limitations of
-retrievals from active instruments may be used to better assess  the benefits of synergistic retrievals.  Shown in
-top of row of Figure \ref{f7} are the distributions of IWP estimated from lidar-only and radar-only observations 
-as a function of their true values. As apparent in the figure, and as expected, the lidar-only retrievals tend to
-be accurate for IWP values smaller than 50 g/m$^2$, while the radar-only retrievals tend to be reliable only for large
-IWP values on the order of hundreds of g/m$^2$.  Not that the radar-only retrievals exhibit a bimodal distributions,
-with a broad distribution of real IWP values associated with a single retrieved IWP value.  This is a consequence
-of the fact that there is a large number of atmospheric profiles characterized by not necessarily small IWP values that
-are not associated with detectable radar signals. The results shown in the top row of Figure \ref{f7} are conditional on
-either the CALIOP lidar or CS radar observations detecting ice, and so are the radar-only retrievals. However, the 
-indiscriminate application of a Ku-band radar-only retrieval procedure to the entire dataset may result in significant
-overestimation because clear skies would be associated with the same IWP value as ice clouds with reflectivity signal 
-below the detection threshold. On the other hand, limiting the application of the radar-only retrieval procedure to atmospheric profiles
-with detectable ice clouds would result in a significant underestimation, as a large number of atmospheric profiles
-with ice clouds would be associated with zero IWP.  From this perspective, the derivation of  Ku-band radar-only ice estimates
-is not useful if coincident radiometer observations are available.
+Although retrievals from the lidar-only or radar-only observations are not expected to be as accurate as those from radiometer-only observations, the lidar observations being subject to severe attenuation while the radar observations being limited by sensitivity, they are nevertheless insightful.
+This is because quantifying the limitations of retrievals from active instruments may be used to better assess  the benefits of synergistic retrievals.
+Shown in the left column of Figure \ref{f7} are the distributions of IWP estimated from lidar-only and radar-only observations as a function of their true values.
+As apparent in the figure, and as expected, the lidar-only retrievals tend to be accurate for IWP values smaller than 50 g/m$^2$, while the radar-only retrievals tend to be reliable only for large IWP values on the order of hundreds of g/m$^2$.
+Not that the radar-only retrievals exhibit a bimodal distributions,with a broad distribution of real IWP values associated with a single retrieved IWP value.
+This is a consequence of the fact that there is a large number of atmospheric profiles characterized by not necessarily small IWP values that are not associated with detectable radar signals.
+The results shown in Figure \ref{f7} are conditional on either the CALIOP lidar or CS radar observations detecting ice, and so are the radar-only retrievals.
+However, the  indiscriminate application of a Ku-band radar-only retrieval procedure to the entire dataset may result in significant overestimation because clear skies would be associated with the same IWP value as ice clouds with reflectivity signal  below the detection threshold.
+On the other hand, limiting the application of the radar-only retrieval procedure to atmospheric profiles with detectable ice clouds would result in a significant underestimation, as a large number of atmospheric profiles with ice clouds would be associated with zero IWP.
+From this perspective, the derivation of  Ku-band radar-only ice estimates is not useful if coincident radiometer observations are available.
 
-Shown in the bottom row of Figure \ref{f7} are the distributions of IWC gravity centers
-estimated from lidar-only and radar-only observations as a function of their true values. Results are similar to those
-obtained for IWP, in the sense that the radar-only retrieval distribution is bimodal and reliable only for ice clouds
-with Ku-band radar observations above the detection threshold.
-The lidar-only retrievals produce a much broader IWC gravity center distribution, with values that exhibit moderate 
-correlation with the true IWC gravity centers.
+Shown in left column of Figure \ref{f7} are the distributions of IWC gravity centers estimated from lidar-only and radar-only observations as a function of their true values.
+Results are similar to those obtained for IWP, in the sense that the radar-only retrieval distribution is bimodal and reliable only for ice clouds with Ku-band radar observations above the detection threshold.
+The lidar-only retrievals produce a much broader IWC gravity center distribution, with values that exhibit moderate  correlation with the true IWC gravity centers.
+
+The synergistic retrievals of IWP and IWC gravity centers from synergistic lidar and radar observations are shown in the bottom row of Figure \ref{f7}.
+As apparent in the figure, the synergistic lidar-radar retrievals are overall superior to both lidar- and radar-only retrievals.
+The lidar-radar retrievals appear inferior to the lidar-only ones for thin high clouds (characterized low IWP and gravity center above 8.0 km above the freezing level).
+This is that a consequence of the fact, to minimize overall errors, the lidar and lidar-radar estimation procedure interpret lidar observations differently.
 
@@
     \includegraphics[width=.481\linewidth]{fig07a.rev.png}
+    \includegraphics[width=.481\linewidth]{fig07c.rev.png}
     \includegraphics[width=.481\linewidth]{fig07b.rev.png}
-    \includegraphics[width=.481\linewidth]{fig07c.rev.png}
     \includegraphics[width=.481\linewidth]{fig07d.rev.png}
-    \caption{Top Row: Density plots of IWP retrievals from lidar-only (left), and radar-only (right) observations as a function
-    of true IWP. Bottom Row: Same as in the top row but for IWC gravity center.}\label{f7}
+    \includegraphics[width=.481\linewidth]{fig7e0.png}
+    \includegraphics[width=.481\linewidth]{fig7e.png}
+    \caption{Top Row: Density plots of lidar-only IWP retrievals (left), and retrieved IWC gravity centers (right) observations as a function of their true values.
+Middle Row: Same as in the top row but for radar-only retrievals.
+Bottom Row: Same as in the top row but for lidar-radar retrievals}\label{f7}
 \end{figure}
@@
 
-The synergy of the instrument on the estimates may be investigated by simply incorporating lidar 
-and radar observations
-into the retrieval process and comparing the results with the radiometer-only estimates.  Although
-the lidar observations are subject to attenuation, they are able to provide information about the vertical
-distribution of ice in clouds, mostly at the top of the clouds. The radar observations, on the other hand,
-are able to provide information in the bottom part of the clouds, where the lidar signal is below the noise
-level due to attenuation. Therefore, the combined used of lidar and radar observations is expected to 
-provide a more complete characterization of the vertical distribution of ice in clouds and enable the
-derivation of more specific estimates than those derived from radiometer-only observations. It should be mentioned that, 
-although deficiencies 
-and potential biases in the simulated observations may distort conclusions to some degree, the forward models
-used in this study are state-of-the-art and are expected to enable a realistic characterization of the impact
-of individual instruments on the synergistic retrievals.
+The synergy of the instrument on the estimates may be investigated by simply incorporating lidar and radar observations into the retrieval process and comparing the results with the radiometer-only estimates.
+Although the lidar observations are subject to attenuation, they are able to provide information about the vertical distribution of ice in clouds, mostly at the top of the clouds.
+The radar observations, on the other hand, are able to provide information in the bottom part of the clouds, where the lidar signal is below the noise level due to attenuation.
+Therefore, the combined used of lidar and radar observations is expected to provide a more complete characterization of the vertical distribution of ice in clouds and enable the derivation of more specific estimates than those derived from radiometer-only observations.
+It should be mentioned that, although deficiencies and potential biases in the simulated observations may distort conclusions to some degree, the forward models used in this study are state-of-the-art and are expected to enable a realistic characterization of the impact of individual instruments on the synergistic retrievals.
 
-Shown in Figure \ref{f8} is the distribution of the synergistic IWP estimates as a function of their true
-values.  As apparent in the figure, the synergistic IWP estimates are more accurate than the radiometer-only
-estimates. At the same time, as apparent in Figure \ref{f9}, the retrieved conditional mean IWC for the 
-18 classes described section 2a and shown in Figure \ref{f1} are in significantly better
-agreement with the true IWC profiles than those derived from radiometer-only observations.  Moreover,
-as seen in Figure \ref{f10} the synergistic IWC gravity center estimates are in much better agreement with 
-the true IWC gravity center than those derived from single-instrument observations.
+Shown in Figure \ref{f8} is the distribution of the synergistic IWP estimates as a function of their true values.
+As apparent in the figure, the synergistic IWP estimates are more accurate than the radiometer-only estimates.
+At the same time, as apparent in Figure \ref{f9}, the retrieved conditional mean IWC for the 18 classes described section 2a and shown in Figure \ref{f1} are in significantly better agreement with the true IWC profiles than those derived from radiometer-only observations.
+Moreover, as seen in Figure \ref{f10} the synergistic IWC gravity center estimates are in much better agreement with the true IWC gravity center than those derived from single-instrument observations.
 
@@
 
-While the estimates based on all instruments are significantly more accurate than those based on radiometer-only
-observations, it is useful to investigate how the two active instruments (lidar and radar) impact the
-estimates.  For conciseness, we use two statistical scores, namely, the normalized root mean square (NRMS) and
-the classification accuracy, to summarize the performance of the estimates.  The NRMS is defined as
+While the estimates based on all instruments are significantly more accurate than those based on radiometer-only observations, it is useful to investigate how the two active instruments (lidar and radar) impact the estimates.
+For conciseness, we use two statistical scores, namely, the normalized root mean square (NRMS) and the classification accuracy, to summarize the performance of the estimates.
+The NRMS is defined as
 \begin{equation}
@@
 \end{equation}
-where $\delta_{i}$ is a binary variable that is equal to 1 if the estimated IWC class for the $i$-th sample is
-equal to the true IWC class for the $i$-th sample, and 0 otherwise. The performance summary is shown 
-in Table \ref{t1} for several combinations of instruments. It may be observed in the table that the
-performance of the estimates based on all instruments is significantly better than those based on
-radiometer-only observations.  Furthermore, the inclusion of the lidar observations in the retrieval
-process has a larger impact on the retrieval performance than the inclusion of the radar
-observations.  This is expected since the lidar observations are able to provide information about the
-top of the clouds, where the radar observations are above the noise level only occasionally. Nevertheless,
-the inclusion of the radar observations in the retrieval process has a notable impact on the
-accuracy of the IWC estimates relative to radiometer-only retrievals.
+where $\delta_{i}$ is a binary variable that is equal to 1 if the estimated IWC class for the $i$-th sample is equal to the true IWC class for the $i$-th sample, and 0 otherwise.
+The performance summary is shown in Table \ref{t1} for several combinations of instruments.
+It may be observed in the table that the performance of the estimates based on all instruments is significantly better than those based on radiometer-only observations.
+Furthermore, the inclusion of the lidar observations in the retrieval process has a larger impact on the retrieval performance than the inclusion of the radar
+observations.
+This is expected since the lidar observations are able to provide information about the top of the clouds, where the radar observations are above the noise level only occasionally.
+Nevertheless, the inclusion of the radar observations in the retrieval process has a notable impact on the accuracy of the IWC estimates relative to radiometer-only retrievals.
 
-Additional insight may be gained by examining the distribution of the retrieved of IWP and IWC gravity center as
-a function of their true values.  These distributions are show in Figure \ref{f11} for the lidar-radiometer
-and radar-radiometer retrievals.  As apparent in the figure, the inclusion of radiometer observations improves both
-the lidar and radar retrievals. However, neither of the two instrument combinations is able to provide estimates
-that are as accurate as those derived from all instruments.  Moreover, the lidar-radiometer IWP retrievals appear to be
-less accurate than lidar-only retrievals. This may be an artifact of statistical inversion methodology, which may 
-make suboptimal use of correlations in the observations and improve the accuracy of the estimates for some type of 
-profiles at the expense of others.  While other statistical inversion methodologies may be able to eliminate this
-artifact, given the complexity of the problem and the multiple sources of uncertainty, no better methodology is
+Additional insight may be gained by examining the distribution of the retrieved of IWP and IWC gravity center as a function of their true values.
+These distributions are show in Figure \ref{f11} for the lidar-radiometer and radar-radiometer retrievals.
+As apparent in the figure, the inclusion of radiometer observations improves both the lidar and radar retrievals.
+However, neither of the two instrument combinations is able to provide estimates that are as accurate as those derived from all instruments.
+Moreover, the lidar-radiometer IWP retrievals appear to be less accurate than lidar-only retrievals.
+This may be an artifact of statistical inversion methodology, which may make suboptimal use of correlations in the observations and improve the accuracy of the estimates for some type of profiles at the expense of others.
+While other statistical inversion methodologies may be able to eliminate this artifact, given the complexity of the problem and the multiple sources of uncertainty, no better methodology is
 obvious yet.
@@
 \section{Conclusions}
-In this study, we investigate the synergy of elastic backscatter lidar, Ku-band radar, and sub-millimeter-wave radiometer
-measurements in the retrieval of the ice from satellite observations.  The synergy is analyzed through
-the generation of a large dataset of IWC profile and the calculation of lidar, radar and radiometer 
-observations using realistic models. The characteristics of the instruments (e.g. frequencies, sensitivities,
-etc.) are set based on the expected characteristics of instruments of the AOS mission. 
-A cross-validation methodology is used to assess the 
-accuracy of the retrieved IWC profiles from various combinations of observations from the three instruments.
-Specifically, the IWC and associated observations is randomly divided into two datasets, one for
-training and the other for evaluation.  The training dataset is used to train the retrieval algorithm,
-while the evaluation dataset is used to assess the retrieval performance.
+In this study, we investigate the synergy of elastic backscatter lidar, Ku-band radar, and sub-millimeter-wave radiometer measurements in the retrieval of the ice from satellite observations.
+The synergy is analyzed through the generation of a large dataset of IWC profile and the calculation of lidar, radar and radiometer observations using realistic models.
+The characteristics of the instruments (e.g.
+frequencies, sensitivities, etc.) are set based on the expected characteristics of instruments of the AOS mission.
+A hold-out validation methodology is used to assess the accuracy of the retrieved IWC profiles from various combinations of observations from the three instruments.
+Specifically, the IWC and associated observations is randomly divided into two datasets, one for training and the other for evaluation.
+The training dataset is used to train the retrieval algorithm, while the evaluation dataset is used to assess the retrieval performance.
 
-To ensure the self-consistency of results and their relevance to practical applications, the dataset of IWC 
-profiles is derived from CloudSat reflectivity observations and extended with lidar-based estimates from the
-2C-ICE product. Although subject to potential biases
-and uncertainties due to deficiencies in the retrieval models, these profiles are deemed to be more realistic
-than those derived from cloud resolving model simulations. Moreover, they are roughly consistent with
-the 2C-ICE CloudSat product \citep{deng2015}, while relying on assumptions and parameterizations that
-enable the accurate computation of backscatter lidar, Ku-band radar, and sub-millimiter-wave radiometer
-observations.
+To ensure the self-consistency of results and their relevance to practical applications, the dataset of IWC profiles is derived from CloudSat reflectivity observations and extended with lidar-based estimates from the 2C-ICE product.
+Although subject to potential biases and uncertainties due to deficiencies in the retrieval models, these profiles are deemed to be more realistic than those derived from cloud resolving model simulations.
+Moreover, they are roughly consistent with
+the 2C-ICE CloudSat product \citep{deng2015}, while relying on assumptions and parameterizations that enable the accurate computation of backscatter lidar, Ku-band radar, and sub-millimiter-wave radiometer observations.
 
-The retrieval of the ice water content (IWC) profiles from the computed observations is achieved in two
-steps.  In the first step, a class, out of 18 potential classes characterized by different 
-vertical distribution of IWC, is estimated from the observations. The 18 classes are predetermined based on
-k-Means clustering algorithm.  In the second step, the IWC profile is estimated using an Ensemble Kalman
-Smoother (EKS) algorithm that uses the estimated class as a priori information.
+The retrieval of the ice water content (IWC) profiles from the computed observations is achieved in two steps.
+In the first step, a class, out of 18 potential classes characterized by different  vertical distribution of IWC, is estimated from the observations.
+The 18 classes are predetermined based on k-Means clustering algorithm.
+In the second step, the IWC profile is estimated using an Ensemble Kalman Smoother (EKS) algorithm that uses the estimated class as a priori information.
 
-The results of the study show that the synergy of lidar, radar, and radiometer observations is significant
-in the retrieval of the IWC profiles.  The inclusion of the lidar observations in the retrieval process has a
-larger impact on the retrieval performance than the inclusion of the radar observations.  Although results are
-not directly comparable to those from other studies \citep{pfreundschuh2020synergistic,liu2022assessing}, given the differences
-between the instruments in this study relative to those from other studies, it may be concluded that they
-are not inconsistent with previous studies. Specifically, previous studies show some skills in radiometer-only
-retrievals and significant improvements in the retrieval performance when the active observations are incorporated.
+The results of the study show that the synergy of lidar, radar, and radiometer observations is significant in the retrieval of the IWC profiles.
+The inclusion of the lidar observations in the retrieval process has a larger impact on the retrieval performance than the inclusion of the radar observations.
+Although results are not directly comparable to those from other studies \citep{pfreundschuh2020synergistic,liu2022assessing}, given the differences between the instruments in this study relative to those from other studies, it may be concluded that they are not inconsistent with previous studies.
+Specifically, previous studies show some skills in radiometer-only retrievals and significant improvements in the retrieval performance when the active observations are incorporated.
 From this perspective, our findings are consistent with previous studies.
 
-Further work is necessary to assess the impact of sources of uncertainties such as potential biases in the
-forward models, variability in the
-PSD intercept not captured by the current parameterization, differences in the instruments' footprint sizes,
-and non-uniform beam filling on the retrievals of the IWC profiles.  Other sources of uncertainties that
-need be considered include the potential existence of supercooled liquid water in the clouds and uncertainties
-in the electromagnetic scattering properties used in the in instruments' forward models. These uncertainties
-may be best investigated and mitigated through the use of high-quality observations from field campaigns
-such as The Investigation of Microphysics and Precipitation for Atlantic Coast-Threatening Snowstorms (IMPACTS)
-\citep{mcmurdie2022}. To achieve its objectives, which were driven by the need to improve the understanding of snowfall processes, 
-remote sensing of snow, and the prediction of banded snow structures \citep{mcmurdie2022}, IMPACTS relied on a suite 
-of active and passive instruments deployed via a satellite-simulating aircraft. These included multiple radars, one of them
-operating at Ku-band, a 532 nm elastic lidar, and a sub-millimeter-wave radiometer similar to the one considered in this study.
-While the objectives of IMPACTS were snowstorms, the instruments used in the campaign sampled a wide range of clouds, 
-including high ice clouds.  The IMPACTS observations associated with high ice clouds may be used to derive IWC estimates that may be
-directly validating using "in-situ" measurements, as the high-altitude aircraft flew in coordination with a cloud penetrating
-aircraft that carried cloud and ice particle probes. These data, although not fully available yet, are expected to
-provide valuable information on the accuracy of the IWC retrievals from synergistic lidar, Ku-band radar and 
-sub-millimeter-wave radiometer observations and enable the refinement of the retrieval algorithm formulated in this study.
+Further work is necessary to assess the impact of sources of uncertainties such as potential biases in the forward models, variability in the PSD intercept not captured by the current parameterization, differences in the instruments' footprint sizes, and non-uniform beam filling on the retrievals of the IWC profiles.
+Other sources of uncertainties that need be considered include the potential existence of supercooled liquid water in the clouds and uncertainties in the electromagnetic scattering properties used in the in instruments' forward models.
+These uncertainties may be best investigated and mitigated through the use of high-quality observations from field campaigns such as The Investigation of Microphysics and Precipitation for Atlantic Coast-Threatening Snowstorms (IMPACTS) \citep{mcmurdie2022}.
+To achieve its objectives, which were driven by the need to improve the understanding of snowfall processes, remote sensing of snow, and the prediction of banded snow structures \citep{mcmurdie2022}, IMPACTS relied on a suite of active and passive instruments deployed via a satellite-simulating aircraft.
+These included multiple radars, one of them operating at Ku-band, a 532 nm elastic lidar, and a sub-millimeter-wave radiometer similar to the one considered in this study.
+While the objectives of IMPACTS were snowstorms, the instruments used in the campaign sampled a wide range of clouds, including high ice clouds.
+The IMPACTS observations associated with high ice clouds may be used to derive IWC estimates that may be directly validating using "in-situ" measurements, as the high-altitude aircraft flew in coordination with a cloud penetrating aircraft that carried cloud and ice particle probes.
+These data, although not fully available yet, are expected to provide valuable information on the accuracy of the IWC retrievals from synergistic lidar, Ku-band radar and sub-millimeter-wave radiometer observations and enable the refinement of the retrieval algorithm formulated in this study.
 As such, they are considered a priority for future work.

### 2023-08-22 RST/Submission/RSTManuscript_rev2.tex
revision 2

@@
 In this study, we parameterize $N_w$ as a function of temperature based on the CloudSat 2C-ICE product \citep{deng2010tropical,deng2013evaluation}.
-A scatter-plot analysis of relationships between the 2C-ICE IWC and the associated reflectivity suggests that the multiplicative coefficient in a power low ice-reflectivity relationship is parameterized as a function of temperature (or equivalently, as a function of height relative to the freezing level) in the default 2C-ICE retrievals.
-The default multiplicative coefficient (the value the provides the IWC estimate prior to the ingestion of the lidar observations) may be simply estimated by regressing the 2C-ICE IWC on the associated CloudSat reflectivity as a function of $H$, defined as height relative the freezing level.
+A scatter-plot analysis of relationships between the 2C-ICE IWC and the associated reflectivity suggests that the multiplicative coefficient in a power law ice-reflectivity relationship is parameterized as a function of temperature (or equivalently, as a function of height relative to the freezing level) in the default 2C-ICE retrievals.
+The default multiplicative coefficient (the value that provides the IWC estimate prior to the ingestion of the lidar observations) may be simply estimated by regressing the 2C-ICE IWC on the associated CloudSat reflectivity as a function of $H$, defined as height relative the freezing level.
 Specifically, a regression of the type $ln(IWC)=ln(a)+b*ln(Z)$ is performed for each bin referenced relative to the zero degree bin is performed, and the variation of $ln(a)$ is parameterized as a function of the relative height.
@@
 reflectivity, attenuation, extinction, scattering-albedo, etc.) to PSD parameters such as $IWC$ and $D_m$.
-The tables are derived for a single of $N_w$, but are usable with any value of
-$N_w$ using the "normalization" operations described in \citep{grecu2011}.
+The tables are derived for a single value of $N_w$, but are usable with any value of $N_w$ using the "normalization" operations described in \citep{grecu2011}.
 \item Derivation of $N_w$-relative height parameterization using the 2C-ICE product.
@@
 The radar observations, on the other hand, are able to provide information in the bottom part of the clouds, where the lidar signal is below the noise level due to attenuation.
-Therefore, the combined used of lidar and radar observations is expected to provide a more complete characterization of the vertical distribution of ice in clouds and enable the derivation of more specific estimates than those derived from radiometer-only observations.
+Therefore, the combined use of lidar and radar observations is expected to provide a more complete characterization of the vertical distribution of ice in clouds and enable the derivation of more specific estimates than those derived from radiometer-only observations.
 It should be mentioned that, although deficiencies and potential biases in the simulated observations may distort conclusions to some degree, the forward models used in this study are state-of-the-art and are expected to enable a realistic characterization of the impact of individual instruments on the synergistic retrievals.

### 2023-10-15 mathPaper/pompeiuTriangles.tex
pompeiu triangles

%% use article style

\documentclass[12pt]{article}
\usepackage{amsmath,amssymb,amsthm,amsfonts}
\usepackage{graphicx}
\title{A new look at Pompeiu Triangles}
\begin{document}
\maketitle
\section{Introduction}
Ptolemey's theorem states that if $ABCD$ is a cyclic quadrilateral, then $AB\cdot CD+BC\cdot DA=AC\cdot BD$. When $D$ does not belong to the circumcircle of $ABC$, then the following inequality holds: $AB\cdot CD+BC\cdot DA>AC\cdot BD$. If $ABC$ is an equilateral triangle (and using $P$ instead of $D$ to denote the fourth point), then $PC+PA \geq PB$. This is because $AB=BC=CA$ and $P$ does not belong to the circumcircle of $ABC$. Given that the inequality above (called the triangle inequality) holds, one can construct a triangle with sides $PA$, $PB$ and $PC$.  Such a triangle is called a Pompeiu triangle. The name comes from the Romanian mathematician Dimitrie Pompeiu, who studied these triangles in 1929. An interesting property of Pompeiu triangles is that their areas can be calculated using elementary geometry. Moreover, their areas are related to other geometric properties. In this paper, we will present a new way of deriving the formula to 
calculate the area of a Pompeiu triangle. The derivation involves circular inversions. We also derive a formula to relate this area to the distance of point D to the circumcircle of $ABC$. A derivation based on inversions was presented by Benyi and Casu (2009). However, the derivation presented here is simpler and more elegant.

\section{Circular inversions}
Inversions are a powerful tool in geometry. They are used to prove many theorems. In this section, we will present the basic properties of inversions. We will also present a formula to calculate the area of a triangle using inversions.

In plane geometry the inverse of a point $P$ with respect to a reference circle with center $O$ and radius $k$ is a point $P'$, lying on the ray from $O$ through $P$ such that

\begin{equation}
OP\cdot OP'=k^2
\label{eq:inversion}
\end{equation}

A remarkable property of inversions is that they transform circles that go through the center of inversion into straight lines. This is a property that we will use in this paper to derive the formula for the area of a Pompeiu triangle. For a proof of this property, see https://artofproblemsolving.com/wiki/
index.php/Circular\_Inversion.
Another property that we will use is that if $A'$ and $B'$ are the inverses of points $A$ and $B$ then

\begin{equation}
{A'B'}={AB}\frac{k^2}{OA\cdot OB}
\label{eq:inversion2}
\end{equation}

\section{The area of a Pompeiu triangle}

%% insert figure 1 here
For a given equilateral triangle $\Delta ABC$, we can apply an inversion with respect to a reference circle with center $A$ and radius $k$, and transform points $A$ and $B$ into point $A'$ and $B'$. A point $P$ interior to triangle $\Delta ABC$ is transformed into $P'$, which is exterior to triangle $\Delta AB'C'$. This is illustrated in Figure \ref{fig:pompeiuTriangles1}. It may be noticed in this figure that the area of $\Delta B'P'C'$ can be determined from the areas of triangles $\Delta AB'P'$, $\Delta AP'C'$ and $\Delta AB'C'$.  More specifically, if we denote the area of any given triangle $\Delta ABC$ by $S(ABC)$, then

\begin{equation}
S(B'P'C')=S(AB'P')+S(AP'C')-S(AB'C')
\label{eq:area1}
\end{equation}

To relate equation ref{eq:area1} to the area of a Pompeiu triangle, we need to express the areas of triangles $\Delta B'P'C'$, $\Delta AB'P'$, $\Delta AP'C'$ and $\Delta AB'C'$ in terms of the untransformed distances among points $A$,$B$,$C$ and $P$.

According to equation \ref{eq:inversion2}, $B'C'=BC\frac{k^2}{AB\cdot AC}$, $B'P'=BP\frac{k^2}{AB\cdot AP}$ and $P'C'=PC\frac{k^2}{AP\cdot AC}$. 
However, $AB$=$BC$=$AC$=$a$. Therefore, $B'C'=\frac{k^2}{a}$, $B'P'=BP\frac{k^2}{a\cdot AP}$ and $P'C'=PC\frac{k^2}{a\cdot AP}$. Rewriting $B'C'$ as $B'C'=PA \frac{k^2}{a \cdot AP}$, we can notice that the sides of triangle $\Delta B'P'C'$ are proportional to the sides of triangle formed with segments $PA$, $PB$, $PC$. The scale factor is $\frac{k^2}{a \cdot AP}$. Therefore, the area of $\Delta B'P'C'$ is proportional to the area of Pompeiu triangle $(PA,PB,PC)$. More specifically, $S(B'P'C')=S(P(PA,PB,PC))(\frac{k^2}{a PA})^2$.

\begin{figure}[h]
\centering
\includegraphics[width=0.5\textwidth]{pompeiuTriangle.png}
\caption{A Pompeiu triangle}
\label{fig:pompeiuTriangles1}
\end{figure}

To express the area of triangles $\Delta AB'P'$ as a function of $S(ABP)$, we can use $S(AB'P')=\frac {AB'\cdot AP' \cdot \sin(B'AP')}{2}$=
$\frac{1}{2} \cdot \frac {k^2} {AB} \cdot \frac {k^2}{AP} \cdot \sin(BAP)$=$\frac {k^4} {AB^2 PA^2} \cdot \frac {AB \cdot AP \cdot \sin(BAP)} {2}$=$\frac {k^4}{a^2 PA^2} S(ABP)$. Similarly, $S(AP'C')=\frac {k^4}{a^2 PC^2} S(APC)$. 
\end{document}

### 2023-10-18 mathPaper/pompeiuTriangles.tex
vector updates

@@
 \centering
-\includegraphics[width=0.5\textwidth]{pompeiuTriangle.png}
+\includegraphics[width=0.75\textwidth]{pompeiuTriangle.png}
 \caption{A Pompeiu triangle}
@@
 $\frac{1}{2} \cdot \frac {k^2} {AB} \cdot \frac {k^2}{AP} \cdot \sin(BAP)$=$\frac {k^4} {AB^2 PA^2} \cdot \frac {AB \cdot AP \cdot \sin(BAP)} {2}$=$\frac {k^4}{a^2 PA^2} S(ABP)$.
-Similarly, $S(AP'C')=\frac {k^4}{a^2 PC^2} S(APC)$.
+Similarly, $S(AP'C')=\frac {k^4}{a^2 PA^2} S(APC)$.
+
+As $S(P(PA,PB,PC))$ depends on $S(ABP)$, $S(APC)$, $S(ABC)$ and $MA$, it follows that a simple but general way of expressing $S(ABP)$, $S(APC)$, and $AP$ in terms of the side of the equilateral triangle $a$ and $MA$ needed.
+This can be accomplished through elementary vector analysis (Weatherburn, 1926).
+
+\subsection{Elements of vector analysis}
+
+The theory of vectors was developed in the second half of the 18th in response to the need to rigorously and efficiently investigate problems in physics.
+As many physically properties such as forces, velocities, 3D displacements, etc.
+can be represented by vectors, cannot be described by single numbers (scalars), the theory of vectors was developed to provide a mathematical framework to study these properties.
+For example, a force cannot be described by a single number related to its magnitude because its direction is important as well.
+Similarly, position of a point relative to another cannot be described by the distance between them because the orientation relative to a given direction is a defining factor as well.
+However, the direction defined as value of an angle is not useful because it precludes the use of algebraic operations such as addition and subtraction.
+Therefore, instead of angular values, vectors rely on the concept of decomposition along two independent directions.
+To describe the decomposition of a vector along two independent, we need to first define vectors and the addition of two vectors.
+In geometry, vectors are just directed segments.
+For example, vector $\vec {AB}$ in Figure \ref{fig:vectors} is just segment $AB$ oriented from $A$ to $B$.
+The addition of two vectors is defined as the vector obtained by placing the second vector at the end of the first vector.
+For example, the sum of vectors $\vec {AF}$ and $\vec {FB}$ is vector $\vec {AC}$.
+Equation
+\begin{equation}
+\vec {AB}=\vec {AF} + \vec {FB}
+\label{eq:vectorAddition}
+\end{equation}
+is called the triangle law of addition and is true irrespective of where $F$ is located in the plane.
+However because $F$ is on line $d1$ and $FB$ is parallel to line $d2$, we can say that vectors $\vec{AF}$ and $\vec{FB}$ are the components of $\vec{AB}$ along directions $d1$ and $d2$.
+The decomposition of a vector along two independent directions is unique.
+This means that if we know the decomposition of a vector along two independent directions, we can determine the vector.
+Note that a different vector, i.e.
+$\vec{AC}$ with same magnitude as $\vec{AB}$, but different direction, has different components on $d1$ and $d2$.
+In short, to define a vector, we need to specify its components along two independent (arbitrary but defined) directions.
+The determination of a vector from its components and the inverse operation are trivial.
+
+In addition to decomposition and vector addition, another important operation is its multiplication by a scalar.
+The multiplication of a vector by a scalar is defined as the vector obtained by multiplying the magnitude of the vector by the scalar and keeping the direction of the vector.
+For example, the vector $2\vec{AB}$ is twice as long as $\vec{AB}$ and has the same direction.
+If we multiply by a negative scalar, in addition to changing the magnitude of the vector, we also invert its direction.
+
+An extremely important vector operation is the dot product between two vectors $\vec{a}$ and $\vec{b}$, defined as 
+\begin{equation}
+\vec{a}\cdot\vec{b}=|a||b|\cos(\theta)
+\label{eq:dotProduct}
+\end{equation}
+where $|a|$ and $|b|$ are the magnitudes of vectors $\vec{a}$ and $\vec{b}$ and $\theta$ is the angle between them.
+While extremely important, the significance of the dot product is not immediately obvious.
+A major reason for the introduction of the dot product is its utility in deriving the components of a vector with respect to two new directions $d1_{new}$ and $d2_{new}$ from its components with respect to current directions $d1$ and $d2$.
+Another benefit is in easily calculating distances, as it will be obvious in the article.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.75\textwidth]{Vectors.png}
+    \caption{Two vectors and their decompositions.}
+    \label{fig:vectors}
+    \end{figure}
+\section{Reference}
+Weatherburn, Charles Ernest. Elementary Vector Analysis: With application to geometry and physics. Bell and Sons, 1926.
 \end{document}

### 2023-10-18 mathPaper/pompeiuTriangles.tex
centroid

@@
 %% insert figure 1 here
+\subsection{Some preliminary results}
 For a given equilateral triangle $\Delta ABC$, we can apply an inversion with respect to a reference circle with center $A$ and radius $k$, and transform points $A$ and $B$ into point $A'$ and $B'$.
@@
 Similarly, $S(AP'C')=\frac {k^4}{a^2 PA^2} S(APC)$.
+
+On the other hand, $S(AB'C')=\frac {k^4}{a^4}S(ABC)$.  Therefore,
+
+\begin{equation}
+    S(P(PA,PB,PC))=S(ABP)+ S(APC)-\frac{PA^2}{a^2}S(ABC)
+    \label{eq:area2}
+\end{equation}
 
@@
     \end{figure}
+
+Although not directly used in this article the cross product between two vectors $\vec{a}$ and $\vec{b}$ is also important. We mention it here without a definition.
+
+A concept that is used in the article to derive various formula used in the simplification of Equation \ref{eq:area2} is that of centroid.
+The centroid of a system of points $P_1$, $P_2$, ..., $P_n$ associated with masses $m_1$, $m_2$, ..., $m_n$ is the point $G$ defined by the equation
+\begin{equation}
+\vec{OG}=\frac{1}{\sum_{i=1}^n m_i}\sum_{i=1}^{n}m_i\vec{OP_i}
+\label{eq:centroid}
+\end{equation}
+where $G$ is the centroid location and $O$ is the origin of the coordinate system.
+
+It can be proven that the location of $G$ does not depend on that of $O$ (Weatherburn, 1926).
+
+\subsection{Some useful formulas}
+Consider an arbitrary triangle $ABC$ to which we attach masses $k_1$, $k_2$ and $k_3$.
+According to Eq.
+\ref{eq:centroid} when the origin is point $A$, the centroid of the system consistent of masses $k_2$ and $k_3$ is
+
+\begin{equation}
+\vec{AD}=\frac{1}{k_2+k_3}(k_2\vec{AB}+k_3\vec{AC})
+\end{equation}
+If we denote the centroid of the system of masses $k_1$, $k_2$ and $k_3$ by $P$, then
+
+\begin{equation}
+\vec{AP}=\frac{1}{k_1+k_2+k_3}(k_2\vec{AB}+k_3\vec{AC})
+\label{eq:centroid2}
+\end{equation}
+Note that $\vec{AP}$ and $\vec{AD}$ are both proportional to $k_2 \vec{AB}+ k_3\vec{AC}$.
+They have the same origin and same direction, but they have different magnitudes (on lengths).
+The ratio of their magnitudes is $\frac{k_1+k_2+k_3}{k_2+k_3}$.
+Therefore, $P$ is located on $AD$ at a distance $\frac{k_2+k_3}{k_1+k_2+k_3} AD$ from $A$.
+Moreover, $P$ is located at a distance of $\frac{k_1}{k_1+k_2+k_3} AD$ from $D$.
+It follows that the area of triangle $\Delta BPC$ is $\frac{k_1}{k_1+k_2+k_3} S(ABC)$.
+Similarly, the area of triangle $\Delta APC$ is $\frac{k_2}{k_1+k_2+k_3} S(ABC)$ and the area of triangle $\Delta APB$ is $\frac{k_3}{k_1+k_2+k_3} S(ABC)$.
+
+The converse is also true.
+That is, given a point $P$ in the interior of a triangle, there are masses $k_1$, $k_2$, $k_3$ that can be attached to points $A$, $B$, $C$ such that the centroid of the system is $P$.
+The masses are given by $k_1=\frac{S(BPC)}{S(ABC)}$, $k_2=\frac{S(APC)}{S(ABC)}$ and $k_3=\frac{S(APB)}{S(ABC)}$.
+
+\subsection{Derivation of the formula for the area of a Pompeiu triangle}
+Equation \ref{eq:area2} can be then rewritten as
+
+\begin{equation}
+S(P(PA,PB,PC))=S(ABC)\left(\frac{k_2+k_3}{k_1+k_2+k_3}-\frac{PA^2}{a^2}\right)
+\label{eq:area3}
+\end{equation}
+
+To further simplify Eq. \ref{eq:area3}, we can calculate the dot product of $\vec{AP}$ with itself.  This is given by
+\begin{equation}
+\vec{AP}\cdot\vec{AP}=\frac{1}{(k_1+k_2+k_3)^2}(k_2\vec{AB}+k_3\vec{AC})\cdot(k_2\vec{AB}+k_3\vec{AC})
+\label{eq:dotProduct2}
+\end{equation}
+Expanding the dot product in Eq. \ref{eq:dotProduct2} and using the fact that $\vec{AB}\cdot\vec{AB}=a^2$, $\vec{AC}\cdot\vec{AC}=a^2$ and $\vec{AB}\cdot\vec{AC}=\frac{a^2}{2}$, we obtain
+\begin{equation}
+AP^2=\frac{1}{(k_1+k_2+k_3)^2}(k_2^2a^2+k_3^2a^2+k_2k_3a^2)
+\label{eq:dotProduct3}
+\end{equation}
+\begin{figure}
+\centering
+\includegraphics[width=0.75\textwidth]{centroid.png}
+\caption{An arbitrary triangle $ABC$ with masses $k_1$, $k_2$, $k_3$ attached at points $A$, $B$, $C$.}
+\label{fig:centroid}
+\end{figure}
+
+Therefore, Eq. \ref{eq:area3} can be rewritten as
+\begin{equation}
+S(P(PA,PB,PC))=S(ABC)\left(\frac{k_2+k_3}{k_1+k_2+k_3}-\frac{k_2^2+k_3^2+k_2k_3}{(k_1+k_2+k_3)^2}\right)
+\label{eq:area4}
+\end{equation}
+or
+\begin{equation}
+S(P(PA,PB,PC))=S(ABC)\left(\frac{k_1k_2+k_1k_3+k_2k_3}{(k_1+k_2+k_3)^2}\right)
+\label{eq:area5}
+\end{equation}
+Benyi and Casu (2009) derived a formula similar Eq.
+\ref{eq:area5} also relying on inversions.
+However, they do not use vector geometry and the steps from Eq.
+\ref{eq:area2} to the final form are significantly more algebraic in nature.
+Moreover, the algebraic steps are not as intuitive as the vector geometry steps presented here and do not enable the derivation of the formula for the distance of point $P$ to the circumcircle of $ABC$.
+
+\subsection{A remarkable property of Pompeiu triangles}
+In Andreescu et al.
+(2009), the reader is asked to find the geometric locus of points P that have the same areas.
+The original author of the problem is unknown, but the problem is believed to have been known for quite a while before it was published in the book.
+
 \section{Reference}
+Andreescu, T., Gelca, R. and Saul, M., 2009. Mathematical olympiad challenges (Vol. 216). Boston, MA: Birkhäuser.
+Bényi, Á. and Caşu, L., 2009. Pompeiu's Theorem Revisited. The College Mathematics Journal, 40(4), pp.252-258.
 Weatherburn, Charles Ernest. Elementary Vector Analysis: With application to geometry and physics. Bell and Sons, 1926.

### 2023-10-18 mathPaper/pompeiuTriangles.tex
final

@@
 
+Using Eq. (\ref{eq:centroid2}) and the equivalents for $\vec{BP}$ and $\vec{CP}$, after multiplication by $k_1$,$k_2$, and $k_3$ and summation, we obtain
+
+\begin{equation}
+k_1\vec{AP}+k_2\vec{BP}+k_3\vec{CP}=0
+\label{eq:psum}
+\end{equation}
+
+However, $\vec{AP}=\vec{AO}+\vec{OP}$, $\vec{BP}=\vec{BO}+\vec{OP}$ and $\vec{CP}=\vec{CO}+\vec{OP}.$, where $O$ is the center of the circumcircle of $\Delta ABC$.
+Therefore, Eq.
+(\ref{eq:psum}) can be rewritten as
+
+\begin{equation}
+k_1\vec{AO}+k_2\vec{BO}+k_3\vec{CO}=-(k_1+k_2+k_3)\vec{OP}
+\label{eq:psum2}
+\end{equation}
+
+Taking the dot product of Eq. (\ref{eq:psum2}) with itself, we obtain
+\begin{equation}
+\begin{split}
+k_1^2 AO^2+k_2^2 BO^2 +k_3^2 CO^2+\\
+2k_1k_2\vec{AO}\cdot\vec{BO}+2k_1k_3\vec{AO}\cdot\vec{CO}+2k_2k_3\vec{BO}\cdot\vec{CO}=\\
+(k_1+k_2+k_3)^2 OP^2
+\end{split}
+\label{eq:psum3}
+\end{equation}
+
+However, $AO=BO=CO=R$, where $R$ is the radius of the circumcircle of $\Delta ABC$, and 
+$\vec{AO}\cdot\vec{BO}=\vec{BO}\cdot\vec{CO}$=
+$\vec{CO}\cdot\vec{AO}=$
+$-\frac{1}{2}R^2$. Therefore, Eq. (\ref{eq:psum3}) can be rewritten as
+\begin{equation}
+OP^2=\frac{k_1^2+k_2^2+k_3^2}{(k_1+k_2+k_3)^2}R^2-\frac{k_1k_2+k_1k_3+k_2k_3}{(k_1+k_2+k_3)^2}R^2
+\label{eq:psum4}
+\end{equation}
+or
+\begin{equation}
+    OP^2=\left(1-3\frac{k_1k_2+k_1k_3+k_2k_3}{(k_1+k_2+k_3)^2}\right) R^2
+    \label{eq:psum5}
+\end{equation}
+Therefore if $S(P(PA,PB,PC))$ is constant, $\frac{k_1k_2+k_1k_3+k_2k_3}{(k_1+k_2+k_3)^2}$ is also constant.
+Consequently, $OP$ is constant.
+This means that the locus of points $P$ with constant area is a circle centered at $O$ with radius $OP$.
+
 \section{Reference}
-Andreescu, T., Gelca, R. and Saul, M., 2009. Mathematical olympiad challenges (Vol. 216). Boston, MA: Birkhäuser.
-Bényi, Á. and Caşu, L., 2009. Pompeiu's Theorem Revisited. The College Mathematics Journal, 40(4), pp.252-258.
+
+Andreescu, T., Gelca, R. and Saul, M., 2009. Mathematical olympiad challenges (Vol. 216). Boston, MA: Birkhäuser.\\
+Bényi, Á. and Caşu, L., 2009. Pompeiu's Theorem Revisited. The College Mathematics Journal, 40(4), pp.252-258.\\
 Weatherburn, Charles Ernest. Elementary Vector Analysis: With application to geometry and physics. Bell and Sons, 1926.

### 2023-10-18 mathPaper/pompeiuTriangles.tex
generalized

@@
 
+Note that Eq.
+(\ref{eq:psum4}) is valid for any triangle $ABC$ and any point $P$ in the interior of the triangle.
+If triangle $ABC$ is not equilateral, one can use formula $\cos(2x)=1-2\sin(x)^2$ along with the law of sines to derive a more general formula for the distance of $P$ to the circumcircle of $ABC$, i.e.
+
+\begin{equation}
+OP^2=R^2-\frac{k_1k_2c^2+k_1k_3b^2+k_2k_3a^2}{(k_1+k_2+k_3)^2}
+\label{eq:psum6}
+\end{equation}
+where $a$, $b$ and $c$ are the sides of triangle $ABC$.
+Combined with the formula for the area of a generalized Pompeiu triangle (Benyi and Casu 2009), this formula can be used to show that the locus of a point associated with a generalized Pompeiu triangle with constant area is still a circle with the same center as the circumcircle of $\Delta ABC$.
+The derivation of the formula for the area of a generalized Pompeiu triangle using the approach described in the previous sections for a regular Pompeiu triangle is straightforward and requires the extension of Eqs.
+(\ref{eq:area2}) and (\ref{eq:centroid2}) to their general forms.
+
 \section{Reference}

### 2023-11-09 RST/Submission/responseReviewers_Rev2.3.tex
revision 3

\documentclass[12pt]{article}
\usepackage{amsmath,amssymb,amsthm}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{longtable}
\usepackage[margin=1in]{geometry}
\author{Mircea Grecu and John E. Yorks}
\title{Synergistic retrievals of ice in high clouds\\
Response to Reviewers\\Second Revision}
\date{}
\begin{document}
\maketitle

\noindent \textbf{Reviewer 1}\\

\noindent\textit{The authors should state which version of the 2C-ICE and 2B-GEOPROF products they are using (R05?). Other than that, I think this is a good paper that is 
ready for publication.}\\

\noindent We thank the reviewer for the positive feedback. We added a sentence to the manuscript to clarify that we use the R05 version of the 2C-ICE and 2B-GEOPROF products.\\

\noindent \textbf{Reviewer 2}\\

\noindent\textit{Although I would prefer not to drag out this review process even longer, I feel that the authors failed to adequately address some of my previous comments. Most importantly, the published code does not contain the retrieval you describe in your paper. The repository only includes a retrieval based on a similarity search, which yields better results than the results presented in the manuscript. If these results are representative, they do put much of the presented results into question. I, therefore, suggest the authors provide the actual code used to produce the results presented in the manuscript instead.}\\

\noindent We thank the reviewer for the perspective.\\

\noindent \textbf{Other comments:}\\

\noindent\textit{\textbf{1.}Please note that the sentence above clearly states that $N_w$ depends on $a_{norm}$. I do understand that, in general, $N_w$ does not depend on $a_{norm}$. In your study, however, you are trying to infer both $N_w$ and $a_{norm}$ from the Z-IWC relationship in CS 2C-ICE. You describe how you use these values to derive the variation of $ln(a)$ and b, but that is still insufficient to determine $N_w$ and its altitude variation. Please provide the final parametrization of $N_w$ that you use in your simulations --prior to perturbing $N_w$ -- and how you calculated it. Currently, the manuscript only states that you calculate  Z for a fixed $N_w$, but it is unclear how you use that to determine $a_{norm}$.}\\

\noindent We are sorry about the confusion. Parameter $a_{norm}$ is derived using analytical normalized gamma distributions with $N_w$=0.08cm$^{-4}$. The use of analytical distributions in the derivation of the scattering lookup tables is already explained in the manuscript, but their relevance to $a_{norm}$ is insufficiently described. We added a few sentences to the manuscript to clarify this point.  Specifically, given that normalized particle size distributions are functions of type $N(D)=N_w\cdot F(D/D_m)$, normalization by $N_w$ transforms the relation between normalized $IWC$ and $Z$ to a relation between $IWC_{norm}=\int_0^{\infty} F(D/D_m)m(D)dD$ and $Z_{norm}=\frac {\lambda ^4} {\pi ^5 |K_w|^2} \int_0^{\infty} F(D/D_m) \sigma_b(D) dD$, where $m(D)$ and $\sigma_b(D)$ are the mass and electromagnetic backscattering cross-section of particles of mass $D$. As apparent in the above integrals, there is no $N_w$ dependence in the relations between normalized variables. A relationship of the type $IWC_{norm}=a_{norm}Z_{norm}^b$ becomes $IWC/N_w=a_{norm}(Z/N_w)^b$ for a PSD with a scaling intercept of $N_w$, or $IWC=a_{norm}N_w^{1-b}Z^b$. The values associated with our scattering lookup tables are $a_{norm}$ $\approx$0.985 and b$\approx$0.64, which result in parameterization $log_{10}(N_w)=8.87+0.161H$.\\

\noindent\textit{\textbf{2.}
$>$ l. 347: The new distribution of retrieved gravity centers cannot be described as narrow.\\
You did not change anything in the manuscript.}\\

\noindent We apologize for the omission. We certainly intended to, but did not keep proper track of all changes meant to be included in the submission. We added a sentence to the manuscript to clarify this point.\\

\noindent\textit{Minor mistakes:\\
l. 94: Formatting of reference\\
l. 108: Space after full stop.\\
l. 121: Spelling mistake\\
l. 361: There $\->$ The\\
l. 441: 'Of' repeated\\
l. 462: Is $\->$ are.} \\

\noindent We thank the reviewer for pointing out these issues. We addressed them in the new version of the manuscript. It should be mentioned though that reference formatting is determined by the Latex compiler at submission time and some formatting aspects are not under the author's control.\\

\begin{thebibliography}{999}

\bibitem{arlot2010survey}
Arlot, S., A survey of cross-validation procedures for model selection. Statistics Surveys. 2010, 4, 40-79.

%\bibitem{ferreira2001}
%Ferreira, F., P. Amayenc, S. Oury, and J. Testud, 2001: Study and Tests of Improved Rain Estimates from the TRMM Precipitation Radar. J. Appl. Meteor. Climatol., 40, 1878–1899, https://doi.org/10.

\end{thebibliography}

\end{document}

### 2023-11-09 RST/Submission/RSTManuscript_rev3.tex
revision 3

--- 2023-08-22 RST/Submission/RSTManuscript_rev2.tex
@@
 This ensures the consistency between the particle distribution assumptions and the electromagnetic scattering properties used in the CS reflectivity processing and those used the simulation of the lidar, Ku-band radar and radiometer observations.
-Our CS-based ice product is optimized to be consistent with the synergistic Cloudsat and CALIPSO Ice Cloud Property Product (2C-ICE) of \cite{deng2015}.
+Our CS-based ice product is optimized to be consistent with the synergistic Cloudsat and CALIPSO Ice Cloud Property Product (2C-ICE) version R05 of \cite{deng2015}.
 When the CALIPSO lidar detects echo associated with clouds but the CS radar signal is below the noise level, we use the 2C-ICE product to extend our CS-based estimates.
@@
 \end{equation}
-(where $a_{norm}$ and $b$ are constants) explains almost perfectly the relationships between IWC and Z calculated from observed PSDs.However, equation (\ref{eq:1}) is not  sufficient to derive accurate, unbiased estimates of ice water contents, because $N_w$ varies considerably in time and space.
-Nevertheless, multiple studies showed that it is beneficial to parameterize $N_w$ as a function  of various variables, such as temperature \citep{hogan2006retrieval, delanoe2008variational,deng2010tropical}, rather than using $N_w$ independent reflectivity ice relations.
-In this study, we parameterize $N_w$ as a function of temperature based on the CloudSat 2C-ICE product \citep{deng2010tropical,deng2013evaluation}.
+(where $a_{norm}$ and $b$ are constants) explains almost perfectly the relationships between IWC and Z calculated from observed PSDs.
+Parameters $a_{norm}$ and $b$ are usually derived from regression of $IWC/N_w$ against $Z/N_w$ in the log-log space.
+Given that normalized PSDs are functions of type $N(D)=N_w\cdot F(D/D_m)$, normalization by $N_w$ removes any $N_w$ dependence from $IWC/N_w$ against $Z/N_w$ regressions, as both  $IWC/N_w$ and $Z/N_w$ are exclusive functions of $F(D/D_m)$.
+In this study, we derive $a_{norm}$ and $b$ from IWC and associated Z calculated from analytical normalized PSDs.
+While details are present in subsequent paragraphs, it should be mentioned that the actual values of $a_{norm}$ and $b$ are 0.985 and 1.64, with $a_{norm}$ valid when the units of IWC, $N_w$ and $Z$ are g/m$^3$, m$^{-4}$, and dBZ respectively.
+However, equation (\ref{eq:1}) is not  sufficient to derive accurate, unbiased estimates of ice water contents, because $N_w$ varies considerably in time and space.
+Nevertheless, multiple studies showed that it is beneficial to parameterize $N_w$ as a function  of various variables, such as temperature or height \citep{hogan2006retrieval, delanoe2008variational,deng2010tropical}, rather than using $N_w$ independent reflectivity ice relations.
+In this study, we parameterize $N_w$ as a function of height based on the CloudSat 2C-ICE product \citep{deng2010tropical,deng2013evaluation}.
 A scatter-plot analysis of relationships between the 2C-ICE IWC and the associated reflectivity suggests that the multiplicative coefficient in a power law ice-reflectivity relationship is parameterized as a function of temperature (or equivalently, as a function of height relative to the freezing level) in the default 2C-ICE retrievals.
 The default multiplicative coefficient (the value that provides the IWC estimate prior to the ingestion of the lidar observations) may be simply estimated by regressing the 2C-ICE IWC on the associated CloudSat reflectivity as a function of $H$, defined as height relative the freezing level.
-Specifically, a regression of the type $ln(IWC)=ln(a)+b*ln(Z)$ is performed for each bin referenced relative to the zero degree bin is performed, and the variation of $ln(a)$ is parameterized as a function of the relative height.
+Specifically, a regression of the type $\log_{10}(IWC)=\log_{10}(a)+b\cdot\log_{10}(Z)$ is performed for each bin referenced relatively to the zero degree bin is performed, and the variation of $\log_{10}(a)$ is parameterized as a function of the relative height $H$.
 The results are shown in Figure \ref{f2}.
@@
 We, therefore, use the data in Figure \ref{f2} to parameterize $N_w$ as a function of the relative height.
-Specifically, given that $(1-b)ln(N_w)+ln(a_{norm})=ln(a)$ and $ln(a)=ln(a_0)+s*H$, with $a_0$ and $s$ the intercept and the slope of a regression for the data in Figure \ref{f2}, we can express $N_w$ as a function of the relative height, slope $s$, and parameters $a_0$ and $a_{norm}$.
+Specifically, given that $(1-b)\log_{10}(N_w)+\log_{10}(a_{norm})=\log_{10}(a)$ and $\log_{10}(a)=\log_{10}(a_0)+s\cdot H$, with $a_0$ and $s$ the intercept and the slope of a regression for the data in Figure \ref{f2}, we can express $N_w$ as a function of the relative height, slope $s$, and parameters $a_0$ and $a_{norm}$.
+The actual expression derived in this study is $\log_{10}N_w=8.87+0.161\cdot H$ with $N_w$ expressed in m$^{-4}$ and $H$ expressed in km.
 
@@
 Shown in Figure \ref{f6} is the frequency of IWC gravity center estimated from radiometer-only observations as a function of its true value.
-It may be observed in the figure that while the true IWC gravity center exhibits quite a broad distribution, the one retrieved from the radiometer-only observations exhibits a multimodal narrow distribution.
-Moreover, there correlation between the retrieved and the true IWC gravity center is rather low.
+It may be observed in the figure that while the true IWC gravity center exhibits quite a broad distribution, the one retrieved from the radiometer-only observations is characterized by a narrower range with most values occurring between 2.0 km above the freezing level to about 5.5 km above the freezing level.
+Moreover, the correlation between the retrieved and the true IWC gravity center is rather low.
 This is another indication that, while the total amount of ice may be reasonably estimated from radiometer-only observations, its vertical distribution can not be accurately determined from radiometer-only observations.
@@
 It may be observed in the table that the performance of the estimates based on all instruments is significantly better than those based on radiometer-only observations.
-Furthermore, the inclusion of the lidar observations in the retrieval process has a larger impact on the retrieval performance than the inclusion of the radar
-observations.
+Furthermore, the inclusion of the lidar observations in the retrieval process has a larger impact on the retrieval performance than the inclusion of the radar observations.
 This is expected since the lidar observations are able to provide information about the top of the clouds, where the radar observations are above the noise level only occasionally.
@@
 
-Additional insight may be gained by examining the distribution of the retrieved of IWP and IWC gravity center as a function of their true values.
+Additional insight may be gained by examining the distribution of the retrieved IWP and IWC gravity center as a function of their true values.
 These distributions are show in Figure \ref{f11} for the lidar-radiometer and radar-radiometer retrievals.
@@
 However, neither of the two instrument combinations is able to provide estimates that are as accurate as those derived from all instruments.
-Moreover, the lidar-radiometer IWP retrievals appear to be less accurate than lidar-only retrievals.
+Moreover, the lidar-radiometer IWC gravity center retrievals appear to be less accurate than lidar-only retrievals.
 This may be an artifact of statistical inversion methodology, which may make suboptimal use of correlations in the observations and improve the accuracy of the estimates for some type of profiles at the expense of others.
@@
 % NRMS & 0.73 &  0.59  &  0.32 & 0.32 \\
-NRMS & 0.84	& 0.67 & 0.64 & 0.61 &	0.56 &	0.54 &	0.48 \\
+IWC NRMS & 0.84	& 0.67 & 0.64 & 0.61 &	0.56 &	0.54 &	0.48 \\
+IWP NRMS & 1.12 & 0.54 & 0.4  & 0.48 &  0.37 &  0.32 &  0.3  \\
+GC NRMS  & 0.21 & 0.58 & 0.46 &  0.26 &  0.23 &  0.42 &  0.19 \\
 %0.73 &  0.59  &  0.32 & 0.22 \\
@@
 A hold-out validation methodology is used to assess the accuracy of the retrieved IWC profiles from various combinations of observations from the three instruments.
-Specifically, the IWC and associated observations is randomly divided into two datasets, one for training and the other for evaluation.
+Specifically, the IWC and associated observations are randomly divided into two datasets, one for training and the other for evaluation.
 The training dataset is used to train the retrieval algorithm, while the evaluation dataset is used to assess the retrieval performance.
@@
 \datastatement
-The CloudSat data can be accessed at: https://www.cloudsat.cira.colostate.edu/.
+The CloudSat 2C-ICE and 2B-GEOPROF version R05 data can be accessed at: https://www.cloudsat.cira.colostate.edu/.
+Code to simulate the synthetic observations from the CloudSat products and investigate the synergy of the instruments may be accessed at:  https://github.com/mirgrecu/synergIceRetrievals/.
+
 %  The data availability statement is where authors should describe how the data underlying 

### 2023-11-10 RST/Submission/responseReviewers_Rev2.3.tex
updated rev3

@@
 \title{Synergistic retrievals of ice in high clouds\\
-Response to Reviewers\\Second Revision}
+Response to Reviewers\\Third Revision}
 \date{}
@@
 
-\noindent We thank the reviewer for the perspective.\\
+\noindent We thank the reviewer for the perspective.
+We agree, we should have done a better job with the code repository.
+To address that, we included a notebook containing the definition of the classification network and code to estimate the parameters of the ensemble filter.
+Code to train the classification network using the subset of the simulated observations and associated IWC profiles placed in the repository is also included in the file.
+Moreover, the application of the estimation framework is exemplified in the same notebook.
+These being said, we don't think that the inclusion of a somewhat simpler retrieval methodology in the initial code release invalidates our work.
+While a rigorous evaluation may reveal which methodology results is better, it is expected that the difference between them be smaller than the uncertainties in either.
+Moreover, both approaches are based on similarity evaluations, as the ensemble filter is applied after a classification trained based on clustering.
+Currently, testing the forward models and the associated retrievals using real observations is a higher priority.
+We included the similarity search retrieval in the initial release because it was more stimulating and convenient that to clean and test the code for the ensemble filter in a new environment, but we removed it during the revision to avoid potential confusions.\\
 
@@
 A relationship of the type $IWC_{norm}=a_{norm}Z_{norm}^b$ becomes $IWC/N_w=a_{norm}(Z/N_w)^b$ for a PSD with a scaling intercept of $N_w$, or $IWC=a_{norm}N_w^{1-b}Z^b$.
-The values associated with our scattering lookup tables are $a_{norm}$ $\approx$0.985 and b$\approx$0.64, which result in parameterization $log_{10}(N_w)=8.87+0.161H$.\\
+The values associated with our scattering lookup tables are $a_{norm}$ $\approx$0.985 and b$\approx$0.64, which result in parameterization $log_{10}(N_w)=8.87+0.161H$ after combination with the parameterization of $a$ described in the manuscript.\\
 
@@
 \noindent We apologize for the omission.
-We certainly intended to, but did not keep proper track of all changes meant to be included in the submission.
+We certainly intended to but did not keep proper track of all changes meant to be included in the submission.
 We added a sentence to the manuscript to clarify this point.\\
@@
 
-\begin{thebibliography}{999}
+%\begin{thebibliography}{999}
 
-\bibitem{arlot2010survey}
-Arlot, S., A survey of cross-validation procedures for model selection. Statistics Surveys. 2010, 4, 40-79.
+%\bibitem{arlot2010survey}
+%Arlot, S., A survey of cross-validation procedures for model selection. Statistics Surveys. 2010, 4, 40-79.
 
@@
 
-\end{thebibliography}
+%\end{thebibliography}
 
